### 2024-01-28 Findings (Shared Folder)/Chapters/PapersReviewed.tex
new comments from Leah

\chapter{Papers Reviewed}

% Redefine section command for this chapter
\let\oldsection\section
\renewcommand{\section}{\newreviewsection}

% Your custom sections here

In this section I will keep the notes of the papers I have reviewed, or reproduced their results.

\section{Topological data analysis distinguishes parameter regimes in the Anderson-Chaplain model of angiogenesis}{Nardini, Byrne}{2021-PLOS CB}

\subsection{Introduction}
This paper studies the Anderson Chaplain \cite{Anderson1998} model of angiogenesis and partitions the parameter spaces based on the morphology of the vascular structure generated by the model. In other words, let $P = R^d$ be the parameter space of the model, $M$ the space of all possible morphology for the vascular networks. Also, define the equivalence relation $\sim$ defined on the parameter space $P$ to be
\[ \text{for } p_1, p_2 \in P \wh p_1 \sim p_2 \qquad iff \qquad \mathcal{A}(p_1) \equiv \mathcal{A}(p_2), \]
where $\mathcal{A}: P \to M$ a mapping from the parameter space to the morphology space. The $\equiv$ is yet another equivalence relation defined on the morphology space $M$ where for $m_1, m_2 \in M$ we write $m_1 \equiv m_2$ if and only if $m_1$ and $m_2$ has the same topological characterization. These topological characterizations are computed using the topological data analysis techniques.

\subsection{Method}
Chaplain-Anderson model of angiogenesis used in this paper keeps track of the spatio-temporal evolution of three variables: endothelial tip cells, tumor angiogenesis factor, and fibronectin.

Topological data analysis: Two filtration methods were used: sweeping plane method, and flooding filtration. The filtration is performed on the binary images generated with the Chaplain-Anderson model.

\subsection{Useful facts}
\begin{itemize}
	\item The growth factors the cancer cells release when under low nutritient and oxygen: vascular endothelial growth factor (VEGF), platelet derived growth factor (PDGF), and basic fibroblast growth factor (bFGF).
\end{itemize}

\subsection{Points that are not clear yet}
\begin{enumerate}[(a)]
	\item In the introduction, the authors claim that ``The morphology of a vascular network can reveal the presence of an underlying disease, or predict the response of a patient to treatment'', without any citation of explanation. I think this needs more discussion.
\end{enumerate}

\subsection{Useful papers cited}
\begin{itemize}
	\item Papers related to biology of the tumor induced angiogenesis \cite{Gupta2003,Folkman1971}.
	\item More modern descriptions of the angiogenesis \cite{Lugano2020,Saman2020}
	\item The role of the mechanical stress on the angiogenesis \cite{Li2005,Li2002,Vavourakis2017}
	\item Some old and classic models for the angiogenesis \cite{Anderson1998,Balding1985,Byrne1995,Stokes1991}.
	\item More detailed theoretical models for angiogenesis \cite{Byrne2010,Hadjicharalambous2021,Metzcar2019,Scianna2013}.
	\item Alternative models of angiogenesis \cite{Vilanova2017,Stepanova2021,Perfahl2017,Grogan2017,Vavourakis2017,Cai2017,Sefidgar2015}
	\item Statistical and single scale methods to quantify the vascular networks \cite{Perfahl2017,Folarin2010,Kannan2018,Konerding1999,Konerding2001}
	\item Biological angiogenesis experiments \cite{Bauer2007}
	\item Topological data analysis \cite{Carlsson2009,Zomorodian2005}
	\item Applied Algebraic Topology \cite{2020,2023}
	\item Application of topological data analysis \cite{Nicolau2011,Nielson2015,Qaiser2016,Xia2014}
	\item Topological data analysis and agent based models \cite{Topaz2015,McGuirl2020}
\end{itemize}

\subsection{Results}

% Reset section command back to original after the chapter
\let\section\oldsection

### 2024-01-28 Findings (Shared Folder)/Findings.tex
new comments from Leah

%\documentclass[12pt,a4paper]{book}
%\usepackage{commands}
%
%\begin{document}
%
%	
%\end{document}
\documentclass{labdiary}
\usepackage{commands}
\usepackage[numbers]{natbib}

\title{My Findings}
\author{Ali Fele Paranj}
\date{\today}

\begin{document}

	\maketitle

	\tableofcontents

	\include{Chapters/StatisticalPhysicsForBiology.tex}
	\include{Chapters/PapersReviewed.tex}

\section{Leah Comments Jan 24, 2024}
Please change the citation style to Author et al (year) in place of [number], so it is easier to see who you are citing without having to flip to the bibliography. Thanks for linking the bibliography to the URLs of the papers so it's possible to scan them.

When citing papers, it is best if you can also say 1-2 sentences about those papers, even based on their abstract (in your own words, of course, never copied directly). For example:

Byrne (2010) reviews theoretical cancer models and demonstrates the advantages of colaboration between modelers and experimentalists.

\subsection{Suggested research style and flavour}

Since it is unlikely that we will get data for the detailed mechanochemical mechanisms for blood vessel growth in the prostate tumors, it makes sense to (a) start simple from very minimal models that can be linked to data and (b) avoid introducing variable that we have no hope of measuing in the obtained data. My understanding is that (for now) we will have to make do with at best some bulk properties of the blood vessels, so models with a lot of detail will hardly help us.

Here is a possible minimalistic stepwise approach, where we start very simple and gradually build up more detail, starting with simple assumptions.

\textbf{Definitions:}

$n(t)$= density of tip cells in area of interest, (number per unit area)

$\rho(t)$ = density of blood vessels (length per unit area),

$c(t)$ = concentration of drug delivered to region by blood vessels

\subsection{Step 1: Bulk model}

Ignoring spatial structure, we only track the density of vessels. Assume everything is spatially uniform, so there will be no spatial derivatives to consider. We construct an ODE model, and make an elementary assumption.

\textbf{Assumption 1:} The drug is delivered by diffusion from the capillaries into the tissue. Hence, as a rough approximation, and (for now) neglecting the detailed structure of the vessel network, the amount of drug delivered to the region per unit time is proportional to the density of the blood vessels.

\textbf{Step 1a: Elementary model:}

Assume that tips extend at some rate $v$ (units of length/time), creating additional length of capillaries as they extend. Assume capillaries may also have some loss rate $\delta$ (per unit time). Write down an ODE for the rate at which capillary density changes with time.

\begin{subequations}
\label{eq:SimplestModel}
\begin{equation}
\label{eq:vesseldensODE}
 \frac{d \rho}{dt}= ??   
\end{equation}

Assume that new tip cells are created by branching along sides of vessels (or possibly by splitting of existing tip cells) at a rate $\beta$ per unit length per unit time, and that tips disappear when they reconnect to a capillary at some rate $\kappa$ to form a loop. [Note: reconnection requires the interaction of tips with capillaries, and would thus be handled as mass action term. What are the units of $\kappa$?]
Write down an equation for the rate of change of tip density.
\begin{equation}
\label{eq:tipdensODE}
 \frac{d n}{dt}= ??   
\end{equation}

Complete the ODE model equations \eqref{eq:SimplestModel}. Analyse the model so far by determining the steady state densities $\rho_{ss} , n_{ss}$, and how they depend on the parameters $v,\delta,\beta, \kappa$. Determine stability of SS. Create a phase plane diagram that shows the expected dynamics. Simulate the simple ODE system assuming some values of the parameters.

We made the assumption that drug delivery is roughly proportional to the vessel density. Write down an approximate ODE for concentration of drug in the region. 
\begin{equation}
\label{eq:drugODE}
 \frac{d c}{dt}= ??   
\end{equation}	
\end{subequations}Explain how this level of drug depends on the vessel branching and growth parameters.

So far, the blood vessels affect the drug but not the other way around.

\bigskip

\textbf{Step 1b: Coupling vessel dynamics to drug}

Consider how the level of drug might affect the vessel parameters (branching or growth rate or death rate, etc). This will introduce feedback from the drug to the vessel density.

Write down one or two variants of such a model and analyse them fully (including steady states, simulations, and some interpretation of what it means for overall treatment of the tissue.)

Note that the drug dynamics would be fast on the timescale of vessel growth, so there is some time-scale separation that you can take advantage of.

\subsection{(Optional) Step 2: Simple spatially distributed 1D system}
We continue with simplest model but now take spatial growth of vessels into a region. So we consider $\rho(x,t), n(x,t), c(x,t)$ as variables of interest.
We make the same assumptions as above, but now we take into account the fact that there is a flux of tips growing into a region,
\[
J=nv.
\]
Explain why this is a flux.
The equations will be modified to form PDEs. Use the 1D balance equation to create that equation for $n$. Explain whether you need to add any spatial derivatives to the equation for $\rho$. The drug diffusion in the spatial variable will also introduce spatial derivatives in the equation for $c$. Write down the modified 1D spatial model. Note that we do not assume anything like chemotaxis or other fancy mechanisms for the tip motion at this point.

\begin{subequations}
\label{eq:SpatialtModel}
\begin{align}
\frac{\partial \rho}{\partial t }&= ?\\
\frac{\partial n}{\partial t }&= ?\\
\frac{\partial c}{\partial t }&= ?
\end{align}
\end{subequations}

Remark: see above for timescale separation.

\textbf{Step 2a: Analysis of wave of invasion}

Consider looking for traveling wave solutions of the $\rho,n$ system on its own to ask how blood vessels spread along a 1D direction and invade a tissue. (Write down ODEs by transforming variables to $z=x-ct$ where $c$ is wave speed, then analyze existence of traveling was in the $\rho n$ phase plane. See one of my books or ask Jack Hughes for help if you are not yet familiar with this idea.)

\textbf{Step 2b: Simulations}
For simulations of the whole system: You will need to assume some boundary conditions on $n$ and on $c$, as well as some initial distribution in order to simulate this system.

\subsection{Step 3: An agent-based (CPM) model:}
Look up the simplest work on Merks and Rens and co and find their CPM model. Ask whether a Morpheus xml file already exists for this model (can ask the Morpheus team or Merks). If not, create one.

Set up this model and adapt it to describing a simple branching vessel structure, similar to what we have above.

ADD: assume that the cells in this network ``secrete'' drug that then diffuses into the tissue and has some decay time. Find ways of plotting properties of the vessels and the drug concentration.

Here you can get creative, and assume that the tip cell growth etc are affected by drug level, etc. (Again, time scale separation is important.)

Your role will be to extend the Merks model to include this drug aspect.

NOTE: some of Merk's work includes the dynamics of an ECM. I would suggest to avoid extending the model with such a dynamic variable, and to assume instead, that it is a static field or vector-field that affects the rate or direction of tip cell motion.

\subsection{Step 4: Look for data}

This can be done in parallel with other steps: look for specific data on blood vessel density in normal and cancerous tissue. There may be animal studies in which the vessel density is tracked over time.

Find if there is data that we can use to help constrain any of these simple models.

For sure it's easier to find bulk vessel density than to find its spatial distribution and the chemical factors liek VEGF that are modeled in some papers.

\subsection{Step 5: More details and other variants}

You can later (after all the early steps) extend and improve the model in various ways. Some suggestions include the following:

\begin{itemize}
    \item Write down an equation for the number of loops that accumulate as tips reconnect to blood vessels (extend simple model).
    \item Find a way to associate these with ``tortuosity'' of vessel network that could affect its conductivity of drug to tissue.
    \item Consider some kind of D'Arcy's Law (porous medium) as a measure of how vessel structure can reduce net drug delivery.
    \item Vessels have various radii and sizes. You may want to consider how this affects the model as well as the implications on drug delivery. A PDE model with a distribution of vessel diameters would likely be a bit newer than the above simple branching equations.
\end{itemize}

	%\backmatter
	% Appendices, bibliography, etc.

	% Bibliography
	\bibliographystyle{unsrtnat} % The bibliography style
	\bibliography{References/references} % The filename of the .bib file (without the extension)

\end{document}

### 2024-01-28 Findings (Shared Folder)/Chapters/MeetingLog.tex

\chapter{Meeting log}

\subsection{29 Jan Meeting with Leah}
Things to discuss
\begin{itemize}
	\item Citation with Year-Author style, or Number Style?
\end{itemize}

### 2024-01-28 Findings (Shared Folder)/Chapters/MolecularBiology.tex

\chapter{Molecular Biology}
Here in this chapter, I will be covering the basics of the relevant molecular biology concepts. This chapter will serve as a reference for the biological claims throughout the document, as well as the foundation for the review chapters of my thesis.

\section{Molecular Mechanism of Angiogenesis}
There are two general balancing forces acting on the angiogenesis
\begin{itemize}
	\item Inhibitors:
	\begin{itemize}
		\item endostatin
		\item angiostatin
		\item thrombospondin
	\end{itemize} 
	\item Angiogens
	\begin{itemize}
		\item VEGF: Vascular Endothelial Growth Factors.
		\item bFGF: Basic Fibroblast Growth Factor.
		\item PDGF: Platelet Driven Growth Factor.
	\end{itemize}
\end{itemize}

\subsubsection*{Controlling Capillary Joining Process}
In the following text from \cite{Alberts2002}, there is some vague hints about the mechanisms that are controlling capillary joining to each other.

\begin{quote}
	Observations such as these reveal that endothelial cells that are to form a new capillary grow out from the side of an existing capillary or small venule by extending long pseudopodia, pioneering the formation of a capillary sprout that hollows out to form a tube (Figure 22-25). This process continues until the sprout encounters another capillary, with which it connects, allowing blood to circulate. Endothelial cells on the arterial and venous sides of the developing network of vessels differ in their surface properties, in the embryo at least: the plasma membranes of the arterial cells contain the transmembrane protein ephrin-B2 (see Chapter 15), while the membranes of the venous cells contain the corresponding receptor protein, Eph-B4, which is a receptor tyrosine kinase (discussed in Chapter 15). These molecules mediate a signal delivered at sites of cell-cell contact, and they are essential for the development of a properly organized network of vessels. One suggestion is that they somehow define the rules for joining one piece of growing capillary tube to another.

\end{quote}

\subsubsection*{Formation of tube structures by endothelial cells}
It was one of my main concerns that what is the process in which a single lining of endothelial cells following a tip cell forms a hallow tube (i.e. vessel). The following text from \cite{Alberts2002} explains this clearly. This process has also been described in \cite{angiogenesisYoutube}.
\begin{quote}
	Experiments in culture show that endothelial cells in a medium containing suitable growth factors will spontaneously form capillary tubes, even if they are isolated from all other types of cells (Figure 22-26). The capillary tubes that develop do not contain blood, and nothing travels through them, indicating that blood flow and pressure are not required for the initiation of a new capillary network.

	Endothelial cells in culture spontaneously develop internal vacuoles that appear to join up from cell to cell, giving rise to a network of capillary tubes. These photographs show successive stages in the process. 
\end{quote}

\section{Biological Assays to Study Angiogenesis}

\subsection{Corneal Micropocket Assay}
This is one of the simple and reproducible assays to study angiogenesis in a eye. The process involves introducing growth factors in the eye ball of mouse, and then letting the vascular network to form. This is a video from JOVE explaining the details of the protocol \citep{conealMicroPocketAssayJOVE}

### 2024-01-28 Findings (Shared Folder)/Chapters/StatisticalPhysicsForBiology.tex

\chapter{Statistical Physics and Biology}

\section{Introduction}

The future of medicine, Time magazine, Jan 1999: ``Ring farewell to the century of physics, the one on which we split the atom and turned silicon into computing power. It's time to ring in the century of biotechnology \cite{Sung2018}.'' Despite thee tremendous importance of life science and biotechnology nowadays as the above statements proclaim, at this stage their knowledge appears to be largely phenomenological, and thus undeniably calls for fundamental and quantitative understandings of the complex phenomena. It will be timely to ring in the century of a new physical science to meet this challenge.

The concept of self-organization is a central theme in biology and there are many great review paper on the subject (like \cite{Pochan2021,McManus2016,Whitelam2015}). In short \textbf{biological components self-assemble themselves to function. To perform the biological self-organization, then oftern cross over the energy barriers that seem to be insurmountable in the view point of simple physics. To this end, there are two physical characteristics that feature in the mesoscopic biological systems introduced above. The first one is their aqueous environment and thermal fluctuations therein}. The water has many outstanding properties among all liquids. It has a very high heat capacity, meaning that it can act as a heat reservoir with a negligible temperature change. Also, it has a very high (compared to other liquids) di-electric constant. Because of this, water can reduce electrostatic energy of the interactions to the level of thermal energy. This unique property of water originates from the weak hydrogen bonds between molecules. Although, the hydrogen bounds can be broken with the thermal energy of the environment, it causes long range correlation between water molecules. As a result, the liquid water manifests a quasi-critical state where it responds collectively and sensitively to external stimuli \cite{Sung2018,Wiggins1990,Bagchi2013}. The biological systems in mesoscale characterized by the \textit{soft inter connectivity and weak interactions} may appropriately be called the \textbf{bio-soft condensed matter} \cite{Sung2018}.

\section{Keywords}
Quasi-critical state, collective behaviour, self-assembly, bio-soft condensed matter, Coarse graining, non-equilibrium thermodynamics,

### 2024-01-28 Findings (Shared Folder)/Chapters/modelAttempts.tex

\chapter{Modeling Attempts}

\section{Simple ODE model (First Iteration)}
Here, we develop a model that keeps track of the following variables
\begin{quote}
	$n(t)$= density of tip cells in area of interest, (number per unit area).

	$\rho(t)$ = density of blood vessels (length per unit area).

	$c(t)$ = concentration of drug delivered to region by blood vessels.
\end{quote}
An updating list of model parameters
\begin{itemize}
	\item $v$ [length/time]: The rate at which the tip cells move and extends the blood vessels.
	\item $\delta_v$ [1/time]: The rate at which the vascular structure gets degraded.
	\item $\lambda_s$ [1/time]: Tip cell division rate (splitting rate).
	\item $\lambda_b$ [1/time/length]: Tip cell emerging rate from stalk cells.
	\item $\delta_t$ [1/time]: Tip cell death/deactivation rate.
	\item $\kappa$ [area/length/time]: Re-connection of tip cells to the other capillaries to form loops.  
\end{itemize}

\subsubsection*{Studying dynamics of vessel formation}
\[ \frac{d\rho}{dt} = ??. \]
The active tip cells extend the vascular structure as they move. Assuming the tip cells move at rate $v$, then 
\[ \frac{d \rho}{dt} = vn+ ??. \]
Also, assuming the vascular structure degrades with rate $\delta$ [per unit time], we can add the degradation term 
\[ \boxed{\frac{d\rho}{dt} = vn + \delta_v \rho} . \]

\subsubsection*{Studying the Dynamics of Tip Cells}
Things important in the dynamics of the tip cells
\begin{itemize}
	\item Generation of the tip cells: There are at least two ways for new tip cell generation listed as follows:
	\begin{enumerate}[(i)]
		\item Splitting mechanics: When the tip cells splits new vascular stem gets two heads. This should be proportional to the density of tip cells. The parameter $\lambda_s$ [per unit time] reflects this mechanism.
		\item Branching: New tip cells can form out the the endothelial stalk cells. This process should be proportional to the density of blood vessels. The parameter $\lambda_b$ [per unit time per unit length] reflects this mechanism
	\end{enumerate}
	\item Loss of tip cells
	\begin{enumerate}[(i)]
		\item Death of the tip cells or getting deactivated: Reflected by the parameter $\delta_t$
		\item Joining the other branches of vascular network: When a tip cell reconnects another capillary branch, then they disappear. The parameter $\kappa$ is for this mechanism. Note that the re-connection term is proportional to both number of tips cells, as well as the density of blood vessels. Thus the units of $\kappa$ should be [area/length/time]. 
	\end{enumerate}
	\item The movement of tip cells and formation of new vascular networks along the way.
\end{itemize}

\[ \boxed{\frac{dn}{dt} = (\lambda_s - \delta_t) n + \lambda_b \rho + \kappa n \rho}.  \]

\subsubsection*{Nondimensionalization}
In order the analyze the model more easily, we nondimensionalize the system with the following change of variable
\[ \rho = R \tilde{\rho}, \qquad n = N \tilde{n}, \qquad t = T \tau. \]
There are many possible choice to choose the scaling factors $R, N, T$. However, we will choose them in a way that they are always positive, and the system of ODE becomes as simple as possible. Substituting the change of variable above in the ODE system, we will get
\begin{align*}
	&\frac{d\tilde{\rho}}{d\tau} = \frac{vNT}{R}\tilde{n} + \delta_v T \tilde{\rho},\\
	&\frac{d\tilde{n}}{d\tau} = T(\lambda_s-\delta_t)\tilde{n} + \frac{\lambda_b T R}{N} \tilde{\rho} + T\kappa R \tilde{n}\tilde{\rho}.
\end{align*}
We choose the following values for $T,N$, and $R$
\[ T = \frac{1}{\delta_v}, \qquad R = \frac{\delta_v}{\kappa}, \qquad N = \frac{\lambda_b}{\kappa}. \]
This is a very suitable moment to pause and check the dimensions if they match (I did it and all of them matches!). With these choices from the coefficients, the system of ODEs will be
\[ \frac{d\tilde{n}}{d \tau} = \frac{\lambda_s - \delta_t}{\delta_v} \tilde{n} + \tilde{\rho} + \tilde{n}\tilde{\rho}, \qquad
\frac{d\tilde{\rho}}{d\tau} = \frac{v\lambda_b}{\delta_v^2}\tilde{n} + \tilde{\rho}. \]
To make the ODEs simpler to work with, we will write $n,\rho$ in place of $\tilde{n}$ and $\tilde{\rho}$, and also we introduce the following parameters
\[ \alpha = \lambda_s - \delta_t, \qquad \beta =\delta_v, \qquad \gamma = v\lambda_b. \]
Then we can write

\begin{equation*}
	\boxed{
		\begin{aligned}
			&\dot{n} = \frac{\alpha}{\beta}n + \rho + n\rho, \\
			&\dot{\rho} = \frac{\gamma}{\beta^2}n +\rho.
		\end{aligned}
	}
	\tag{\smiley}
\end{equation*}

\newpage

\section{Some Ideas to Try}
This section might have very simple, basic and sometimes silly ideas that came into my mind during developing some models and I thought they might worth trying
\begin{itemize}
	\item Developing a model for a weighted graph generation. I suspect a weighted graph might have all the necessary information we want.
\end{itemize}

### 2024-01-28 Findings (Shared Folder)/Findings.tex

@@
 \usepackage{commands}
-\usepackage[numbers]{natbib}
+\usepackage{natbib} 
+%\setcitestyle{authoryear, open={([},close={)]}}
+%\setcitestyle{numbers,square}
+\bibliographystyle{dinat}
 
@@
 
-	\include{Chapters/StatisticalPhysicsForBiology.tex}
+%	\include{Chapters/StatisticalPhysicsForBiology.tex}
+	\include{Chapters/modelAttempts.tex}
 	\include{Chapters/PapersReviewed.tex}
-
-\section{Leah Comments Jan 24, 2024}
-Please change the citation style to Author et al (year) in place of [number], so it is easier to see who you are citing without having to flip to the bibliography.
-Thanks for linking the bibliography to the URLs of the papers so it's possible to scan them.
-
-When citing papers, it is best if you can also say 1-2 sentences about those papers, even based on their abstract (in your own words, of course, never copied directly). For example:
-
-Byrne (2010) reviews theoretical cancer models and demonstrates the advantages of colaboration between modelers and experimentalists.
-
-\subsection{Suggested research style and flavour}
-
-Since it is unlikely that we will get data for the detailed mechanochemical mechanisms for blood vessel growth in the prostate tumors, it makes sense to (a) start simple from very minimal models that can be linked to data and (b) avoid introducing variable that we have no hope of measuing in the obtained data.
-My understanding is that (for now) we will have to make do with at best some bulk properties of the blood vessels, so models with a lot of detail will hardly help us.
-
-Here is a possible minimalistic stepwise approach, where we start very simple and gradually build up more detail, starting with simple assumptions.
-
-\textbf{Definitions:}
-
-$n(t)$= density of tip cells in area of interest, (number per unit area)
-
-$\rho(t)$ = density of blood vessels (length per unit area),
-
-$c(t)$ = concentration of drug delivered to region by blood vessels
-
-\subsection{Step 1: Bulk model}
-
-Ignoring spatial structure, we only track the density of vessels.
-Assume everything is spatially uniform, so there will be no spatial derivatives to consider.
-We construct an ODE model, and make an elementary assumption.
-
-\textbf{Assumption 1:} The drug is delivered by diffusion from the capillaries into the tissue.
-Hence, as a rough approximation, and (for now) neglecting the detailed structure of the vessel network, the amount of drug delivered to the region per unit time is proportional to the density of the blood vessels.
-
-\textbf{Step 1a: Elementary model:}
-
-Assume that tips extend at some rate $v$ (units of length/time), creating additional length of capillaries as they extend.
-Assume capillaries may also have some loss rate $\delta$ (per unit time).
-Write down an ODE for the rate at which capillary density changes with time.
-
-\begin{subequations}
-\label{eq:SimplestModel}
-\begin{equation}
-\label{eq:vesseldensODE}
- \frac{d \rho}{dt}= ??   
-\end{equation}
-
-Assume that new tip cells are created by branching along sides of vessels (or possibly by splitting of existing tip cells) at a rate $\beta$ per unit length per unit time, and that tips disappear when they reconnect to a capillary at some rate $\kappa$ to form a loop.
-[Note: reconnection requires the interaction of tips with capillaries, and would thus be handled as mass action term.
-What are the units of $\kappa$?]
-Write down an equation for the rate of change of tip density.
-\begin{equation}
-\label{eq:tipdensODE}
- \frac{d n}{dt}= ??   
-\end{equation}
-
-Complete the ODE model equations \eqref{eq:SimplestModel}.
-Analyse the model so far by determining the steady state densities $\rho_{ss} , n_{ss}$, and how they depend on the parameters $v,\delta,\beta, \kappa$.
-Determine stability of SS.
-Create a phase plane diagram that shows the expected dynamics.
-Simulate the simple ODE system assuming some values of the parameters.
-
-We made the assumption that drug delivery is roughly proportional to the vessel density. Write down an approximate ODE for concentration of drug in the region. 
-\begin{equation}
-\label{eq:drugODE}
- \frac{d c}{dt}= ??   
-\end{equation}	
-\end{subequations}Explain how this level of drug depends on the vessel branching and growth parameters.
-
-So far, the blood vessels affect the drug but not the other way around.
-
-\bigskip
-
-\textbf{Step 1b: Coupling vessel dynamics to drug}
-
-Consider how the level of drug might affect the vessel parameters (branching or growth rate or death rate, etc). This will introduce feedback from the drug to the vessel density.
-
-Write down one or two variants of such a model and analyse them fully (including steady states, simulations, and some interpretation of what it means for overall treatment of the tissue.)
-
-Note that the drug dynamics would be fast on the timescale of vessel growth, so there is some time-scale separation that you can take advantage of.
-
-\subsection{(Optional) Step 2: Simple spatially distributed 1D system}
-We continue with simplest model but now take spatial growth of vessels into a region. So we consider $\rho(x,t), n(x,t), c(x,t)$ as variables of interest.
-We make the same assumptions as above, but now we take into account the fact that there is a flux of tips growing into a region,
-\[
-J=nv.
-\]
-Explain why this is a flux.
-The equations will be modified to form PDEs.
-Use the 1D balance equation to create that equation for $n$.
-Explain whether you need to add any spatial derivatives to the equation for $\rho$.
-The drug diffusion in the spatial variable will also introduce spatial derivatives in the equation for $c$.
-Write down the modified 1D spatial model.
-Note that we do not assume anything like chemotaxis or other fancy mechanisms for the tip motion at this point.
-
-\begin{subequations}
-\label{eq:SpatialtModel}
-\begin{align}
-\frac{\partial \rho}{\partial t }&= ?\\
-\frac{\partial n}{\partial t }&= ?\\
-\frac{\partial c}{\partial t }&= ?
-\end{align}
-\end{subequations}
-
-Remark: see above for timescale separation.
-
-\textbf{Step 2a: Analysis of wave of invasion}
-
-Consider looking for traveling wave solutions of the $\rho,n$ system on its own to ask how blood vessels spread along a 1D direction and invade a tissue.
-(Write down ODEs by transforming variables to $z=x-ct$ where $c$ is wave speed, then analyze existence of traveling was in the $\rho n$ phase plane.
-See one of my books or ask Jack Hughes for help if you are not yet familiar with this idea.)
-
-\textbf{Step 2b: Simulations}
-For simulations of the whole system: You will need to assume some boundary conditions on $n$ and on $c$, as well as some initial distribution in order to simulate this system.
-
-\subsection{Step 3: An agent-based (CPM) model:}
-Look up the simplest work on Merks and Rens and co and find their CPM model. Ask whether a Morpheus xml file already exists for this model (can ask the Morpheus team or Merks). If not, create one.
-
-Set up this model and adapt it to describing a simple branching vessel structure, similar to what we have above.
-
-ADD: assume that the cells in this network ``secrete'' drug that then diffuses into the tissue and has some decay time. Find ways of plotting properties of the vessels and the drug concentration.
-
-Here you can get creative, and assume that the tip cell growth etc are affected by drug level, etc. (Again, time scale separation is important.)
-
-Your role will be to extend the Merks model to include this drug aspect.
-
-NOTE: some of Merk's work includes the dynamics of an ECM.
-I would suggest to avoid extending the model with such a dynamic variable, and to assume instead, that it is a static field or vector-field that affects the rate or direction of tip cell motion.
-
-\subsection{Step 4: Look for data}
-
-This can be done in parallel with other steps: look for specific data on blood vessel density in normal and cancerous tissue.
-There may be animal studies in which the vessel density is tracked over time.
-
-Find if there is data that we can use to help constrain any of these simple models.
-
-For sure it's easier to find bulk vessel density than to find its spatial distribution and the chemical factors liek VEGF that are modeled in some papers.
-
-\subsection{Step 5: More details and other variants}
-
-You can later (after all the early steps) extend and improve the model in various ways. Some suggestions include the following:
-
-\begin{itemize}
-    \item Write down an equation for the number of loops that accumulate as tips reconnect to blood vessels (extend simple model).
-    \item Find a way to associate these with ``tortuosity'' of vessel network that could affect its conductivity of drug to tissue.
-    \item Consider some kind of D'Arcy's Law (porous medium) as a measure of how vessel structure can reduce net drug delivery.
-    \item Vessels have various radii and sizes.
-You may want to consider how this affects the model as well as the implications on drug delivery.
-A PDE model with a distribution of vessel diameters would likely be a bit newer than the above simple branching equations.
-\end{itemize}
-
-	%\backmatter
-	% Appendices, bibliography, etc.
+	\include{Chapters/MolecularBiology.tex}
+	\include{Chapters/Comments.tex}
+	\include{Chapters/MeetingLog.tex}
 
 	% Bibliography
-	\bibliographystyle{unsrtnat} % The bibliography style
+%	\bibliographystyle{plainnat}
+
 	\bibliography{References/references} % The filename of the .bib file (without the extension)

### 2024-01-28 Findings (Shared Folder)/Chapters/modelAttempts.tex
new comments added

@@
 To make the ODEs simpler to work with, we will write $n,\rho$ in place of $\tilde{n}$ and $\tilde{\rho}$, and also we introduce the following parameters
-\[ \alpha = \lambda_s - \delta_t, \qquad \beta =\delta_v, \qquad \gamma = v\lambda_b. \]
+\[ \alpha = \lambda_s - \delta_t, \qquad \beta =\delta_v, \qquad \gamma = v\lambda_b. \tag{$\clubsuit$} \]
 Then we can write
-
 \begin{equation*}
@@
 \end{equation*}
+After some algebra, it turns out that there are two equilibrium points for this system.
+\[ p^0_1 = (0,0), \qquad p^0_2 = (\frac{\alpha\beta}{\gamma} - 1, \frac{\gamma}{\beta^2}-\frac{\alpha}{\beta}) = (\frac{\alpha\beta-\gamma}{\gamma},\frac{\gamma - \alpha\beta}{\beta^2}). \]
+\begin{strangeObs}
+	Something does not make sense!
+We know that the variables $n$ and $\rho$ must all be positive, as they are both representing biological values.
+However, by looking closely at the second component of $p^0_2$, it turns out th at $\gamma - \alpha\beta$ should be positive.
+In order to have the first component positive as well, then $\gamma$ should be negative, which is not possible.
+That is because by $(\clubsuit)$ we know that $\gamma>0$.
+There is at least one solution for this paradox, which is letting $\alpha\beta - \gamma =0$.
+Then this equilibrium point will be the same as $p^0_1$, showing the fact that the system has only one equilibrium point with biological meaning, which is the origin (not very interesting)!
+\\
+	However, I am going to ignore this observation for now and proceed with the stability analysis.
+
+	\label{strangeObs:SimpleODEequiliNegative}
+\end{strangeObs}
+
+In order to analyze the stability of these equilibrium points, we first need to calculate the Jacobian matrix of the ODE system 
+\[ DF = \matt{\alpha/\beta+\rho}{1+n}{\gamma/\beta^2}{1}. \]
+
+\subsubsection{Stability Analysis of $p^0_2$}
+By evaluating the Jacobian matrix at the equilibrium point we will have
+\[ DF[p^0_2] = \matt{\gamma/\beta^2}{\alpha\beta/\gamma}{\gamma/\beta^2}{1}. \]
+The trace and determinant of this matrix is
+\[  \Delta = \gamma/\beta_2 - \alpha/\beta, \qquad \sigma=\gamma/\beta^2 + 1.  \]
+By close inspection, it turns out that $\Delta$ is the same as the first component of $p^0_2$, which should be positive.
+This implies $\Delta > 0$.
+So the sign of the trace of the Jacobian matrix will determine the stability.
+From $(\clubsuit)$, $\gamma>0$.
+Thus $\sigma > 0$.
+This indicates that the equilibrium point $p_0^2$ is unstable (look at \autoref{strangeObs:SimpleODEequiliNegative} for another ongoing issue regarding this equilibrium point).
+
+\subsubsection{Stability Analysis of $p^0_1$}
+The Jacobian matrix evaluated at $p^0_1$ is
+\[ DF[p^0_1] = \matt{\alpha/\beta}{1}{\gamma/\beta^2}{1}. \]
+The determinant and the trace will be
+\[  \Delta = \alpha/\beta - \gamma/\beta^2, \qquad \sigma = \alpha/\beta+1. \]
+Since the determinant is minus the first component of $p^0_2$. Thus $\delta<0$. This implies that regardless of the value of $\sigma$, the equilibrium $p^0_1$ is a saddle point.
+
+\begin{strangeObs}
+	This equilibrium point $p^0_1$ has no choice but to be stable, if we expect the model to represent real biological situation.
+But our analysis above, reveals that this equilibrium point is in fact a saddle point.
+However, I can argue that if that the eigenvector with negative eigenvalue for the Jacobian matrix evaluated at $p^0_2$, lies in the first quadrant, then $p^0_2$ will ``look'' stable for the points at the first quadrant.
+But I am not sure if this is something normal to happen for a biological model.
+\end{strangeObs}
 

### 2024-01-29 Findings (Shared Folder)/Chapters/modelAttempts.tex
new details added

--- 2024-01-28 Findings (Shared Folder)/Chapters/modelAttempts.tex
@@
 
-	$c(t)$ = concentration of drug delivered to region by blood vessels.
+	$c(t)$ = concentration of drug delivered to region by blood vessels (nano mole per unit area).
 \end{quote}
-An updating list of model parameters
+An updating list of model parameters:
 \begin{itemize}
@@
 	\item $\kappa$ [area/length/time]: Re-connection of tip cells to the other capillaries to form loops.  
+	\item $\mu$ [1/time]: The rate at which 
+	\item $\sigma$ [area/length]: The coverage of the blood vessels in the region.
+\end{itemize}
+And a list input functions
+\begin{itemize}
+	\item $f(t)$: [nmol/length]: The amount of drug inside the capillary.
 \end{itemize}
 
-\subsubsection*{Studying dynamics of vessel formation}
+\subsection*{Studying dynamics of vessel formation}
 \[ \frac{d\rho}{dt} = ??. \]
@@
 Also, assuming the vascular structure degrades with rate $\delta$ [per unit time], we can add the degradation term 
-\[ \boxed{\frac{d\rho}{dt} = vn + \delta_v \rho} . \]
+\[ \boxed{\frac{d\rho}{dt} = vn - \delta_v \rho} . \]
 
@@
 
-\[ \boxed{\frac{dn}{dt} = (\lambda_s - \delta_t) n + \lambda_b \rho + \kappa n \rho}.  \]
+\[ \boxed{\frac{dn}{dt} = (\lambda_s - \delta_t) n + \lambda_b \rho - \kappa n \rho}.  \]
 
@@
 \begin{align*}
-	&\frac{d\tilde{\rho}}{d\tau} = \frac{vNT}{R}\tilde{n} + \delta_v T \tilde{\rho},\\
-	&\frac{d\tilde{n}}{d\tau} = T(\lambda_s-\delta_t)\tilde{n} + \frac{\lambda_b T R}{N} \tilde{\rho} + T\kappa R \tilde{n}\tilde{\rho}.
+	&\frac{d\tilde{\rho}}{d\tau} = \frac{vNT}{R}\tilde{n} - \delta_v T \tilde{\rho},\\
+	&\frac{d\tilde{n}}{d\tau} = T(\lambda_s-\delta_t)\tilde{n} + \frac{\lambda_b T R}{N} \tilde{\rho} - T\kappa R \tilde{n}\tilde{\rho}. \tag{\twonotes}
 \end{align*}
@@
 This is a very suitable moment to pause and check the dimensions if they match (I did it and all of them matches!). With these choices from the coefficients, the system of ODEs will be
-\[ \frac{d\tilde{n}}{d \tau} = \frac{\lambda_s - \delta_t}{\delta_v} \tilde{n} + \tilde{\rho} + \tilde{n}\tilde{\rho}, \qquad
-\frac{d\tilde{\rho}}{d\tau} = \frac{v\lambda_b}{\delta_v^2}\tilde{n} + \tilde{\rho}. \]
+\[ \frac{d\tilde{n}}{d \tau} = \frac{\lambda_s - \delta_t}{\delta_v} \tilde{n} + \tilde{\rho} - \tilde{n}\tilde{\rho}, \qquad
+\frac{d\tilde{\rho}}{d\tau} = \frac{v\lambda_b}{\delta_v^2}\tilde{n} - \tilde{\rho}. \]
 To make the ODEs simpler to work with, we will write $n,\rho$ in place of $\tilde{n}$ and $\tilde{\rho}$, and also we introduce the following parameters
-\[ \alpha = \lambda_s - \delta_t, \qquad \beta =\delta_v, \qquad \gamma = v\lambda_b. \]
+\[ \alpha = \lambda_s - \delta_t, \qquad \beta =\delta_v, \qquad \gamma = v\lambda_b. \tag{$\clubsuit$} \]
 Then we can write
-
 \begin{equation*}
@@
 		\begin{aligned}
-			&\dot{n} = \frac{\alpha}{\beta}n + \rho + n\rho, \\
-			&\dot{\rho} = \frac{\gamma}{\beta^2}n +\rho.
+			&\dot{n} = \frac{\alpha}{\beta}n + \rho - n\rho, \\
+			&\dot{\rho} = \frac{\gamma}{\beta^2}n -\rho.
 		\end{aligned}
@@
 \end{equation*}
+After some algebra, it turns out that there are two equilibrium points for this system.
+\[ p^0_1 = (0,0), \qquad p^0_2 = (\frac{\alpha\beta}{\gamma} + 1, \frac{\gamma}{\beta^2}+\frac{\alpha}{\beta}) = (\frac{\alpha\beta+\gamma}{\gamma},\frac{\gamma + \alpha\beta}{\beta^2}). \]
+%\begin{strangeObs}
+%	Something does not make sense!
+We know that the variables $n$ and $\rho$ must all be positive, as they are both representing biological values.
+However, by looking closely at the second component of $p^0_2$, it turns out th at $\gamma - \alpha\beta$ should be positive.
+In order to have the first component positive as well, then $\gamma$ should be negative, which is not possible.
+That is because by $(\clubsuit)$ we know that $\gamma>0$.
+There is at least one solution for this paradox, which is letting $\alpha\beta - \gamma =0$.
+Then this equilibrium point will be the same as $p^0_1$, showing the fact that the system has only one equilibrium point with biological meaning, which is the origin (not very interesting)!\\
+%	However, I am going to ignore this observation for now and proceed with the stability analysis.
+%	
+%	\label{strangeObs:SimpleODEequiliNegative}
+%\end{strangeObs}
+
+In order to analyze the stability of these equilibrium points, we first need to calculate the Jacobian matrix of the ODE system 
+\[ DF = \matt{\alpha/\beta-\rho}{1-n}{\gamma/\beta^2}{-1}. \]
+
+\subsubsection{Stability Analysis of $p^0_2$}
+By evaluating the Jacobian matrix at the equilibrium point we will have
+\[ DF[p^0_2] = \matt{-\gamma/\beta^2}{-\alpha\beta/\gamma}{\gamma/\beta^2}{-1}. \]
+The trace and determinant of this matrix is
+\[  \Delta = \gamma/\beta^2 + \alpha/\beta, \qquad \sigma=-\gamma/\beta^2 - 1.  \]
+By close inspection, it turns out that $\Delta$ is the same as the first component of $p^0_2$, which should be positive.
+This implies $\Delta > 0$.
+So the sign of the trace of the Jacobian matrix will determine the stability.
+From $(\clubsuit)$, $\gamma>0$.
+Thus $\sigma < 0$.
+This indicates that the equilibrium point $p_0^2$ is stable equilibrium.
+Also, note that since $\sigma$ can never transversally become positive from being negative (i.e.
+passing through $\sigma=0$, transversally), thus we can rule out the existence of any Hopf bifurcation with this particular model.
+
+\begin{observation}[stability of $p^0_2$]
+	The Jacobian matrix evaluated at $p^0_2$ has
+	\[ \Delta \geq 0, \qquad \sigma < 0. \]
+	Thus equilibrium point $p^0_2$ is a hyperbolic sink (when $\delta > 0$). When $\Delta>0$, then we have two imaginary eigenvalues
+	\[ \lambda_1 = \sigma/2+i\sqrt{\Delta}, \qquad \lambda_2 = \sigma/2-i\sqrt{\Delta}. \]
+	where $\sigma$ is always negative. This implies that $p^0_2$ will always be a stable focus and the equilibrium point will be approached in a damping oscillatory way.\\
+	In the special case where $\Delta =0$, we will have $p^0_2=p^0_1$, thus it will inherit the stability of $p_0^1$. 
+\end{observation}
+
+\begin{observation}[No sustained oscillations]
+	Note that since $\sigma<0$ for all values of the parameters of the model, then there is no chance to observe a Hopf bifurcation, thus ruling out any sustained oscillations in the model.
+\end{observation}
+
+\subsubsection{Finding Lyapunov Function}
+In attempting to find a Lyapunov function, I thought it might be a good idea to have a different choices for the non-dimensionalization scaling so that I can have control on the nonlinear part $n\rho$.
+But it seems that there are no possible ways to achieve this.
+That is because in $(\twonotes)$, we can not make the first term of RHS of $\dot{\tilde{\rho}}$ and the second term of RHS of $\dot{\tilde{n}}$ simultaneously to be 1.
+Thus there are no choices for the scaling factors to make the coefficient of $\tilde{n}\tilde{\rho}$ to be 1.
+
+\subsubsection{Stability Analysis of $p^0_1$}
+The Jacobian matrix evaluated at $p^0_1$ is
+\[ DF[p^0_1] = \matt{\alpha/\beta}{1}{\gamma/\beta^2}{-1}. \]
+The determinant and the trace will be
+\[  \Delta = -\alpha/\beta - \gamma/\beta^2, \qquad \sigma = \alpha/\beta-1. \]
+\begin{observation}
+	The determinant $\Delta$ is negative the first term of the $p^0_2$.
+Thus $\Delta\leq0$.
+When $\Delta<0$, then regardless of value of $\sigma$, $p^0_1$ is a hyperbolic saddle.
+However, when $\Delta=0$, then $p^0_1$ is non-hyperbolic and further analysis is required to determine the stability.
+\end{observation}
+The following shows the phase portrait of the system with some values for the parameters.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.5\linewidth]{images/simpleODEModel1PhasePortrait}
+
+	\label{fig:simpleodemodel1phaseportrait}
+\end{figure}
+
+%\begin{strangeObs}
+%	This equilibrium point $p^0_1$ has no choice but to be stable, if we expect the model to represent real biological situation.
+But our analysis above, reveals that this equilibrium point is in fact a saddle point.
+However, I can argue that if that the eigenvector with negative eigenvalue for the Jacobian matrix evaluated at $p^0_2$, lies in the first quadrant, then $p^0_2$ will ``look'' stable for the points at the first quadrant.
+But I am not sure if this is something normal to happen for a biological model.
+%\end{strangeObs}
+
+\subsection*{Adding the drug-vessel interaction}
+In order to bring the drug-vessel interaction into play, we develop a ODE for $c(t)$ add approporite terms to the RHS of $\dot{n}$ and $\dot{\rho}$.
+\[ \frac{dc}{dt} = \mu (\rho(t)f(t) - \sigma c(t)\rho(t)) - \boxed{\lambda c} \]
+where $\mu$ has the unit [1/time], and $f(t)$ is the amount drug inside the capillary that has the unit [nmol per unit length].
+Note that we have assumed the exchange of drug between the vessels and the region of interest is proportional to the difference in the concentration of drug in two different environments.
+Furthermore, the coefficient $\sigma$ has the units of [area/length] which is an indicator of the area coverage of the blood vessels.
+This parameter somehow characterizes the space filling and fractal structure of the blood vessels.
+This parameter should have some relations with the fractal dimension of a given vascular structure.
+Considering the dynamics of this parameter can possibly reflect some of the topological and non-local characterizations of the vascular network.
 

### 2024-01-29 Findings (Shared Folder)/Chapters/modelAttempts.tex
new details updated

@@
 	\item $\kappa$ [area/length/time]: Re-connection of tip cells to the other capillaries to form loops.  
-	\item $\mu$ [1/time]: The rate at which 
+	\item $\mu$ [1/time]: Permeability of the capillary to the drug. 
 	\item $\sigma$ [area/length]: The coverage of the blood vessels in the region.
@@
 
-\subsection*{Adding the drug-vessel interaction}
+\subsection*{Drug delivery}
 In order to bring the drug-vessel interaction into play, we develop a ODE for $c(t)$ add approporite terms to the RHS of $\dot{n}$ and $\dot{\rho}$.
@@
 
+\subsection*{Adding the drug-vessel interaction}
+Presence of drug in the environment can have many different effects.
+It can kill/deactivate the existing tip cells (increasing $\delta_t$).
+Or it can change the rate at which the endothelial cells turn into the tip cells (changing the value of $\delta_b$).
+Or it can affect the tip cell division rate ($\lambda_s$).
+It can also affect the cellular migration of the tip cells and change the value of parameter $v$.
+I am not sure which of these effects will be more significant than the others.
+
 \newpage

### 2024-02-05 Findings (Shared Folder)/Chapters/modelAttempts.tex
new updates added

@@
 	\[ \Delta \geq 0, \qquad \sigma < 0. \]
-	Thus equilibrium point $p^0_2$ is a hyperbolic sink (when $\delta > 0$). When $\Delta>0$, then we have two imaginary eigenvalues
-	\[ \lambda_1 = \sigma/2+i\sqrt{\Delta}, \qquad \lambda_2 = \sigma/2-i\sqrt{\Delta}. \]
-	where $\sigma$ is always negative. This implies that $p^0_2$ will always be a stable focus and the equilibrium point will be approached in a damping oscillatory way.\\
-	In the special case where $\Delta =0$, we will have $p^0_2=p^0_1$, thus it will inherit the stability of $p_0^1$. 
+	Thus equilibrium point $p^0_2$ is a hyperbolic sink (when $\delta > 0$).
+This hyperbolic sink can be of the type stable node (with purely real eigenvalues) or stable focus (with complex valued eigenvalues whose real part is negative).
+But since these two kind of stability are topologically equivalent, we don't do further analysis to distinguish them at this point.
+%	When $\Delta>0$, then we have two imaginary eigenvalues
+%	\[ \lambda_1 = \sigma/2+i\sqrt{\Delta}, \qquad \lambda_2 = \sigma/2-i\sqrt{\Delta}. \]
+%	where $\sigma$ is always negative. This implies that $p^0_2$ will always be a stable focus and the equilibrium point will be approached in a damping oscillatory way.\\
+%	In the special case where $\Delta =0$, we will have $p^0_2=p^0_1$, thus it will inherit the stability of $p_0^1$. 
 \end{observation}
@@
 
+Furthermore, the following diagram shows the nullclines and the sign of the vector field at the different regions of the phase portrait.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.5\linewidth]{images/nullClines.png}
+
+	\label{fig:simpleodemodel1phaseportrait}
+\end{figure}
+
 %\begin{strangeObs}
@@
 
+\subsubsection*{Basic and Simplified delivery scenario}
+We assume that there is an infinite pool of drug (i.e. the patient is getting injected continuously) so the amount of drug per unit length in the capillary is constant $C_0$ [nmol per unit length].
+\[ \frac{dc}{dt} = \mu \rho(t)(C_0 - \sigma c(t)).  \]
+
+Also, note that we have ignored the radiation decay term to keep stuff simpler at this stage.
+The reason behind this choice is that at this stage, in adding the drug-vessel interaction, we will only consider a mass-action type interaction (i.e.
+chemical interactions), and no radio-biological interaction will be assumed (which has more complexity).
+
+Another assumption that we make is that since the molecules of the drug are much smaller and simpler than the cells in the body, then we assume that they reach to equilibrium much more faster than the characteristic time scale of tip cell movement, and the death/generation of the cells.
+This basically means that we can simply assume $dc/dt=0$ to arrive at the following algebraic equation for $c$.
+\[ C_0 = \sigma c(t). \]
+
 \subsection*{Adding the drug-vessel interaction}
@@
 It can also affect the cellular migration of the tip cells and change the value of parameter $v$.
-I am not sure which of these effects will be more significant than the others.
+The following lists the parameters corresponding to the amplitude of each of these interactions
+\begin{itemize}
+	\item $a_1$: changing the tip cell movement/migration
+	\item $a_2$: killing/deactivating the tip cells
+	\item $a_3$: changing the endothelial-to-tip cell conversion rate
+	\item $a_4$: changing the tip cell division rate
+\end{itemize}
+
+\subsection*{Some discussions on drug interaction}
+In mathematical modeling of biological processes, incorporating drug interactions can be complex, as the drug can influence various parameters of the system in different ways.
+The choice of interaction term often depends on the nature of the drug action and the available experimental data.
+Here are a few common approaches to include drug interactions in the model:
+
+\begin{enumerate}
+	\item \textbf{Linear Interaction}: If the drug effect is proportional to its concentration, we can model the interaction linearly. For instance:
+	\begin{align*}
+		v(c) &= v_0 - a_1 c, \\
+		\delta_v(c) &= \delta_{v0} + a_2 c,\\
+		\lambda_b(c) &= \lambda_{b0} + a_3 c,\\
+		\lambda_s(c) &= \lambda_{s0} - a_4 c.
+	\end{align*}
+	Here, \( v_0 \) and \( \delta_{v0} \) are the baseline motility and degradation rates without the drug, while \( a_1 \) and \( a_2 \) represent the sensitivity of these rates to the drug concentration.
+Similarly \( \lambda_{b0} \) and \( \lambda_{s0} \) are the baseline rates without the drug, and \( a_3 \), \( a_4 \) are the sensitivities of these rates to the drug concentration.
+
+	\item \textbf{Hill Function}: If the drug effect exhibits saturation -- meaning it has a maximum effect regardless of concentration -- a Hill function can be appropriate:
+	\begin{align*}
+		v(c) &= v_0 \left(1 - \frac{a_1 c^h}{K_d^h + c^h}\right), \\
+		\delta_v(c) &= \delta_{v0} \left(1 + \frac{a_2 c^h}{K_d^h + c^h}\right),\\
+		\lambda_b(c) &= \lambda_{b0} \left(1 + \frac{a_3 c^h}{K_{d3}^h + c^h}\right),\\
+		\lambda_s(c)& = \lambda_{s0} \left(1 - \frac{a_4 c^h}{K_{d4}^h + c^h}\right),
+	\end{align*}
+	Here, \( h \) is the Hill coefficient that determines the steepness of the response curve, and \( K_d \) is the drug concentration at which the effect is half of its maximum.
+With \( h \) as the Hill coefficient, and \( K_{d3} \), \( K_{d4} \) as the half-maximal effective concentrations for \( \lambda_b \) and \( \lambda_s \), respectively.
+
+	\item \textbf{Michaelis-Menten Kinetics}: If the drug interaction is enzyme-like, you can model it using Michaelis-Menten kinetics:
+	\begin{align*}
+		v(c) &= v_0 \left(1 - \frac{a_1 c}{K_m + c}\right), \\
+		\delta_v(c) &= \delta_{v0} \left(1 + \frac{a_2 c}{K_m + c}\right),\\
+		\lambda_b(c) &= \lambda_{b0} \left(1 + \frac{a_3 c}{K_m + c}\right),\\	
+		\lambda_s(c) &= \lambda_{s0} \left(1 - \frac{a_4 c}{K_m + c}\right),
+	\end{align*}
+	Where \( K_m \) is the Michaelis constant, representing the drug concentration at which the rate of reaction is half of its maximum.
+where \( K_m \) is the Michaelis constant, indicative of the concentration at which the reaction rate is half its maximum.
+
+	\item \textbf{Exponential or Sigmoidal Functions}: For more complex drug effects, such as those that have a threshold effect or exhibit a sigmoidal dose-response, exponential or sigmoidal functions can be used.
+\end{enumerate}
+
+These interaction terms would be incorporated into the model by modifying the differential equations as follows:
+\[
+\frac{d\rho}{dt} = v(c)n - \delta_v(c) \rho,
+\]
+
+\[
+\frac{dn}{dt} = (\lambda_s(c) - \delta_t) n + \lambda_b(c) \rho - \kappa n \rho,
+\]
+
+where \( \lambda_b(c) \) and \( \lambda_s(c) \) are now functions of the drug concentration that reflect the modulation of the endothelial-to-tip cell conversion rate and the tip cell division rate by the drug.
+
+The form of the interaction should be chosen based on the biological mechanism of the drug action, the type and quality of experimental data available, and the ability to estimate the additional parameters introduced by these functional forms with the available data.
+To decide which model to use, consider the following:
+
+\begin{itemize}
+	\item \textbf{Biological Mechanism}: Does the drug interact with its target in a manner that is competitive, non-competitive, or does it follow some form of cooperative binding?
+This will guide whether you use linear, Hill, or Michaelis-Menten kinetics.
+	\item \textbf{Data Availability}: What kind of data do you have?
+If you have dose-response data, you can fit these models to the data to estimate parameters like \( a_1 \), \( a_2 \), \( h \), \( K_d \), or \( K_m \).
+	\item \textbf{Parameter Estimation}: Can you estimate the additional parameters introduced by these functions? More complex models require more data for accurate parameter estimation.
+\end{itemize}
 

### 2024-02-09 Findings (Shared Folder)/Chapters/MeetingLog.tex
new updates

@@
 \chapter{Meeting log}
+\section{Meetings with Leah}
+\subsection{29 Jan Meeting}
 
-\subsection{29 Jan Meeting with Leah}
-Things to discuss
-\begin{itemize}
-	\item Citation with Year-Author style, or Number Style?
+\begin{itemize}[itemsep=0pt, parsep=0pt]
+	\item Fixing some errors in the eigenvalues for the main differential equations. 
+	\item Add the nullclines plot.
+	\item Add the possible interactions between vascular networks and the drug
 \end{itemize}
+
+\subsection{5 Feb Meeting}
+\begin{itemize}[itemsep=0pt, parsep=0pt]
+	\item Thinking again about including the decay of radiopharmaceuticals: This has both pros and cons. The pros is that 
+	\begin{itemize}[itemsep=0pt, parsep=0pt]
+		\item more realistic model,
+		\item it is always good to have same sort of decay in the model to ensure the stability.
+	\end{itemize}
+	However, the downside is that radiation interacts more wildly with the the cells present in the mode.
+It can inhibit them (by killing them) which is in a non mass action or Hill function style.
+The killing mechanism follows the linear quadratic rule.
+Also, the radiation can have activation functionality on the same cells by simply causing crazy genetic mutations.
+\textbf{So decision on including the radiation term in the basic model should be done with extra care.}
+	\item Doing qualitative analysis (not quantitative) with the nullclines and the change in parameters due to the drug interaction.
+This way we can capture the possible interactions with out considering the actual functional form of the interaction.
+	\item After doing the qualitative analysis, I need to do some literature review to see what are the possible drug-vessel interactions (both radiation and chemical)
+	\item Consider adding the tumor compartment. The tumor compartment can interact with the vascular network by
+	\begin{itemize}[itemsep=0pt, parsep=0pt]
+		\item increasing the mobility of the tip cells: both by increasing chemotaxis agents, and also by loosens the extracellular matrix,
+		\item Any other interaction that needs to be determined carefully.
+	\end{itemize}
+	\item Adding the condition under which the stability of $p^0_2$ is focus or node.
+\end{itemize}
+
+\section{Meetings with Arman}

### 2024-02-09 Findings (Shared Folder)/Chapters/modelAttempts.tex
new updates

@@
 \end{equation*}
+In order to find the equilibrium points, we demand $\dot{n} = 0$ as well as $\dot{\rho} = 0$. This will lead to the following equations
+\begin{align*}
+	\dot{n} = 0&: \qquad \frac{\alpha}{\beta} n + \rho - n\rho = 0, \\
+	\dot{\rho} =0&: \qquad \frac{\gamma}{\beta^2}n - \rho = 0.
+\end{align*}
 After some algebra, it turns out that there are two equilibrium points for this system.
-\[ p^0_1 = (0,0), \qquad p^0_2 = (\frac{\alpha\beta}{\gamma} + 1, \frac{\gamma}{\beta^2}+\frac{\alpha}{\beta}) = (\frac{\alpha\beta+\gamma}{\gamma},\frac{\gamma + \alpha\beta}{\beta^2}). \]
+\[ p^0_1 = (0,0), \qquad p^0_2 = (\frac{\alpha\beta}{\gamma} + 1, \frac{\gamma}{\beta^2}+\frac{\alpha}{\beta}) = (\frac{\alpha\beta+\gamma}{\gamma},\frac{\gamma + \alpha\beta}{\beta^2}).
+\tag{E.1.1} \]
 %\begin{strangeObs}
@@
 
+\FloatBarrier
+
 %\begin{strangeObs}
@@
 %\end{strangeObs}
+
+\subsection*{Quantitative analysis with nullclines}
+Drawing the phase portrait and including the nullclines helps in understanding the quantitative effect of change in parameters (which causes by the drug-vessel interaction).
+To draw the nullclines, we require
+\[ \vectt{\dot{n}}{\dot{\rho}} = \vectt{f_1(n,\rho)}{f_2(n,\rho)} = \vectt{\frac{\alpha}{\beta}n + \rho - n\rho}{\frac{\gamma}{\beta^2}n - \rho} = \vectt{0}{0}. \]
+After a little bit of algebra, we get
+\begin{align*}
+	\dot{n} = 0:& \qquad \rho = \frac{\alpha}{\beta}\cdot \frac{n}{n-1} \quad (n\neq 1),\\
+	\dot{\rho} = 0:& \qquad \rho = \frac{\gamma}{\beta^2}n.
+\end{align*}
+
+\begin{observation}
+	The reason that we get the restriction $n\neq 1$ for the $\dot{n} = 0$ nullcline is the following.
+From (E.1.1) we know that $f_1(p^0_2)=0$.
+Thus we can use the implicit function theorem to get a continuous branch of equilibria near $p_2^0$ in the form of $\rho = \hat{\rho}(n)$ where $\hat{\rho}$ is a continuously differentiable function, where $\rho^* = \hat{\rho}(n^*)$ (note $p^0_2 = (n^*, \rho^*)$), and $f_1(n,\hat{\rho}(n)) = 0$ for some neighborhood of $p_0^2$.
+However, we can use this implicit function argument only when $\partial_\rho f_1 \neq 0$, which implies $n\neq 1$.
+\end{observation}
 

### 2024-02-09 Findings (Shared Folder)/Chapters/modelAttempts.tex
new analysis is added

@@
 Furthermore, the following diagram shows the nullclines and the sign of the vector field at the different regions of the phase portrait.
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.5\linewidth]{images/nullClines.png}
-
-	\label{fig:simpleodemodel1phaseportrait}
-\end{figure}
 
@@
 \end{align*}
-
 \begin{observation}
@@
 Thus we can use the implicit function theorem to get a continuous branch of equilibria near $p_2^0$ in the form of $\rho = \hat{\rho}(n)$ where $\hat{\rho}$ is a continuously differentiable function, where $\rho^* = \hat{\rho}(n^*)$ (note $p^0_2 = (n^*, \rho^*)$), and $f_1(n,\hat{\rho}(n)) = 0$ for some neighborhood of $p_0^2$.
-However, we can use this implicit function argument only when $\partial_\rho f_1 \neq 0$, which implies $n\neq 1$.
+However, we can use this implicit function argument only when $\partial_\rho f_1 \neq 0$, which implies $n\neq 1$.\\
+	Also, it is interesting to note that $n=1$ is equivalent to $\alpha=0$. This can be observed from (E.1.1). The consequences of this are summarized in the following observation boxes.
 \end{observation}
+
+We will have two cases of the phase portrait as shown in the figure below.
+Note that the stability of the equilibrium point $p_2^0$ will still remain the same as the value of $\alpha$ passes $\alpha=0$ transversally.
+The results of this section is summarized in the observation box below.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{images/nullClines_positiveNegative.png} % first image
+\end{figure}
+
+\FloatBarrier
+
+\begin{observation}[Two different phase portraits]
+	As the parameter $\alpha$ passes through $\alpha_0 = 0$ transversally, we get two phase portraits that are not topologically equivalent (the results are shown in the figure above).
+Because of this, depending on the sign of $\alpha$, we will observe totally different behaviours from the system as we change the parameter values.
+\\
+	I \textbf{emphasis} that in both of these phase portraits, the stability of both equilibria remains the same.
+\end{observation}
+
+\begin{observation}[Biological meaning of $\alpha$]
+	From $(\clubsuit)$ we see that $\alpha = \lambda_s - \delta_t$, where $\lambda_s$ is the tip cell division rate (which leads to vascular splitting), and $\delta_t$ is the death rate of the tip cells.
+Thus $\alpha >0$ translates to larger division rate compared to the death rate for tip cells, and $\alpha<0$ is the opposite.
+\end{observation}
+
+\subsection*{Quantitative study of Effect of Changing the Parameters}
+This section will lay the foundations for studying the drug-vessel interaction and how that affects the system.
+\subsubsection*{Effect of $\alpha$} 
+Regardless of the sign of $\alpha$ (i.e. being in either of phase portraits) increasing the value of alpha will move the $p^0_2$ higher. This observation is summarized in the following figure.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{images/effectOfAlpha.png} % first image
+\end{figure}
+
+\FloatBarrier
+
+\subsubsection*{Effect of $\gamma$}
 

### 2024-02-09 Findings (Shared Folder)/Chapters/modelAttempts.tex
new analysis added

@@
 \subsubsection*{Effect of $\gamma$}
+$\gamma$ is basically determining the slope of the $\dot{\rho}=0$ nullcline.
+The higher the value of $\gamma$ the more steeper is the slop.
+Thus changing the values of $\gamma$, the equilibrium point $p^0_2$ will move up or down on the $\dot{n}=0$ nullcline.
+The sign of $\alpha$ determines the way $p^0_2$ changes.
+The following figure summarizes the results for the argument above.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{images/effectOfGamma.png} % first image
+\end{figure}
+\FloatBarrier
+
+\subsubsection*{Effect of $\beta$}
+Determining the effect of the parameter $\beta$ is not as straight forward as the other two parameters as it appears in both ODEs.
+However, we can plot the parameterized curve of $p^0_2(\beta)$ using (E.1.1) to see the effect of $\beta$ on the equilibrium point.
+The following figure summarizes the effect of $\beta$ on $p^0_2$.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{images/EffectOfBeta.png} % first image
+\end{figure}
+\FloatBarrier
 

### 2024-03-03 Findings (Shared Folder)/Chapters/PapersReviewed.tex
new changes

@@
 	\item Biological angiogenesis experiments \cite{Bauer2007}
-	\item Topological data analysis \cite{Carlsson2009,Zomorodian2005}
-	\item Applied Algebraic Topology \cite{2020,2023}
-	\item Application of topological data analysis \cite{Nicolau2011,Nielson2015,Qaiser2016,Xia2014}
-	\item Topological data analysis and agent based models \cite{Topaz2015,McGuirl2020}
+%	\item Topological data analysis \cite{Carlsson2009,Zomorodian2005}
+%	\item Applied Algebraic Topology \cite{2020,2023}
+%	\item Application of topological data analysis \cite{Nicolau2011,Nielson2015,Qaiser2016,Xia2014}
+%	\item Topological data analysis and agent based models \cite{Topaz2015,McGuirl2020}
 \end{itemize}
 
-\subsection{Results}
+\section{Quantitative Angiogenesis Assays in vivo – A Review}{J. Hasan}{2004, Angiogenesis}
+
+This paper discusses various angiogenesis assays, highlighting the corneal micropocket and the CAM assay as established methods.
+It emphasizes the importance of selecting complimentary assays to best replicate tumor angiogenesis for studying the effects of pro- or anti-angiogenic compounds.
+The development of non-invasive techniques for quantifying angiogenesis is highlighted as a significant advancement for the field \cite{Hasan2004}
+
+\section{Characterization of lymphocyte-dependent angiogenesis using a SCID mouse: human skin model of psoriasis}{B. Nickoloff}{2000, The journal of investigative dermatology}
+
+This review updates the understanding of angiogenesis in psoriasis, integrating the characterization of endothelial cells in plaques and discussing a novel animal model for triggering neovascularization and plaque formation, providing insights into the angiogenic process in skin disorders \cite{Nickoloff2000}
+
+\section{Integration of experimental and computational approaches to sprouting angiogenesis}{S. Peirce}{2012, Current Opinion in Hematology}
+
+This paper summarizes the integration of experimental tools and computational modeling in studying sprouting angiogenesis, showcasing how such interdisciplinary approaches can lead to new understandings and therapeutic targets by accounting for molecular data and cell-level behaviors \cite{Peirce2012}
+
+\section{Mathematical models of developmental vascular remodelling: A review}{Jessica R. Crawshaw}{2023, PLOS Computational Biology}
+
+Focusing on the less-explored area of developmental remodeling of vascular networks, this review discusses mathematical models that have contributed to understanding the transformation of primitive vessel networks into functional ones, highlighting the multiscale nature of this problem \cite{Crawshaw2023}
+
+\section{Assessment Methods of Angiogenesis and Present Approaches for Its Quantification}{G. J. Khan}{2014, Cancer Research}
+
+This paper provides an overview of angiogenesis assessment methods, including in vitro, in vivo, and in ovo models, focusing on the calculation modes and considerations necessary for concluding the angiogenic or antiangiogenic properties of agents\cite{Khan2014}
+
+\section{Endogenous inhibitors of angiogenesis.}{P. Nyberg, et. al.}{2005, Cancer research}
+
+Highlighting the balance between proangiogenic and antiangiogenic factors, this review explores the role of endogenous inhibitors in the body, providing insight into the potential of leveraging these natural inhibitors for therapeutic purposes in cancer treatment \cite{Nyberg2005}
+
+\section{Mathematical Models of Avascular Tumor Growth}{T. Roose, et. al.}{2007, SIAM Rev}
+
+Offering a comprehensive list and discussion of models for avascular tumor growth, this review emphasizes the importance of mathematical modeling in understanding tumor development and outlines potential future directions for research in this area \cite{Roose2007}
+
+\section{Mathematical Modelling of Angiogenesis}{Chaplain}{2000, Journal of Neuro-Oncology}
+discusses a variety of mathematical models used to describe capillary network formation, focusing on a model that generates 2D and 3D vascular structures.
+This model incorporates the migratory response of endothelial cells to tumor angiogenic factors, cell proliferation, and interactions with extracellular matrix macromolecules, among other factors \cite{Chaplain2000}
+
+\section{Computational and Mathematical Modeling of Angiogenesis}{S. Peirce}{2008, Microcirculation}
+
+reviews mathematical and computational models developed over two decades to study various aspects of angiogenesis.
+This work emphasizes the insights gained from these models in normal physiological growth, tumorigenesis, wound healing, and therapeutic strategy design \cite{Peirce2008}
+
+\section{Mathematical modelling of angiogenesis using continuous cell-based models}{F. D. Bookholt, et. al}{2016, Biomechanics and Modeling in Mechanobiology}
+
+by Bookholt et al.
+(2016) introduces a 3D in vitro model simulating early stages of angiogenesis.
+The model addresses endothelial cell migration due to chemotaxis and durotaxis and includes various proteins impacting angiogenesis \cite{Bookholt2016}
+
+\section{Mathematical modelling of dynamic adaptive tumour-induced angiogenesis: clinical implications and therapeutic targeting strategies}{S.
+McDougall, et.
+al.}{2006, Journal of theoretical biology}
+presents a model that couples vessel growth with blood flow, offering insights into the adaptive and dynamic nature of tumor-induced angiogenesis and identifying new therapeutic targets for tumor management \cite{McDougall2006}
+
+\section{Angiogenesis—Understanding the Mathematical Challenge}{Pamela F Jones}{2006, Angiogenesis}
+
+Explains the mathematical modelling strategy in biological terms, aiming to bridge the gap between mathematics and life sciences.
+The paper discusses the assumptions and simplifications foundational to modeling and their implications for understanding angiogenesis \cite{Jones2006}
+
+\section{On the mathematical modeling of wound healing angiogenesis in skin as a reaction-transport process}{Samik Ghosh}{2015, Frontiers in Physiology}
+
+provides a comprehensive review of mathematical models of angiogenesis in skin wound healing.
+It introduces the continuum reaction-transport framework as a useful tool for exploring unresolved questions in angiogenesis research \cite{Ghosh2015}
+
+\section{Multiscale Agent-based Model of Tumor Angiogenesis}{Megan M, et. al.}{2013}
+
+Olsen and Siegelmann (2013) developed a three-dimensional multiscale ABM focusing on breast cancer.
+The model encompasses cellular (genetic control), tissue (cells, blood vessels, angiogenesis), and molecular (VEGF, diffusion) levels.
+A novel discrete approach to model angiogenesis is proposed to decrease computational cost, offering potential new directions for modeling in cancer research \cite{Olsen2013}
+
+\section{Simulating cancer growth with multiscale agent-based modeling.}{Zhihui Wang, et. al.}{2015, Seminars in cancer biology}
+discuss the utility of ABMs in simulating diverse cancer behaviors, including tumor morphology, adaptation to the microenvironment, angiogenesis, and response to therapies.
+The review highlights the capability of ABMs to simulate the complex interplay between tumor cells and their microenvironment, paving the way for new therapeutic insights \cite{Wang2015}
+
+\section{Agent-based model of angiogenesis simulates capillary sprout initiation in multicellular networks.}{Joseph Walpole, et. al.}{2015, Integrative biology}
+
+present an ABM that incorporates both stochastic and deterministic rules to simulate the initiation of sprouting angiogenesis.
+The model accurately simulates sprout initiation frequency and location, offering a deeper understanding of the balance between stochasticity and determinism in biological processes \cite{Walpole2015}
+
+\section{Agent-Based Modeling of Vascularization in Gradient Tissue Engineering Constructs}{E. S. Bayarak, et. al.}{2015, IFAC-PapersOnLine}
+
+develop an ABM to simulate vascular growth in engineered biomaterials, investigating the influence of growth factor release rate on angiogenesis.
+The model's results, validated against experimental studies, suggest microsphere properties that promote angiogenesis, offering insights into tissue engineering applications \cite{Bayrak2015}
+
+\section{A cell-based model exhibiting branching and anastomosis during tumor-induced angiogenesis.}{A. Bauer, et. al.}{2007, Biophysical journal}
+describe a cell-based ABM that integrates endothelial cell migration, growth, division, and the evolving structure of the stroma at the cellular Potts model level.
+The model successfully reproduces various morphologies of capillary sprouts observed in vivo, demonstrating the emergence of branching and anastomosis without prescribed rules \cite{Bauer2007a}
+
+\section{Coupled mathematical model of tumorigenesis and angiogenesis in vascular tumours}{M. Cooper, et. al.}{2010, Cell Proliferation}
+
+developed a model that combines the processes of avascular tumor growth and the development of capillary networks through tumor-induced angiogenesis.
+This comprehensive model offers insights into the growth and development mechanisms of vascular tumors \cite{Cooper2010}
+
+\section{Tree topology analysis of the arterial system model}{V. Kopylov}{2018, Journal of Physics}
+
+presented an algorithm for constructing an arterial system model with physiologically significant geometric properties.
+Their analysis of the bifurcation exponent's effect on the arterial network's topology provides valuable insights into the optimal network topology for efficient vascular function  \cite{Kopylova2018}
+
+\section{Mathematical Model of Blood Flow in an Anatomically Detailed Arterial Network of the Arm}{Sansuke M, el. al.}{2013, Mathematical Modelling and Numerical Analysis}
+
+Watanabe, Blanco, and Feijóo (2013) developed a detailed model for hemodynamics simulations in the arm's arterial network.
+Their model includes a comprehensive arterial topology and offers a systematic estimation of involved parameters, allowing for accurate simulations of blood flow and pressure \cite{Watanabe2013}
+
+\section{An integrated approach to quantitative modelling in angiogenesis research}{A. J. Connor, et. al.}{2015, Journal of The Royal Society Interface}
+
+discuss a multidisciplinary approach combining experiments, image processing, analysis, and mathematical modeling focused on angiogenesis in the cornea micropocket assay.
+This approach aims to provide mechanistic insights into the action of angiogenic factors through quantitative data extraction and model parametrization \cite{Connor2015}
+
+\section{Integration of experimental and computational approaches to sprouting angiogenesis}{S. Peirce, et. al.}{2012, Current Opinion in Hematology}
+
+Peirce et al.
+(2012) summarize recent advancements in computational modeling of angiogenesis, driven by detailed molecular data and experimental tools.
+These models help predict hypothetical experiment outcomes and generate new hypotheses for understanding angiogenesis at a system-wide level \cite{Peirce2012}
+
+\section{A Computational Tool for Quantitative Analysis of Vascular Networks}{E. Zudaire, et. al.}{2011, PLoS ONE}
+developed AngioTool, a user-friendly software for the quantification of vascular networks in microscopic images.
+AngioTool computes several morphological and spatial parameters and is open source, available for free download, facilitating standardized analysis in angiogenesis research \cite{Zudaire2011}
+
+\section{Consensus guidelines for the use and interpretation of angiogenesis assays}{Many authors}{2018, Angiogenesis}
+published the first edition of consensus guidelines for the use and interpretation of angiogenesis assays.
+This collaborative work aims to serve as a reference for current and future angiogenesis research, promoting standardized methodologies across the field \cite{NowakSliwinska2018a}
+
+\section{Zebrafish as an Emerging Model Organism to Study Angiogenesis in Development and Regeneration}{Myra N Chávez, et. al.}{2016, Front Physiol}
+ocus on the zebrafish (Danio rerio) as an emerging model organism for studying angiogenesis in development and regeneration, highlighting its potential for understanding vascularization in artificial tissues and organs, as well as for drug discovery \cite{Chavez2016}
 

### 2024-03-03 Findings (Shared Folder)/Chapters/modelAttempts.tex
new changes

@@
 From (E.1.1) we know that $f_1(p^0_2)=0$.
-Thus we can use the implicit function theorem to get a continuous branch of equilibria near $p_2^0$ in the form of $\rho = \hat{\rho}(n)$ where $\hat{\rho}$ is a continuously differentiable function, where $\rho^* = \hat{\rho}(n^*)$ (note $p^0_2 = (n^*, \rho^*)$), and $f_1(n,\hat{\rho}(n)) = 0$ for some neighborhood of $p_0^2$.
+Thus we can use the implicit function theorem to get a continuous branch of equilibria near $p_2^0$ in the form of $\rho = \hat{\rho}(n)$ where $\hat{\rho}$ is a continuously differentiable function, where $\rho^* = \hat{\rho}(n^*)$ (note $p^0_2 = (n^*, \rho^*)$), and $f_1(n,\hat{\rho}(n)) = 0$ for some open neighborhood containing $p_2^0$.
 However, we can use this implicit function argument only when $\partial_\rho f_1 \neq 0$, which implies $n\neq 1$.\\
@@
 
+\section{Simple Spatially Distributed 1D System}
+
+In this section, we extend a simple ordinary differential equation (ODE) model for angiogenesis to include spatial dynamics.
+The original model captures the temporal evolution of the density of tip cells, the density of blood vessels, and the concentration of a drug delivered to a region.
+We now consider these variables as functions of both space and time, \(\rho(x,t)\), \(n(x,t)\), and \(c(x,t)\), to develop a partial differential equation (PDE) model that accounts for spatial growth and diffusion processes.
+
+\subsection{Flux of Tip Cells}
+A key concept in extending the model to incorporate spatial dynamics is the introduction of the flux of tip cells, denoted as \(J\).
+Flux is defined as the rate of flow of a property per unit area, which in this context is the flow of tip cells moving into a region.
+Mathematically, the flux \(J\) is given by the product of the density of tip cells \(n(x,t)\) and their velocity \(v\), i.e., \(J = nv\).
+This formulation allows us to quantify how the movement of tip cells contributes to the spatial development of blood vessels.
+
+\subsection{Spatial Model Development}
+We now proceed to develop the spatial model by modifying the equations from the original ODE model to incorporate spatial derivatives, reflecting the spatial dynamics of vessel formation, tip cell movement, and drug diffusion.
+
+\subsubsection{Equation for Blood Vessel Density}
+The blood vessel density \(\rho(x,t)\) is primarily affected by the extension of vascular structures due to the movement of tip cells and the degradation of these structures over time.
+Assuming homogeneous conditions without specific spatially-dependent growth mechanisms, the equation for \(\rho\) remains similar to the non-spatial model but now includes a spatial component:
+
+\begin{equation}
+	\frac{\partial \rho}{\partial t} = vn - \delta_v \rho,
+\end{equation}
+where \(\delta_v\) represents the rate of vascular degradation.
+
+\subsubsection{Equation for Tip Cell Density}
+The spatial dynamics of tip cell density \(n(x,t)\) are influenced by their flux across the spatial domain.
+Incorporating the concept of flux, the balance equation for \(n\) in one dimension is given by:
+
+\begin{equation}
+	\frac{\partial n}{\partial t} + \frac{\partial (nv)}{\partial x} = (\lambda_s - \delta_t) n + \lambda_b \rho - \kappa n \rho,
+\end{equation}
+where the terms on the right-hand side represent the generation and loss of tip cells, and the spatial derivative term accounts for their movement.
+
+\subsubsection{Equation for Drug Concentration}
+The drug concentration \(c(x,t)\) is affected by diffusion, permeation through the capillary walls, and any external sources of the drug. The equation accounting for these processes is:
+
+\begin{equation}
+	\frac{\partial c}{\partial t} = D\frac{\partial^2 c}{\partial x^2} - \mu c + f(x,t),
+\end{equation}
+where \(D\) is the diffusion coefficient, \(\mu\) represents the drug's permeability through capillaries, and \(f(x,t)\) is a source term for the drug.
+
 \newpage

### 2024-03-10 Findings (Shared Folder)/Chapters/MeetingLog.tex
new changes
new comments for the mol_bio

@@
 
+\subsection{12 Feb Meeting}
+\begin{itemize}
+	\item Add the axis labels for the qualitative analysis, and edit the title.
+	\item Consider the non-dimensionalization process to see if you can control the non-linear term.
+	\item Add the section for drug-tissue interaction. Consider the possible simplifications of the model.
+\end{itemize}
+
+\subsection{8 March Meeting}
+\begin{itemize}
+	\item To develop the PDEs for the model: Note the following facts: The blood vessels are not moving, but they are deposited by the moving endothelial cells.
+	\item There should be no diffusion or advection terms in the PDE for $ \rho $.
+	\item Although the tip cells are diffusing as well, but we can consider their dynamics to be advection dominated diffusion. Thus no diffusion term at the moment.
+	\item The advection speed of the tip cells is the same as the speed that they move in the bulk model.
+	\item Try the traveling wave solution to the derived PDE.
+	\item Be careful about the terms that can be simplified by the time scale separation argument. 
+\end{itemize}
+
 \section{Meetings with Arman}

### 2024-03-10 Findings (Shared Folder)/Chapters/MolecularBiology.tex
new comments for the mol_bio

@@
 \section{Molecular Mechanism of Angiogenesis}
+
+Blood vessels and the vascular structure are formed by the differentiation of the cells in the mesoderm layer during the embryo development (the layer which also give rise to blood cells, kidney, liver, connective tissue, etc.) \cite{Alberts2002}.
+
+\subsection{A Brief Anatomy of Vessels}
+Endothelial cells line all of the vessels.
+Blood vessels (like the arteries and the veins that are the largest vessels of the body) have a thick and tough wall of connective tissue with several layers of smooth muscles.
+The wall is lined by a very thin layer of endothelial cells (i.e.
+the endothelium) separated from the outer surrounding layers by basal lamina \cite{Alberts2002}.
+It is worth noting that the amount of connective tissue and smooth muscle depends on the diameter of the blood vessel as well as its function, \textbf{but the endothelial lining is always present}.
+In the finest branches of the vasculature (i.e.
+capillaries and sinusoids) the wall is just made up of endothelial cells and basal lamina.
+One of the major roles of the endothelial cells is to control to transport of material in an out of the bloodstream.
+
+A study of embryo development reveals that the even larger vessels (like arteries and veins) start developing from smaller vessels that has only endothelial cells and basal lamina.
+The connective tissue, smooth muscles and pericytes are added later on, by the signaling from endothelial cells.
+In particular, the recruitment of pericytes are driven by PDGF (platelet driven growth factor) secreted by the endothelial cells.
+
+\begin{figure}[h!]
+	\centering
+	% First bottom figure, occupying the left half
+	\begin{minipage}{0.4\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{images/AnatomyOfVessel.jpg}
+		% Optional caption for the first bottom image
+		% \caption{Caption for the first bottom image.}
+		% \label{fig:bottom-left-image}
+	\end{minipage} % Space between the two bottom figures
+	% Second bottom figure, occupying the right half
+	\begin{minipage}{0.4\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{images/perycitesForVascular.jpg}
+		% Optional caption for the second bottom image
+		% \caption{Caption for the second bottom image.}
+		% \label{fig:bottom-right-image}
+	\end{minipage}
+\end{figure}
+\vspace{-15pt}
+\begin{figure}[h!]
+	% Top figure spanning the whole width
+	\centering
+	\includegraphics[width=0.8\textwidth]{images/ElectronMicroscopeOfVessel.jpg}
+	% Optional caption for the top image
+	\caption{\textbf{Figure Top Left:} This figure shows the anatomy of a large vessel, like vein or arteries.
+Note that smaller vessels, like capillaries as well as sinusoids consists of only endothelial cells and basal lamina, except for some scattered pericytes wrapped around the walls (see figure Top Left).
+\textbf{Figure Top Right:} Electron micro graph showing small pericytes wrapped around small blood vessels.
+\textbf{Figure Bottom Left:} A capillary that its wall consists of only endothelial cell and basal lamina.
+\textbf{Figure Bottom Right:} Electron micro graph showing a cross section of small capillary in pancreas.
+All of the figures are from \cite{Alberts2002}}
+
+\end{figure}
+\FloatBarrier
+
+Also, the following figure summarizes the cross section of different types of vasculature.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.8\textwidth]{images/CrossSectionOfVessels.png}
+	\caption{The cross section of vessels in the form of arteries, capillaries, and vein. Note the single lining of the endothelial cells for the capillary.}
+\end{figure}
+
+\subsection{Molecular Biology of Vascular Structure}
+
+New vessels in the adults originate as capillaries, which sprout from the existing small vessels.
+Endothelial cells on the arterial and venous side of the developing networks of vessels differ in their surface properties.
+In the embryo at least, the plasma membrane of the arterial cells contains trans membrane protein ephrine-B2, while the membrane of the venous cells contain the corresponding receptor protein Eph-B4, which is a receptor tyrosine kinase.
+These molecules mediate a signal delivered at sites of cell-cell contact, and they are essential for the development of a properly organized network of vessels.
+One suggestion is that they somehow define the rules for joining one piece of growing capillary tube to another \cite{Alberts2002}.
+
  There are two general balancing forces acting on the angiogenesis
@@
 
+Endothelial cells that are to form a new capillary, grow out from the side of an existing capillary by forming long pseudopodia pioneering the formation of new capillary sprout that hallows out to form a tube.
+This process continues until the sprout encounters another capillary, where they merge.
+
 \subsubsection*{Controlling Capillary Joining Process}
-In the following text from \cite{Alberts2002}, there is some vague hints about the mechanisms that are controlling capillary joining to each other.
+In the following text from \cite{Alberts2002}, there is some vague hints about the mechanisms that are controlling capillary joining to each other
 
@@
 This process has also been described in \cite{angiogenesisYoutube}.
-\begin{quote}
-	Experiments in culture show that endothelial cells in a medium containing suitable growth factors will spontaneously form capillary tubes, even if they are isolated from all other types of cells (Figure 22-26).
+	\begin{wrapfigure}{r}{0.6\textwidth}
+	\centering
+	\includegraphics[width=0.6\textwidth]{images/hallowStructureEndothelialCells.jpg}
+	\caption{The endothelial cells, when supported by suitable growth medium and signals, start to form hallow structure, that do not contain any blood, and not fluid passes through them.
+This indicates that the no mechanical trigger (i.e.
+pressure) is required to form the hallow structure for the new vessels.
+Image from \cite{Alberts2002}.}
+\end{wrapfigure}
+
+	``Experiments in culture show that endothelial cells in a medium containing suitable growth factors will spontaneously form capillary tubes, even if they are isolated from all other types of cells (Figure 22-26).
 The capillary tubes that develop do not contain blood, and nothing travels through them, indicating that blood flow and pressure are not required for the initiation of a new capillary network.
-
 Endothelial cells in culture spontaneously develop internal vacuoles that appear to join up from cell to cell, giving rise to a network of capillary tubes.
-These photographs show successive stages in the process.
-\end{quote}
+These photographs show successive stages in the process.''
 

### 2024-03-10 Findings (Shared Folder)/Chapters/PapersReviewed.tex
Updates from Overleaf

@@
 \section{Mathematical Modelling of Angiogenesis}{Chaplain}{2000, Journal of Neuro-Oncology}
-discusses a variety of mathematical models used to describe capillary network formation, focusing on a model that generates 2D and 3D vascular structures.
+Discusses a variety of mathematical models used to describe capillary network formation, focusing on a model that generates 2D and 3D vascular structures.
 This model incorporates the migratory response of endothelial cells to tumor angiogenic factors, cell proliferation, and interactions with extracellular matrix macromolecules, among other factors \cite{Chaplain2000}
@@
 
-reviews mathematical and computational models developed over two decades to study various aspects of angiogenesis.
+Reviews mathematical and computational models developed over two decades to study various aspects of angiogenesis.
 This work emphasizes the insights gained from these models in normal physiological growth, tumorigenesis, wound healing, and therapeutic strategy design \cite{Peirce2008}
@@
 
-by Bookholt et al.
+By Bookholt et al.
 (2016) introduces a 3D in vitro model simulating early stages of angiogenesis.
@@
 al.}{2006, Journal of theoretical biology}
-presents a model that couples vessel growth with blood flow, offering insights into the adaptive and dynamic nature of tumor-induced angiogenesis and identifying new therapeutic targets for tumor management \cite{McDougall2006}
+Presents a model that couples vessel growth with blood flow, offering insights into the adaptive and dynamic nature of tumor-induced angiogenesis and identifying new therapeutic targets for tumor management \cite{McDougall2006}
 
@@
 
-provides a comprehensive review of mathematical models of angiogenesis in skin wound healing.
+Provides a comprehensive review of mathematical models of angiogenesis in skin wound healing.
 It introduces the continuum reaction-transport framework as a useful tool for exploring unresolved questions in angiogenesis research \cite{Ghosh2015}
@@
 \section{Simulating cancer growth with multiscale agent-based modeling.}{Zhihui Wang, et. al.}{2015, Seminars in cancer biology}
-discuss the utility of ABMs in simulating diverse cancer behaviors, including tumor morphology, adaptation to the microenvironment, angiogenesis, and response to therapies.
+Discusses the utility of ABMs in simulating diverse cancer behaviors, including tumor morphology, adaptation to the microenvironment, angiogenesis, and response to therapies.
 The review highlights the capability of ABMs to simulate the complex interplay between tumor cells and their microenvironment, paving the way for new therapeutic insights \cite{Wang2015}
@@
 
-present an ABM that incorporates both stochastic and deterministic rules to simulate the initiation of sprouting angiogenesis.
+Presents an ABM that incorporates both stochastic and deterministic rules to simulate the initiation of sprouting angiogenesis.
 The model accurately simulates sprout initiation frequency and location, offering a deeper understanding of the balance between stochasticity and determinism in biological processes \cite{Walpole2015}
@@
 
-develop an ABM to simulate vascular growth in engineered biomaterials, investigating the influence of growth factor release rate on angiogenesis.
+Develops an ABM to simulate vascular growth in engineered biomaterials, investigating the influence of growth factor release rate on angiogenesis.
 The model's results, validated against experimental studies, suggest microsphere properties that promote angiogenesis, offering insights into tissue engineering applications \cite{Bayrak2015}
@@
 \section{A cell-based model exhibiting branching and anastomosis during tumor-induced angiogenesis.}{A. Bauer, et. al.}{2007, Biophysical journal}
-describe a cell-based ABM that integrates endothelial cell migration, growth, division, and the evolving structure of the stroma at the cellular Potts model level.
+Describes a cell-based ABM that integrates endothelial cell migration, growth, division, and the evolving structure of the stroma at the cellular Potts model level.
 The model successfully reproduces various morphologies of capillary sprouts observed in vivo, demonstrating the emergence of branching and anastomosis without prescribed rules \cite{Bauer2007a}
@@
 
-developed a model that combines the processes of avascular tumor growth and the development of capillary networks through tumor-induced angiogenesis.
+Developed a model that combines the processes of avascular tumor growth and the development of capillary networks through tumor-induced angiogenesis.
 This comprehensive model offers insights into the growth and development mechanisms of vascular tumors \cite{Cooper2010}
@@
 
-presented an algorithm for constructing an arterial system model with physiologically significant geometric properties.
+Presented an algorithm for constructing an arterial system model with physiologically significant geometric properties.
 Their analysis of the bifurcation exponent's effect on the arterial network's topology provides valuable insights into the optimal network topology for efficient vascular function  \cite{Kopylova2018}
@@
 
-discuss a multidisciplinary approach combining experiments, image processing, analysis, and mathematical modeling focused on angiogenesis in the cornea micropocket assay.
+Discusses a multidisciplinary approach combining experiments, image processing, analysis, and mathematical modeling focused on angiogenesis in the cornea micropocket assay.
 This approach aims to provide mechanistic insights into the action of angiogenic factors through quantitative data extraction and model parametrization \cite{Connor2015}
@@
 \section{Consensus guidelines for the use and interpretation of angiogenesis assays}{Many authors}{2018, Angiogenesis}
-published the first edition of consensus guidelines for the use and interpretation of angiogenesis assays.
+Published the first edition of consensus guidelines for the use and interpretation of angiogenesis assays.
 This collaborative work aims to serve as a reference for current and future angiogenesis research, promoting standardized methodologies across the field \cite{NowakSliwinska2018a}
@@
 \section{Zebrafish as an Emerging Model Organism to Study Angiogenesis in Development and Regeneration}{Myra N Chávez, et. al.}{2016, Front Physiol}
-ocus on the zebrafish (Danio rerio) as an emerging model organism for studying angiogenesis in development and regeneration, highlighting its potential for understanding vascularization in artificial tissues and organs, as well as for drug discovery \cite{Chavez2016}
+Focuses on the zebrafish (Danio rerio) as an emerging model organism for studying angiogenesis in development and regeneration, highlighting its potential for understanding vascularization in artificial tissues and organs, as well as for drug discovery \cite{Chavez2016}
 

### 2024-03-10 Findings (Shared Folder)/Chapters/MolecularBiology.tex
new updates to biology notes

@@
 
+\begin{observation}
+	The difference in the surface properties of endothelial cells on the arterial and venous side of the developing networks of vessels control the rate an which one piece of growing capillary tube joins another.
+This becomes very interesting if we consider it along the observations in \cite{Koery2024}.
+They observed that the blunt-ended capillaries with small diameter are more susceptible for degradation after irradiation.
+And since the presence of blunt-ended vessels with small diameter increase the flow resistance of the network, pruning these branches ``normalizes'' the blood flow, hence increase the perfusion after irradiation.
+\end{observation}
+
+\subsubsection{Steps involved in angiogenesis}
+Individual endothelial cells responds to the signals produced by the organ that they invade.
+The signal is complex, but the main part of the signal is vascular endothelial growth factor (\textbf{VEGF}) (which is a distant relative of platelet driven growth factor (\textbf{PDGF})).
+The control on the production of VEGF is through its mRNA stability and its rate of transcription.
+Under a low oxygen concentration, the  intracellular concentration of an active form of gene regulatory protein called \textbf{hypoxia inducible factor 1 (HIF-1)} increases.
+HIF-1 stimulates the transcription of VEGF gene (and the production of other genes that are needed when the oxygen supply is low).
+When the VEGF protein is secreted, it is then diffuses through the tissue and acts on nearby endothelial cells.
+
+Endothelial cells that are to form a new capillary, grow out from the side of an existing capillary by forming long pseudopodia pioneering the formation of new capillary sprout that hallows out to form a tube.
+This process continues until the sprout encounters another capillary, where they merge.
+In the tumor micro environment, The growth rate of tumor increases abruptly as soon as the vessels reach it.
+
 There are two general balancing forces acting on the angiogenesis
@@
 \end{itemize}
+\FloatBarrier
 
-Endothelial cells that are to form a new capillary, grow out from the side of an existing capillary by forming long pseudopodia pioneering the formation of new capillary sprout that hallows out to form a tube.
-This process continues until the sprout encounters another capillary, where they merge.
+\subsubsection{The Response of Endothelial Cells to VEGF}
+The response of endothelial cells to VEGF has four components.
+First, they produce proteases to digest through the basal lamina of the parent vessels.
+For the second step, they migrate towards the source of VEGF, and for the third step they proliferate.
+Finally, they form hallow tubes.
+It is worth mentioning that VEGF stimulates endothelial cells selectively, while other angiogens, like fibroblast growth factor stimulates other cell types as well.
+The following figure summarizes these steps.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.8\textwidth]{images/endothelialResponseToVEGF.jpg}
+	\caption{A summary of the response of the endothelial cells to VEGF.
+Under low oxygen concentration, the intracellular concentration of HIF-1 increases.
+This gene regulatory protein in turn increases the transcription of VEGF protein.
+Then VEGF diffuses through the tissue and stimulates the endothelial cells lining the vessels.
+Figure is from \cite{Alberts2002}.}
+\end{figure}
 
@@
 \subsubsection*{Formation of tube structures by endothelial cells}
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.8\textwidth]{images/hallowStructureEndothelialCells.jpg}
+	\caption{The endothelial cells, when supported by suitable growth medium and signals, start to form hallow structure, that do not contain any blood, and not fluid passes through them.
+This indicates that the no mechanical trigger (i.e.
+pressure) is required to form the hallow structure for the new vessels.
+Image from \cite{Alberts2002}.}
+\end{figure}
+
 It was one of my main concerns that what is the process in which a single lining of endothelial cells following a tip cell forms a hallow tube (i.e.
@@
 This process has also been described in \cite{angiogenesisYoutube}.
-	\begin{wrapfigure}{r}{0.6\textwidth}
-	\centering
-	\includegraphics[width=0.6\textwidth]{images/hallowStructureEndothelialCells.jpg}
-	\caption{The endothelial cells, when supported by suitable growth medium and signals, start to form hallow structure, that do not contain any blood, and not fluid passes through them.
-This indicates that the no mechanical trigger (i.e.
-pressure) is required to form the hallow structure for the new vessels.
-Image from \cite{Alberts2002}.}
-\end{wrapfigure}
 
-	``Experiments in culture show that endothelial cells in a medium containing suitable growth factors will spontaneously form capillary tubes, even if they are isolated from all other types of cells (Figure 22-26).
+\begin{quote}
+	Experiments in culture show that endothelial cells in a medium containing suitable growth factors will spontaneously form capillary tubes, even if they are isolated from all other types of cells (Figure 22-26).
 The capillary tubes that develop do not contain blood, and nothing travels through them, indicating that blood flow and pressure are not required for the initiation of a new capillary network.
 Endothelial cells in culture spontaneously develop internal vacuoles that appear to join up from cell to cell, giving rise to a network of capillary tubes.
-These photographs show successive stages in the process.''
+These photographs show successive stages in the process.
+\end{quote}
 

### 2024-03-17 Findings (Shared Folder)/Chapters/modelAttempts.tex
new notes added

@@
 \end{equation*}
+
+\begin{observation}[Other form of non-dimensionalization]
+	After inspecting the non dimensionalization steps, we can see that the parameter $ \kappa $ somehow disappears, and there is no sign of it in the new parameters $ \alpha, \beta, $ and $ \gamma $.
+I tried another choices for $ R, T, $ and $ N $, but still, I can not have the $ \kappa $ in the new parameters.
+The steps that I followed are the following.
+	\[ T \kappa R = 1, \qquad NvT/R = 1, \qquad \delta_v T = 1. \]
+	Then we will have
+	\[ T = 1/\delta_v, \qquad R = \delta_v/\kappa, \qquad N = \delta_v^2/(v\kappa). 
+	\]
+	Thus the system of ODEs will be
+	\[ \dot{\rho} = n - \rho, \qquad \dot{n}= \alpha/\beta n - n\rho + \gamma/\beta^2 p, \]
+	where
+	\[ \alpha = \lambda_s - \delta_t, \qquad \beta=\delta_v, \qquad \gamma = \lambda_b v. \]
+	Still, there is not $ \kappa $ in the new parameters. 
+\end{observation}
+
 In order to find the equilibrium points, we demand $\dot{n} = 0$ as well as $\dot{\rho} = 0$. This will lead to the following equations

### 2024-03-20 Findings (Shared Folder)/Chapters/MolecularBiology.tex
new notes added

@@
 This is a video from JOVE explaining the details of the protocol \citep{conealMicroPocketAssayJOVE}
+
+\section{Some Histology}
+
+In short, histology is the study of the animal tissue in the microscopic scale (which is also known as the microscopic anatomy or micro anatomy).
+Studying different types of animal tissue falls in the realm of histology.
+
+\noindent There are four types of animal tissue 
+\begin{enumerate}[(i),noitemsep]
+	\item Epithelium
+	\begin{itemize}
+		\item squamous: endothelial lining of the vascular structure is of this type.
+		\item cuboidal
+		\item columnar
+	\end{itemize}
+	\item Muscle tissue
+	\begin{itemize}
+		\item smooth muscle
+		\item skeletal muscle
+		\item cardiac muscle
+	\end{itemize}
+	\item Connective tissue
+	\begin{itemize}
+		\item cartilage
+		\item bone
+		\item blood
+		\item lymph
+		\item hemopoietic
+	\end{itemize}
+	\item Nervous tissue
+	\begin{itemize}
+		\item central nervous system
+		\item peripheral nervous system
+	\end{itemize}
+\end{enumerate}
+
+Among this list of the four basic types of the animal tissue, we will focus on the Epithelium.
+
+\subsection{Epithelium}
+Epithelium forms continuous sheets of cells that line internal surfaces and cover the external surfaces of the organs.
+A \textbf{basement membrane} separates an epithelium from the underlying connective tissue.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.8\linewidth]{images/histologySquamousEpithelial}
+	\captionsetup{width=0.9\textwidth} 
+	\caption{A) A microscopic image of renal corpuscle that contains a glomerulus (a tuft of capillaries) surrounded by Bowman's capsule.
+The interior of the capsule, is lined with a simple squamous epithelial that rests on a thick basement membrane.
+The only part of these cell visible is their nuclei bulging into the interior.
+B) Scanning electron microscope of renal corpuscle that its glomerulus is removed.
+The simple squamous epithelial can be seen in blue (borders of individual cells are not visible).
+Both images are from histologyguide.com.
+}
+	\label{fig:histologysquamousepithelial}
+\end{figure}
+\FloatBarrier
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.8\linewidth]{images/bileDuct-SquamousEpithelialCells.png}
+	\captionsetup{width=0.9\textwidth} 
+	\caption{A pathology image of bile duct (the large lumen at the center).
+There are many blood vessels in the surrounding connective tissue.
+Blood vessels are lined with simple squamous epithelial.
+The only par of these cells visible is their flatten nuclei.
+\textbf{Epithelium that lines blood vessels, heart, and lymphatic vessels is also known as endothelium.}}
+
+\end{figure}

### 2024-03-20 Findings (Shared Folder)/Chapters/MeetingLog.tex
new notes added
new section for feedbacks

@@
 
+\subsection{12 March Meeting}
+\begin{itemize}
+	\item Things to discuss: going through the qualitative study once again, discussing why we can not have $ \kappa $ directly appearing in the non-dimensionalized version, discussing the effect of $ \kappa $ on the dynamics.
+
+	\item 
+	\[ \frac{\partial n}{\partial t} = D \frac{\partial^2 n}{\partial x^2} - \frac{\partial(nv)}{\partial x} + (\lambda_s - \delta_t)n - \kappa_2 n^2. \]
+	This can be thought of as a model that $ \lambda_b \to 0 $.
+\end{itemize}
+
+\subsection{20 March Meeting}
+\begin{itemize}
+	\item Jupiter: More analysis on the vertical asymptotic.
+	\item Laurent: Drawing a diagram summerizing all of the different interactions happening
+	\begin{itemize}
+		\item Effect of drug on vasculature
+		\item Effect of vasculature on tumor response
+		\item Effect of vasculature on drug delivery
+		\item what is the topology of the reaction graph?
+	\end{itemize}
+	\item Jupiter: adaptive systems
+	\item Leah:
+\end{itemize}
+
 \section{Meetings with Arman}

### 2024-03-21 Findings (Shared Folder)/Chapters/MeetingLog.tex
new comments

@@
 		\item Effect of vasculature on drug delivery
+		\item Effect of tumor on all of these
 		\item what is the topology of the reaction graph?
@@
 	\item Jupiter: adaptive systems
-	\item Leah:
+	\item Leah: (1) I agree that vessel density is a good metric, and I suppose that often this is more easy to measure in scans of images than any other network property.
+I seriously doubt that you can estimate the number of loops or tips in 2D images for all but the very low density of vessels.
+	\begin{enumerate}[(i)]
+
+	\item Eric asked: How do veins and arteries connect up. Good to look into this.
+
+	\item He asked what causes more/less resistance to radiotherapy. Eric asked if you were thinking about devising some fractionation protocol that would result in tumor reduction overall.
+
+	There was also a question about the effect of hypoxia - pro or anti tumor, pro or anti vessel growth etc.
+
+	\item Tim asked if you were considering vessels to be tubes. I think it's OK to assume that the tubes emerge once the branches grow.
+
+	\item Eric was asking what the model is describing, e.g.
+what timescale, etc.
+Do the vessels degrade over such timescale?
+Or do you want to consider a model where the growth rate v goes to 0 at steady state?
+[LEK: we are currently exploring a basic simple model with all parameters assumed to be constants..
+but the eventual goal is to include the effect of tumor, and drug..
+etc, and feedback between treatment and vessels.]
+
+	\item You did not yet mention the drug c, and how you might include it.
+
+	\item Talk to Tim about his microtubule branching ABM.
+
+	\item Do you really want to nondimensionalize the model?
+Jupiter as confused why n --> 1 always.
+It's due to the scaling.
+If you keep the original parameters, you can see how they affect the n and rho steady states.
+
+	\item Jimmy said so far you focused on the angiogenesis, what about the tumor. It would be good to mention that this will be considered later.
+
+	\item Laurent: will drug feed back on vasculature? will tumor affect vasculature?
+
+	He and Jack also asked if tips diffuse.
+
+	[We do not want to redo the model by the Helen Byre group, but we could include later the chemotaxis of tips towards the tumor, as well as a bit of tip diffusion.]
+
+	\item Talk to Jupiter about adaptive systems.
+
+	\item General comments from me:
+	\begin{itemize}
+
+	\item On every slide you show, you need to give credit to where you got images, tables, data, graphs etc that you did not create yourself, even if available on the web.
+It can be in small font, but it is essential in all talks.
+
+	\item I know you are still learning about TDA.
+But I am doubtful whether it will be helpful to us since (1) we have few images, if any to analyse (2) the images are all in 2D, so vessels overlap or disappear into the 3rd dimension (what you called blunt ends) making it tricky or impossible to detect true loops or other features.
+Density seems reasonable.
+	\end{itemize}
+\end{enumerate}
 \end{itemize}

### 2024-03-23 Findings (Shared Folder)/Chapters/randomNotes.tex
new notes added

\chapter{Some Notes on Mathematical Modeling}
These are the steps for mathematical modeling that I have gathered by thinking deep and observing what is happening around me. How my supervisors attack a problem, what is the thought process of professor X in dealing with a problem at one glance, what are the concerns about a specific type of problem (like Stock's problem in PDE theory which is a saddle-node type problem), and what are the main tools to address that problem. These steps in modeling a phenomena mathematically gives me a sense of orientation in a busy world full of new things. This section will be developed overtime, and at some point it can be a chapter of my thesis.

\begin{summary}
	The primary purpose of these notes is to locate the stage of any ongoing project. Also, I will do my best to prepare a kind of checklist for each step that can be used to evaluate the progress of the sub-steps of the project.
\end{summary}

\section{STEP 0: Understanding the phenomena we want to study}

\section{STEP I: Converting Natural Phenomena to a Mathematical Problem }
This step is one of the steps that looks very strange for me. Early from calculus courses and elementary ODE, PDE courses, we can see phrase like ``this Dirichlet boundary condition along with an elliptic PDE models a rod that its two ends are covered with ice at temperature 0'', or ``This saddle-point problem named Stokes problem models a non-compressible fluid''. But they never talk about why this is true, and even how this is possible? Why a bunch of numbers and functions should mimic a very complex phenomena?

There are some very interesting papers written to explain why the mathematics is so effective in modeling the natural world (see \autoref{fig:mathmodelingmatheffectiveness}). In my opinion, this step, although innocent looking, is a totally non-trivial step. Because the nature is what it is, and the fact that some abstract notions and objects can mimic those natural phenomena with nuance details (like the things that happens in bifurcation) is very strange.

At this step, we need to first \textbf{identify the questions} that we want to answer. This is the most important step that affects all of the subsequence analysis. We need to ask ourselves \textbf{to what level of accuracy} we want our answer? \textbf{To what level of generality} we want our answer be?

This step is the step in which we make our decisions. We decide what details to include or ignore. We decide what are the concrete questions we want to answer. This will determine the right mathematical tools that we can use.

\begin{summary}[Step I: Converting a natural phenomena to a mathematical problem]
	$ \; $
	\begin{enumerate}[(i)]
		\item The most important thing: What is the question that we want to answer?
		\item We make lots of decisions:
		\begin{itemize}
			\item What is the precision in the answer that we are looking for?
			\item What is the generality that we are after?
			\item What is an appropriate mathematical tool that can be used to model?
		\end{itemize}

	\end{enumerate}
\end{summary}

\begin{figure}
	\centering
	\captionsetup{width=.9\linewidth}
	\includegraphics[width=1\linewidth]{images/EffectinvenessOFMath.pdf}
	\caption{A series of papers written after Wigner's famous article published in 1960 \cite{Wigner1995}. Several famous scientists of the time responded to this paper, by expanding the ideas to particular fields, among which are \cite{Hamming1980},  \cite{Lesk2000},  \cite{Tegmark2008}, and \cite{GrattanGuinness2008}. }
	\label{fig:mathmodelingmatheffectiveness}
\end{figure}

\FloatBarrier

\section{STEP II: Mathematical Analysis of the Model in Hand}
After doing the STEP I appropriately, we will have a mathematical description of the phenomena in our hand. This mathematical formulation on the problem can come in my different flavors and forms, We can categorize them by considering their different aspects, like being a discrete or continuous model, being a deterministic or stochastic model, etc. Each of these formulations will have its own appropriate way of treating. So I will discuss their subsequence steps separately.

\subsection{STEP II: PDE models}
After converting the natural phenomena to a mathematical model containing PDEs, we will have a domain, and a PDE. For instance, let $ \Omega \subset \R^2 $, and we define
\begin{align*}
	-\Delta u  &= g, \qquad \text{on }\Omega, \\
	u &= 0, \qquad \text{on } \partial \Omega.
\end{align*}
In this particular problem, we have a domain, on which we are interested in finding a mapping $ u:\Omega \to \R $ whose partial derivatives satisfies a certain equation. Our first goal is to determine if this problem has a solution (existence), and then to see if the problem as a unique solution (uniqueness), and then to see if the solution is stable (i.e. bounded by the given data).

\begin{summary}
	Given a PDE in hand, it is important to see if the problem is well-posed (in the sense  of Hadamard). I.e.
	\begin{itemize}[noitemsep]
		\item a solution exists,
		\item the solution is unique, and
		\item the solution is stable.

	\end{itemize}
\end{summary}

\subsubsection{Different Formulations of the Problem: Strong vs. Weak Formulation}
To find the answers to these questions, we can formulate our problem in different ways, i.e. strong formulation, or a weak formulation.

### 2024-06-09 Findings (Shared Folder)/Chapters/randomNotes.tex
new notes addeD

@@
 \subsubsection{Different Formulations of the Problem: Strong vs. Weak Formulation}
-To find the answers to these questions, we can formulate our problem in different ways, i.e. strong formulation, or a weak formulation.
+To find the answers to these questions, we can formulate our problem in different ways, i.e.
+strong formulation, or a weak formulation.
+In the weak formation, we have more relaxed requirements on the solutions.
+For instance, for a classic solution to satisfy an elliptic Dirichlet boundary value problem, by definition, it needs to have two derivatives while being zero at the boundary (i.e.
+$ u \in C_c^2(\Omega) $).
+However, we can relax this requirement by a kind of axiomatic extension of the set of possible solutions.
+For instance, we know that all $ C_c^2(\Omega) $ functions has a nice interaction with $ C_c^\infty(\Omega) $ functions through the integration by parts formula.
+\[ \int_\Omega \nabla u \nabla v = -\int_\Omega u \nabla v \qquad \forall v \in C_c^\infty (\Omega).  \]
+Thus we can use this property as a definition of derivative, and define the notion of the weak derivative. We can use this property to formulate the problem in weak sense. 
+\[ \int_\Omega \nabla u \ \nabla v = \int_\Omega g v \qquad \forall v \in H_0^1(\Omega). \]
+We can now write the problem in a more abstract form by defining bi-linear forms
+\[ a: H_0^1 \times H_0^1 \to \R, \qquad l:H_0^1 \to \R \] 
+and define 
+\[ a(u,v) = \int_\Omega \nabla u\  \nabla v, \qquad l(v) = \int_\Omega gv. \]
+Thus we can now write the weak formulation as the following abstract equation, i.e. we say $ u $ satisfies the boundary value problem stated above if it satisfies
+\begin{align*}
+	&a(u,v)  = l(v), \qquad \forall v \in H_0^1,\\
+	&u = 0 \quad \text{on}\ \partial \Omega.
+\end{align*}
+
+\subsubsection{well-Posedness}
+Writing the problem in hand in abstract form, helps us to use other existing mathematical infrastructure to understand if a solution exists, if it is unique, and if it is stable.
+For instance, writing the elliptic problem in the form of $ a(u,v) = l(v) $ allows us to use the infrastructure developed in functional analysis to find the answers to our questions.
+For instance, in this particular case, we can use the Lax-Milgram theorem (which is a kind of generalization to the Ritz-representation theorem) to show that under certain conditions (i.e.
+the bi-linear form $ a $ being coercive and bounded, and the linear form $ l $ being bounded.
+) a solution exists, and is unique and stable (i.e.
+bounded by the data).
+Or in the case of the Stokes problem for the non-compressible fluid, we have the  Ladyzhenskaya-Babuška-Brezzi condition (or LBB in short) which is also often called inf-sup condition, is a set of conditions under which the Stokes problem as a unique solution that is stable.
+
+\begin{summary}[Importance of formulating the mathematical problem appropriately]
+	Formulating a mathematical problem in an appropriate way is very crucial.
+Appropriate formulations allows us to use the existing important theorems in a specific field of mathematics to our advantage.
+For instance, by formulating a PDE in weak form and using suitable forms, we can use
+	\begin{itemize}
+		\item Lax-Milgram theorem (generalized form of Ritz-representation theorem) to show the well-posedness of an elliptic problem under certain conditions.
+		\item Ladyzhenskaya-Babuška-Brezzi theorem (or LBB in short) to show the well-posedness of the Stokes problem for non compressible fluid.
+	\end{itemize}
+\end{summary}
+
+\subsubsection{What kind of solutions do we need?}
+The well-posedness of a problem is a hunting license to find the actual solutions.
+Again, we need to revisits our main question in step I.
+What are we looking for?
+Is having a solutions in closed form necessary?
+Can having the solution in a series form (with a high or low convergence rate) enough?
+Can we get the answers to our question by some sort of qualitative analysis?
+Is a numerical solution necessary?
+\textbf{The exact answer to these questions depends on the questions that we want to answer with mathematical modeling.}
+
+\begin{summary}[Solutions to the PDE]
+	After getting the hunting license and being sure about the well-posedness of the problem, we need to revisit our original question to decide what we want to do with the solution?
+The solution can be found in many ways that might be suitable for some cases and and overkill for other cases.
+	\begin{itemize}
+		\item closed form solution
+		\item series solution (with poor or good convergence rate)
+		\item some contracting map 
+		\item local solution (for local behaviour analysis)
+		\item numerical solution
+		\item etc.
+	\end{itemize}
+\end{summary}
+
+\subsubsection{Numerical Solution}
+After deciding on if we need a numerical solution, we can use different approaches, each of which has its own treatment.
+Here I will only talk about the finite element methods.
+In the finite element method, we find a suitable finite dimensional sub-space of the solution space ($ H_0^1 $ in the case of the elliptic Dirichlet problem) and formulate the problem in discrete form.
+For instance, if we can generate a triangulation of a domain $ \Omega $ and define some local basis functions e.g.
+some continouse and piece-wise affine functions that each of them is non-zero only at one node.
+Thus we can express our function in the weak formulation in this form
+\[ u = \sum_{i=1}^{n}U_i \phi_i(x), \qquad g = \sum_{i=1}^{n}G_i \phi_i(x). \]
+By substituting this in the weak formulation, we will get the following system of linear equations
+\[ A U = G, \]
+where the matrix $ A $ (stiffness matrix) and the vector $  G $ are the given data and we want to find the vector $ U $.
+Or in the case of the Stokes problem, we will get the following system of equations.
+\begin{align*}
+	&AU + BP = F, \\
+	&B^T U = 0.
+\end{align*}
+\textbf{Note}: After projecting the problem to a finite dimensional sub-space, it is not necessarily true the the well-posedness of the problem in the infinite dimensional Hilbert space carries over to the finite dimensional sub-space.
+For instance, in the case of the Stokes problem, the well-posedness does not carry over and we need to study the well-posedness of the discrete problem.
+
+\subsubsection{Error Estimate}
+Projecting an infinite dimensional space to a finite dimensional space introduces some errors that we need to be sure we can quantify and understand how does that error scales with the mesh size (i.e.
+the dimension of the finite dimensional sub-space).
+
+\section{What Is the Flux, Really?}
+Here in this section we will go deeper into the notion of flux in particle transport which is the central part mathematical modeling of living systems.
+I will try to connect this notion with mathematical objects with clear definitions rather than providing vague intuition.
+
+Generally speaking, flux is a surface integral of some vector field, and by designing that vector field accurately, we can use the notion of flux in different contexts.
+Here we will design the notion of flux to study particles and their densities assuming they are non-compressible fluids.
+Let $ \Omega \subset \R^3 $.
+Define
+\[ J = \rho v \]
+where
+\[ \rho: \Omega\times \R \to \R, \qquad v:\Omega\times \R\to\R^3, \]
+representing the density and the velocity of the scalar field respectively.
+But what is the velocity of $ \rho $?
+Here is the part where the notion of volume preserving diffeomorphisms become useful.
+Let $ X_0 \in\Omega $ denote the position of a point particle in the domain.
+Then its trajectory over time is determined by the action of the group $ \set{g^t}_{t \in T} $ on $ X_0 $.
+For instance $ X(t) = g^t X_0 $.
+Since we have assumed that the species are in form of non-compressible fluids, then the group $ \set{g^t} $ is a subgroup of the group of all diffeomorphisms that preserves the volume.
+For a fixed $ X_0 \in \Omega $ its orbit is defined to be
+\[ \operatorname{Orbit}(X_0) = \set{g^t X_0 \ |\ t \in T}. \]
+At any time $ t $, $ v(X,t) $ is defined to be
+\[ v(X,t) = \lim_{\Delta t \to 0} \frac{g^{t+\Delta t} X - g^{t} X}{\Delta t} \]
+This means that for $ X_0 \in A $ we can write
+\[ g^{\Delta t} X_0 = X_0 + \vec{v}(X_0) \Delta t + O(\Delta t ^2) \]
+which means that the effect of $ g^{\Delta t} $ on the point $ X_0 $ just shifting $ X_0 $ in the direction $ \vec{v}(X_0) \Delta t  $.
+In the figure below we can see this approximation on a particular orbit.
+\input{images/orbitApproximation.tex}
+Let $ \sigma $ be any surface given in the domain. We claim that the rate of volume passing through this membrane is given by
+\[ \frac{dV}{dt} = \int_\sigma \vec{v}\cdot \hat{n}\ dA. \]
+To see this $ \Delta A $ be the area of a flat square tangent to the membrane at point $ X_0 $ on the surface. See the figure below.
+\input{images/rateOfVolumePass.tex}
+
+\noindent We observe that the point $ X_0 $ moves to $ X_0 + \vec{v}(X_0)\Delta t + \O(\Delta t^2) $.
+Also, observe that the area any point on $ \Delta A $ distance $ \Delta r $ away from $ X_0 $ will be transferred to
+\[ g^{\Delta t} (X_0 + \Delta r) = X_0 + \Delta r + \vec{v}(X_0 + \Delta r)\Delta t + O(\Delta t^2) = (X_0 + \Delta r) + \vec{v}(X_0) \Delta t + O(\Delta t^2 + \Delta t \Delta r).  \]
+Thus the area of the green surface in the figure above will be the same as $ \Delta A $ up to order
+$ O((\Delta t^2 + \Delta t \Delta r)^2) $.
+Note that by the action of $ g^{\Delta t} $, everything in between the yellow and green surfaces will move through the green surface.
+So the volume that will pass through the green surface is
+\[ \Delta V = (\vec{v}(X_0)\Delta t)\cdot \hat{n} \Delta A + O(\Delta t^2) \]
+Given that the density of species is $ u $, then the number species passed through the green membrane after time $ \Delta t $ will be
+\[ \frac{\Delta N}{\Delta t} = \vec{v}(X_0)\cdot \hat{n}\Delta A + O(\Delta t^2 + \Delta A^2) \]
+I.e. we can write
+\[ \frac{d N }{dt} = \int_\sigma \vec{v}\cdot \hat{n}\ dA \]
+
+\begin{summary}
+	In a nutshell, If we define 
+	\[  J = \rho v \]
+	where $ \rho :\Omega \times \R \to \R $ is a function representing the density, and $ v: \Omega \times \R \to \R^3 $ is the tangent vector to the orbit of particles generated by the action of group of diffeomorphisms that preserve the volume element $ \set{g^t}_{t\in T} $, then, for any surface $ \sigma \subset \Omega $, the rate of particles passing through the membrane will be
+	\[ \frac{dN}{dt} = \int_\sigma J\cdot \hat{n}\ dA. \]
+	Then by assuming that no particles will by destroyed to the thin air, and also no particles will born out of blue sky, then we can derive the continuity equation.
+In any volume $ V $ of our domain enclosed by the surface $ \sigma $, we can write
+	\[ \frac{d}{dt} \int_V \rho\ dV = - \int_\sigma J\cdot \hat{n}\ dA = - \int_V \nabla\cdot J dV. \]
+	By dropping the integration signs we will get
+	\[ \frac{d\rho}{dt} + \nabla\cdot J = 0.  \]
+\end{summary}

### 2024-06-09 Findings (Shared Folder)/images/orbitApproximation.tex
new notes addeD

\begin{figure}[h!]
	\centering

	\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

	\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
		%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300

		%Straight Lines [id:da7253651408536048] 
		\draw [color={rgb, 255:red, 160; green, 160; blue, 160 }  ,draw opacity=1 ][line width=0.75]    (136.83,103.17) -- (149.81,85.61) ;
		\draw [shift={(151,84)}, rotate = 126.47] [color={rgb, 255:red, 160; green, 160; blue, 160 }  ,draw opacity=1 ][line width=0.75]    (6.56,-1.97) .. controls (4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and (4.17,0.84) .. (6.56,1.97)   ;
		%Curve Lines [id:da4933662152703622] 
		\draw    (82,133.33) .. controls (151.8,130.12) and (147.11,57.99) .. (164.66,37.97) ;
		\draw [shift={(166.67,36)}, rotate = 140.19] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (7.14,-3.43) -- (0,0) -- (7.14,3.43) -- (4.74,0) -- cycle    ;
		%Curve Lines [id:da8486116401578401] 
		\draw    (164,38.33) .. controls (173,29.67) and (176.33,28.67) .. (188,22.33) ;
		%Shape: Circle [id:dp6250392978342221] 
		\draw  [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ] (135,103.17) .. controls (135,102.15) and (135.82,101.33) .. (136.83,101.33) .. controls (137.85,101.33) and (138.67,102.15) .. (138.67,103.17) .. controls (138.67,104.18) and (137.85,105) .. (136.83,105) .. controls (135.82,105) and (135,104.18) .. (135,103.17) -- cycle ;
		%Shape: Circle [id:dp5689260707692063] 
		\draw  [fill={rgb, 255:red, 74; green, 144; blue, 226 }  ,fill opacity=1 ] (149.33,70.5) .. controls (149.33,69.49) and (150.15,68.67) .. (151.17,68.67) .. controls (152.18,68.67) and (153,69.49) .. (153,70.5) .. controls (153,71.51) and (152.18,72.33) .. (151.17,72.33) .. controls (150.15,72.33) and (149.33,71.51) .. (149.33,70.5) -- cycle ;

		% Text Node
		\draw (134,107.4) node [anchor=north west][inner sep=0.75pt]  [font=\tiny]  {$X_{0}$};
		% Text Node
		\draw (151.67,89.73) node [anchor=north west][inner sep=0.75pt]  [font=\tiny,color={rgb, 255:red, 116; green, 116; blue, 116 }  ,opacity=1 ]  {$\vec{v}( X_{0})$};
		% Text Node
		\draw (156.67,61.4) node [anchor=north west][inner sep=0.75pt]  [font=\tiny,color={rgb, 255:red, 0; green, 61; blue, 133 }  ,opacity=1 ]  {$g^{\Delta t} X_{0} =X_{0} +V( X_{0}) \Delta t+O\left( \Delta t^{2}\right)$};

	\end{tikzpicture}
\end{figure}

### 2024-06-09 Findings (Shared Folder)/images/rateOfVolumePass.tex
new notes addeD

\begin{figure}[h!]
	\centering

	\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

	\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
		%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300

		%Shape: Polygon [id:ds09456144699087154] 
		\draw  [fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ] (178.67,88.67) -- (202,59.33) -- (202,95) -- (178.67,124.33) -- cycle ;
		%Curve Lines [id:da4604489904212945] 
		\draw    (30.67,78.67) .. controls (61,75.33) and (107.67,54.33) .. (112,30.33) ;
		%Curve Lines [id:da6967720169789406] 
		\draw    (34.33,171.67) .. controls (20,138.33) and (47,121.33) .. (30.67,78.67) ;
		%Curve Lines [id:da35266638580627574] 
		\draw    (34.33,171.67) .. controls (54,154.56) and (86.77,156.23) .. (104.47,141.91) .. controls (109.87,137.54) and (113.87,131.68) .. (115.67,123.33) ;
		%Curve Lines [id:da4240338947405422] 
		\draw    (115.67,123.33) .. controls (123.33,89) and (114.67,70) .. (112,30.33) ;
		%Curve Lines [id:da691890187484379] 
		\draw [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ]   (46.67,164) .. controls (32.33,130.67) and (60,119) .. (43.67,76.33) ;
		%Curve Lines [id:da4043207998593765] 
		\draw [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ]   (57.67,160.33) .. controls (43.33,127) and (71,115.33) .. (54.67,72.67) ;
		%Curve Lines [id:da5252238052351152] 
		\draw [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ]   (69.33,155.33) .. controls (55,122) and (82,111.67) .. (65.67,69) ;
		%Curve Lines [id:da31270908120152363] 
		\draw [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ]   (81,152) .. controls (66.67,118.67) and (93,107.67) .. (76.67,65) ;
		%Curve Lines [id:da6303057039241056] 
		\draw [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ]   (93.67,147.67) .. controls (86.33,112) and (104.33,101.67) .. (88,59) ;
		%Curve Lines [id:da8218053838878148] 
		\draw [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ]   (104.47,141.91) .. controls (109.67,109) and (105.33,64) .. (99,51) ;
		%Curve Lines [id:da8308238051916903] 
		\draw [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ]   (113.67,47) .. controls (103.67,66.33) and (55,89.33) .. (34.67,90.67) ;
		%Curve Lines [id:da02053447085552529] 
		\draw [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ]   (115.67,59.67) .. controls (105.67,79) and (57,102) .. (36.67,103.33) ;
		%Curve Lines [id:da5282822241031992] 
		\draw [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ]   (116.67,73) .. controls (106.67,92.33) and (56.33,115.33) .. (36,116.67) ;
		%Curve Lines [id:da23816183015057169] 
		\draw [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ]   (118.33,86) .. controls (113.67,101.67) and (54,129.33) .. (33.67,130.67) ;
		%Curve Lines [id:da569405412784046] 
		\draw [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ]   (118.67,100.67) .. controls (114,116.33) and (51.67,141.67) .. (31.33,143) ;
		%Curve Lines [id:da11295701912713185] 
		\draw [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ]   (117.67,116.33) .. controls (93,147.67) and (48.33,148) .. (30.33,158.67) ;
		%Straight Lines [id:da4575502778776137] 
		\draw    (178.67,88.67) -- (202,59.33) ;
		%Straight Lines [id:da09259804747392031] 
		\draw    (178.67,124.33) -- (178.67,88.67) ;
		%Straight Lines [id:da5208713033396146] 
		\draw  [dash pattern={on 3pt off 3pt}]  (202,95) -- (202,59.33) ;
		%Straight Lines [id:da5610562629835913] 
		\draw  [dash pattern={on 3pt off 3pt}]  (178.67,124.33) -- (202,95) ;
		%Straight Lines [id:da25649545572729315] 
		\draw [color={rgb, 255:red, 160; green, 160; blue, 160 }  ,draw opacity=1 ][line width=0.75]    (190.17,90.5) -- (208.5,82.47) ;
		\draw [shift={(210.33,81.67)}, rotate = 156.35] [color={rgb, 255:red, 160; green, 160; blue, 160 }  ,draw opacity=1 ][line width=0.75]    (6.56,-1.97) .. controls (4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and (4.17,0.84) .. (6.56,1.97)   ;
		%Shape: Circle [id:dp4347516737268753] 
		\draw  [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ] (188.33,90.5) .. controls (188.33,89.49) and (189.15,88.67) .. (190.17,88.67) .. controls (191.18,88.67) and (192,89.49) .. (192,90.5) .. controls (192,91.51) and (191.18,92.33) .. (190.17,92.33) .. controls (189.15,92.33) and (188.33,91.51) .. (188.33,90.5) -- cycle ;
		%Straight Lines [id:da9970436041671149] 
		\draw    (213.67,64) -- (240.33,30.67) ;
		%Straight Lines [id:da342511901706108] 
		\draw    (213.67,107) -- (213.67,64) ;
		%Straight Lines [id:da5902552900557785] 
		\draw    (240,76.33) -- (240.33,30.67) ;
		%Straight Lines [id:da8931061197488246] 
		\draw    (213.67,107) -- (240,76.33) ;
		%Straight Lines [id:da6468976920553455] 
		\draw    (178.67,124.33) -- (213.67,107) ;
		%Straight Lines [id:da7362591464280432] 
		\draw    (178.67,88.67) -- (213.67,64) ;
		%Straight Lines [id:da5565256227844282] 
		\draw    (202,59.33) -- (240.33,30.67) ;
		%Straight Lines [id:da34026137798256473] 
		\draw  [dash pattern={on 3pt off 3pt}]  (202,95) -- (240,76.33) ;
		%Shape: Polygon [id:ds9396511578559401] 
		\draw  [fill={rgb, 255:red, 80; green, 227; blue, 194 }  ,fill opacity=1 ] (240.33,30.67) -- (240,76.33) -- (235.75,81.28) -- (213.67,107) -- (213.67,64) -- cycle ;
		%Shape: Circle [id:dp5366643183971873] 
		\draw  [fill={rgb, 255:red, 74; green, 144; blue, 226 }  ,fill opacity=1 ] (226.67,69.5) .. controls (226.67,68.49) and (227.49,67.67) .. (228.5,67.67) .. controls (229.51,67.67) and (230.33,68.49) .. (230.33,69.5) .. controls (230.33,70.51) and (229.51,71.33) .. (228.5,71.33) .. controls (227.49,71.33) and (226.67,70.51) .. (226.67,69.5) -- cycle ;
		%Shape: Polygon [id:ds3478569100900799] 
		\draw  [fill={rgb, 255:red, 224; green, 211; blue, 45 }  ,fill opacity=1 ] (95,93) -- (94,107.67) -- (79,114.67) -- (81.67,101.67) -- cycle ;

		% Text Node
		\draw (183,96.73) node [anchor=north west][inner sep=0.75pt]  [font=\tiny]  {$X_{0}$};
		% Text Node
		\draw (227.33,53.73) node [anchor=north west][inner sep=0.75pt]  [font=\tiny,color={rgb, 255:red, 0; green, 61; blue, 133 }  ,opacity=1 ]  {$g^{\Delta t} X_{0} =X_{0} +\vec{v}( X_{0}) \Delta t+O\left( \Delta t^{2}\right)$};
		% Text Node
		\draw (161.33,122.07) node [anchor=north west][inner sep=0.75pt]  [font=\tiny,color={rgb, 255:red, 136; green, 124; blue, 22 }  ,opacity=1 ]  {$\Delta A$};
		% Text Node
		\draw (222.67,102.07) node [anchor=north west][inner sep=0.75pt]  [font=\tiny]  {$\textcolor[rgb]{0.08,0.55,0.12}{g^{\Delta t} \Delta A} =\Delta A+O\left( \| \ ( \Delta r,\Delta t) \| ^{3}\right)$};
		% Text Node
		\draw (166,144.4) node [anchor=north west][inner sep=0.75pt]  [font=\tiny,color={rgb, 255:red, 136; green, 124; blue, 22 }  ,opacity=1 ]  {$\Delta V=(\vec{v}( X_{0}) \cdotp \hat{n}) \Delta t\Delta A$};

	\end{tikzpicture}
\end{figure}

### 2024-06-10 Findings (Shared Folder)/Chapters/randomNotes.tex
new notes added

@@
 \begin{summary}
+	\label{summary:groupActionOfDiffeomorphisms}
 	In a nutshell, If we define 
@@
 \end{summary}
+
+\begin{observation}
+	Reading the summary box above, one might conclude that the flux for diffusion might be something like $ J = - \rho \nabla\rho $ as the particles are moving opposite to their gradient.
+But this is wrong!
+In the case of diffusion, the velocity field is
+	\[ v = -\frac{1}{\rho} \nabla \rho \]
+	so the flux (Fick's law) will be of the form
+	\[  J =  -\nabla\rho. \]
+	At first, it was as very bizarre observation for me. I still do not understand how I can derive this velocity profile (i.e. flow velocity or simply flow).
+\end{observation}
+
+\section{Some cool PDE simulations}
+Living in this level of abstraction, we can now do some stuff with PDEs that might look magic at the first glace. Here, I will show case some of these simulations.
+
+\subsection{Diffusion and Advection}
+
+\begin{quote}
+	{\color{red} Be aware of the context.} Note that in this section, when we talk about the vector field $ F $, we are actually talking about the vector field that determines the advection term of the flux.
+I.e.
+we assume $ F $ is
+	\[ J_\text{advection} = \rho F\]
+\end{quote}
+To have diffusion and advection at the same time, we need to design the flow in a way that can produce this.
+The first term of will be for the diffusion, where according to the observation box above will be of the form
+\[  J = - \nabla\rho + \underbrace{\cdots}_{\text{advection term}} \]
+To design the advection term, we need to observe that the particles should move along side some vector field. So it will be of the form
+\[ \boxed{J_\text{advection} = \rho F(X)}  \]
+where $ F: \Omega\to \R^3 $ is the desired vector field. So the advection-diffusion PDE will be 
+\[ \frac{\partial \rho}{\partial t} = \nabla\cdot (-\nabla\rho + \rho F). \]
+Now we can come up with all sort of creative vector fields.
+\subsubsection{Simple velocity field}
+For instance, for the vector field $ F(x,y)=(-1,0) $, we will have
+\[ \frac{\partial \rho}{\partial t} = \nabla^2 \rho - \frac{\partial \rho}{\partial x}.  \]
+Then the solution of this model will be like the following figure.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.7\linewidth]{images/simpleAdvectionDiffusion1}
+	\caption{Advection diffusion with the vector field $ F(x,y,z) = (-1,0,0). $ The designed PDE forces the particles to move in this velocity field (or velocity profile).}
+	\label{fig:simpleadvectiondiffusion1}
+\end{figure}
+
+\subsubsection{Particles Moving Down the Hill}
+Here, we design a vector field that resembles particles moving down a potential hill to get to their minimum potential state.
+To design such a vector field, we simply first define an appropriate potential function and define the vector field to be its gradient.
+The potential that we will using for this purpose is
+\[ f(x,y) = 1/2(x^2 + y^2). \]
+Then the corresponding vector field will be
+\[ F(x,y) = -\nabla f(x,y) = (x,y). \]
+So the desired PDE will be
+\[ \frac{\partial \rho}{\partial t} = \nabla^2 \rho + x \frac{\partial \rho}{\partial x} + y \frac{\partial \rho}{\partial y} + \rho. \] 
+Simulating this PDE we will get
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.7\linewidth]{images/movingToCenter.png}
+	\caption{Advection diffusion with the vector field $ F(x,y) = (x,y)$ which is the gradient of potential function $ f(x,y) = 1/2(x^2 + y^2) $.}
+	\label{fig:potentialAsGradient}
+\end{figure}
+
+\subsubsection{Moving Around the Origin}
+What if at each point, we rotate the vector vector of the previous example by 90 degrees so that the particles will move around the origin.
+The vector field in previous example was $ F(x,y) = (x,y) $.
+To do the rotation we need ot multiply that in the rotation matrix
+\[ F_\text{new}(x,y)=  \matt{0}{-1}{1}{0}F(x,y) = \vectt{-y}{x}.  \]
+Thus the desired PDE will be
+\[ \frac{\partial \rho}{\partial t} = \nabla^2 \rho - y\frac{\partial \rho}{\partial x} + x \frac{\partial \rho}{\partial y}. \]
+By simulating this PDE we will get
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.7\linewidth]{images/movingAround.png}
+	\caption{Advection diffusion with the vector field $ F(x,y,z) = (-y,x)$.
+This vector field is derived by rotating the vectors in the previous example by 90 degrees.
+Surprisingly, there are no potentials that $ F $ ce be its gradient!
+In other words, we can not derive $ F $ from any potential.}
+	\label{fig:simpleadvectiondiffusion1}
+\end{figure}
+
+\FloatBarrier
+
+\subsubsection{Spiraling in to the Origin}
+We can start with the vector field $ F(x,y) = (x,y) $ and this time, rather time rotating it with a rotation matrix, we can rotate that using a matrix corresponding to a complex number $ z = \mu + i \omega $, i.e.
+\[ F_\text{new}(x,y) = \matt{\mu}{-\omega}{\omega}{\mu}F(x,y). \]
+We can take this complex number to be $ z = -1 + i $. Thus our new vector field will be
+\[ F_\text{new}(x,y) = (-x+y, -x-y).  \]
+So our PDE will read
+\[ \frac{\partial \rho}{\partial t} = \nabla^2 \rho + (-x+y)\frac{\partial \rho}{\partial x} + (-x-y)\frac{\partial \rho}{\partial y} - 2\rho. \]
+Simulating this PDE we will get
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.5\linewidth]{images/spiralAdvectionDiffusion.png}
+	\caption{Advection diffusion with the vector field $ F_{\text{advection}}(x,y,z) = (-x+y, -x-y)$.}
+	\label{fig:spiralAdvectionDiffusion}
+\end{figure}
+\FloatBarrier
+
+Now the interesting question is why do we see this?
+To observe this, remember that the velocity profile is the same as the vector field that we specify for the model (in this case $ F_\text{new}(x,y) $).
+On the other hand, we know that the velocity profile is in fact the right hand side of the ODE for flow
+\[ \frac{d \phi}{d t} = F_\text{new}(x,y). \]
+Note that $ \phi $ is flow, or following the terminology of \autoref{summary:groupActionOfDiffeomorphisms}, it is one element of the group diffeomorphisms that preserve the volume element (in the that terminology we denote it by $ g^t $).
+For the simulation above, ODE for $ \phi:\R^2 \to \R^2 $ is
+\[ \vectt{\dot{\phi_1}}{\dot{\phi_2}} = \matt{-1}{1}{-1}{-1}\vectt{x}{y}. \]
+The phase plane of flow will be
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.2\linewidth]{images/flowLinePlot.png}
+	\caption{The phase plane of flow $ \phi_\text{advection} = (\phi_1,\phi_2) $}
+	\label{fig:SpiralAdvectionDiffusionFlowPhasePortrait}
+\end{figure}
+\FloatBarrier
+
+The following summary box summarizes our observation.
+
+\begin{summary}[Advection-Diffusion PDE]
+	In an advection diffusion PDE for $ \rho $ the flux will have two terms, 
+	\[ J = \underbrace{-D\nabla \rho}_{J_\text{diffusion} = \rho v_\text{diffusion}}+ \underbrace{\rho F}_{J_\text{advection} = \rho v_\text{advection}} \]
+	where $ F $ is a vector field, which is the velocity profile of advection (i.e.
+$ v_\text{advection} = F $).
+Thus the flow (or in the terminology of \autoref{summary:groupActionOfDiffeomorphisms} the diffeomorphism preserving the volume element) will be given with the following ODE
+	\[ \frac{d \phi_\text{advection}}{dt} = v_\text{advection}. \]
+	In the cases where we have small diffusion, the solution of the advection-diffusion PDE will look like the phase plane diagram of $ \phi_\text{advection} $.
+Compare \autoref{fig:spiralAdvectionDiffusion} with  \autoref{fig:SpiralAdvectionDiffusionFlowPhasePortrait} to see this.
+\end{summary}
+
+\begin{observation}
+	Given what we discussed above, I have the following idea to study the phase plane of an ODE system in a wise way! Consider the following ODE system
+	\[ \dot{\Phi} = F(\Phi). \]
+	In the corresponding phase space diagram, we usually plot the vector field $ F(X) $ and roughly draw some important orbits (see \autoref{fig:SpiralAdvectionDiffusionFlowPhasePortrait}).
+However, using the machinery we discussed in the summary box above, we can also evaluate what will happen for a given volume of initial positions (i.e.
+a volume of phase space).
+We can design a PDE whose flow is given by $ \Phi $.
+Thus this PDE will be
+	\[ \frac{\partial \rho}{\partial t} = -\nabla\cdot(\rho F(\Phi)), \]
+	where $ \rho $ represents the density of states in the phase space. Note that the equation above is nothing more that just inserting an appropriate choice of flux into the continuity equation.
+\end{observation}
+
+\begin{observation}[A proof for Liouville's theorem (Hamiltonian)]
+	After doing some research, I found out that the approach I have had in the observation above can be used to prove the Liouville's theorem for Hamiltonian.
+This was an independent discovery and proof of this theorem!
+
+	\begin{proof}
+		consider the following Hamiltonian system
+		\[ \Phi = \vectt{\dot{q}}{\dot{p}} = \vectt{\partial H/ \partial p}{- \partial H / \partial q} = F(\Phi). \]
+		Let $ \rho $ be the density of states in the phase plane, since all of these points are following the flow profile, thus the corresponding flux will be
+		\[  J = \rho \Phi.  \]
+		Inserting this into the continuity equation we will get
+		\[ \frac{\partial \rho}{\partial t} = -\nabla\cdot (\rho \Phi) = - \nabla \rho \cdot \Phi - \rho\nabla\cdot\Phi . \]
+		For the first term in the right hand side we have
+		\[ \nabla\rho\cdot \phi = \frac{\partial \rho}{\partial q}\cdot{q} + \frac{\partial \rho}{\partial p} \cdot{p}. \]
+		And for the second term we can write
+		\[ \rho\nabla\cdot\Phi = \rho(\frac{\partial^2 H}{\partial q\partial p} - \frac{\partial^2 H}{\partial p \partial q}) = 0. \]
+		Thus we can write
+		\[ \frac{\partial\rho}{\partial t} + (\frac{\partial \rho}{\partial q}\cdot{q} + \frac{\partial \rho}{\partial p} \cdot{p}) = 0.  \]
+		This completes the proof.
+	\end{proof}
+\end{observation}
+
+\begin{observation}
+	The time evolution of the density of a die in an \emph{incompressible fluid} due to \emph{advection}, i.e. with the flux given by
+	\[  J=\rho v\]
+	 is analogous to the time evolution of the density of states in the phase plane of its flow $ \Phi $ (determined by the equation $ \dot{\Phi}=v $).
+\end{observation}
+
+In the observation box above, we have only talked about the rationale behind the PDEs that contains only the advection term.
+But what about the diffusion term?
+What are the ODE systems where the density of states in its phase plane will have a diffusion like behaviour?
+The following important observation box answers this important question.
+
+\begin{observation}[Diffusion of the density of states in phase space]
+	The statement of the question is as follows.
+	\begin{quote}
+		I was studying the PDEs governing the time evolution of the concentration of some die in an incompressible fluid.
+I noticed this general framework behind that which can be applied to the time evolution of the density of states in the phase plane of an ODE system.
+Once we know the governing ODE for the state of a system, lets say
+		\[ \dot{X}  = F(X) \] 
+		Then we can draw the phase plane with some important orbits in it.
+However, in order to find out what is the time evolution of the density of states, then we can easily find a corresponding PDE given by first defining $  J = \rho F$ and then writing
+		\[ \rho_t = -\nabla\cdot(\rho F) = -\nabla\cdot(J). \]
+		This gives the time evolution of a given density of states.
+In order to have a diffusion like behaviour in the phase plane, we should have $  J = \rho F = -\nabla \rho$.
+This means that the corresponding RHS for ODE should be $ F = -1/\rho \nabla \rho $ which does not make sense for me, as when writing the ODE for $ X $ the right hand side depends on the density of the states which does not makes sense in an ODE.
+So my question is that in what situation the density of states in and ODE system will have a diffusion like behaviour?
+	\end{quote}
+
+	The answer to this question has very deep roots in the stochastic vs.
+deterministic systems.
+The fact is that in \emph{deterministic} ODEs, we will \emph{never} observe a \emph{diffusion like behaviour} in the density of states of phase plane.
+The diffusion like behaviour of the density of states happens only in the case of Stochastic Differential Equations (SDE).
+As in the case of deterministic ODE systems, defining a suitable flux and inserting it into the continuity equation gives us the PDE for the time evolution of the density of states, doing a similar process gives us the Fokker-Plank equation that describes the time evolution of the density of states (which is called probability density function in this context) of the corresponding SDE.
+Consider the following SDE
+	\[ dX_t = \mu(X_t,t)dt + \sigma(X_t,t)dW_t, \]
+	where $ \mu(X_t,t) $ is the drift and $ D(X_t,t) = \sigma^2(X_t,t)/2 $ is the diffusion coefficient.
+The corresponding Fokker-Plank equation for the probability density function $ p(x,t) $ of the random variable $ X_t $ (analogous to the notion of the density of states in what we had discussed above) is given by
+	\[ \frac{\partial}{\partial t} p(x,t) = -\frac{\partial}{\partial x}[\mu(x,t)p(x,t) - \frac{\partial}{\partial x}(D(x,t)p)]. \]
+	Similarly, for the higher dimensions, the SDE will be of the form
+	\[ d\mathbf{X_t} = \boldsymbol{\mu}(\mathbf{X_t},t)dt + \boldsymbol{\sigma}(\mathbf{X_t},t)d\mathbf{W_t}, \]
+	where $ \mathbf{X_t} $ and $ \boldsymbol{\mu} $ are $ N $ dimensional vectors, and $ \boldsymbol{\sigma} $ is $ N\times M $ matrix, and $ \mathbf{W_t} $ is an $ M $ dimensional Wiener process.
+Then the corresponding Fokker-Plank equation will be given as
+	\[ \frac{\partial p}{\partial t} = -\nabla\cdot(p\boldsymbol{\mu} + \boldsymbol{\nabla}\cdot(p\mathbf{D})) \]
+	where the bold letter $ \boldsymbol{\nabla}\cdot$  applied on a matrix outputs a vector where each row is the divergence of the rows of the input matrix, and the matrix $ \mathbf{D} = \frac{1}{2}\boldsymbol{\sigma}\boldsymbol{\sigma^T} $.
+See examples below for more details.
+\end{observation}
+
+\begin{example}[1D Diffusion Equation]
+	Consider the following SDE
+	\[ dX_t = \sqrt{D}dW_t. \]
+	The PDE governing the density of states (or the probability density function), i.e. the Fokker-Plank equation is given by
+	\[ \frac{\partial p}{\partial t} = -\frac{\partial}{\partial x} (-D \frac{\partial p}{\partial x}) = D\frac{\partial ^2 p}{\partial x^2}, \]
+	which gives rise to the 1D diffusion equation.
+\end{example}
+
+\begin{example}[2D Diffusion Equation]
+	Consider the following system of SDEs
+	\[ \vectt{dX_t}{dY_t} = \matt{1}{0}{0}{1}\vectt{dW^{[1]}_t}{dW^{[2]}_t}. \]
+	The corresponding Fokker-Plank equation will be
+	\[ \frac{\partial p}{\partial x} = \nabla\cdot(\boldsymbol{\nabla}\cdot\matt{p}{0}{0}{p}) 
+	= \nabla\cdot(\vectt{p_x}{p_y}) = p_{xx} + p_{yy},
+	 \]
+	Which is the 2D diffusion equation.
+\end{example}
+
+\begin{example}[Advection Diffusion Equation]
+	For instance, if we have
+	\[ \vectt{dX_t}{dY_t} = \mathbf{F}(X_t)dt +  \matt{1}{0}{0}{1}\vectt{dW^{[1]}_t}{dW^{[2]}_t}. \]
+	Then the corresponding Fokker-Plank equation will be
+	\[ \frac{\partial p}{\partial t} = -\nabla\cdot( -\boldsymbol{\nabla}\cdot(\matt{p}{0}{0}{p}) + p\mathbf{F}) = -\nabla\cdot(-\nabla p + p\mathbf{F}). \]
+	which is nothing but the familiar advection-diffusion system.
+Note that first term inside the parenthesis is the Fick's first law and the second term is taking care of the fact that the particles are advecting according to the vector field $ \mathbf{F} $.
+\end{example}
+
+\begin{example}[Boltzmann Distribution at the Thermodynamic Equilibrium]
+	Consider the following SDE that is the overdamped Langevin equation
+	\[ dX_t = -\frac{1}{k_BT}(\nabla U) dt + dW_t. \]
+	Observe that similar to the example corresponding to \autoref{fig:potentialAsGradient} we defined the vector field for advection (drift) term as the gradient of a potential function.
+Thus similar to the example mentioned above, the particles will following trajectories down the potential well.
+The corresponding Fokker-Plank equation will be
+	\[ \frac{\partial p}{\partial t} = -\frac{\partial}{\partial x}(-\frac{\partial p}{\partial x} - p\frac{1}{k_BT}\nabla U ), \]
+	where we have kept writing the derivative of $ U $ as $ \nabla U $ to emphasis it is the gradient of a potential, but in fact we have $ \nabla U = \frac{\partial U}{\partial x} $.
+The steady state solution for the PDE above is
+	\[ \frac{d p}{d x} = \frac{-p}{k_BT}\frac{d U}{d x}. \]
+	The solution of the ODE above is given by
+	\[\boxed{ p(x) = A e^{-\frac{U(x)}{k_BT}}}. \]
+\end{example}
+
+\begin{example}[Ornstein-Uhlenbeck process]
+	Consider the following SDE
+	\[ dX_t = -X_tdt + dW_t. \]
+	Then the corresponding Fokker-Plank equation will be
+	\[ \frac{\partial p}{\partial t} = -\frac{\partial}{\partial x}(-px - p_x). \]
+	The steady state solution of this PDE will be given as
+	\[ \frac{d p}{dx}  = - p x + c.\]
+	The solution of this differential equation is given as
+	\[ \boxed{p(x) = A e^{-x^2/2}}. \]
+	This model is used to model the financial markets.
+\end{example}

### 2024-06-13 Findings (Shared Folder)/Chapters/modelAttempts.tex
new section for feedbacks
new notes added for the angiogenesis review

@@
 
-\section{Simple ODE model (First Iteration)}
+\section{Simple Model from Scratch}
+\subsection{Simple ODE Model}
 Here, we develop a model that keeps track of the following variables
@@
 
-\section{Simple Spatially Distributed 1D System}
+\subsection{Simple Spatially Distributed 1D System}
 
@@
 
-\subsection{Flux of Tip Cells}
+\subsubsection{Flux of Tip Cells}
 A key concept in extending the model to incorporate spatial dynamics is the introduction of the flux of tip cells, denoted as \(J\).
@@
 
-\subsection{Spatial Model Development}
+\subsubsection{Spatial Model Development}
 We now proceed to develop the spatial model by modifying the equations from the original ODE model to incorporate spatial derivatives, reflecting the spatial dynamics of vessel formation, tip cell movement, and drug diffusion.
@@
 \begin{equation}
-	\frac{\partial \rho}{\partial t} = vn - \delta_v \rho,
+	\boxed{\frac{\partial \rho}{\partial t} = vn - \delta_v \rho},
 \end{equation}
@@
 \begin{equation}
-	\frac{\partial n}{\partial t} + \frac{\partial (nv)}{\partial x} = (\lambda_s - \delta_t) n + \lambda_b \rho - \kappa n \rho,
+	\boxed{\frac{\partial n}{\partial t} + \frac{\partial (nv)}{\partial x} = (\lambda_s - \delta_t) n + \lambda_b \rho - \kappa n \rho},
 \end{equation}
@@
 
+\section{Anderson Chaplain Model of Angiogenesis}
+
+\subsection{Biology Facts and Basics}
+The sequences of angiogenesis is assumed to be
+\begin{enumerate}[(i),noitemsep]
+	\item Tumor angiogenic factors (TAF) are secreted by the tumors under low oxygen stress. 
+	\item Emergence of a chemical gradient of TAF between the tumor and the pre-existing vasculature.
+	\item Endothelial cells lining the interior of the existing vasculature respond to this chemical clue by
+	\begin{itemize}[noitemsep]
+		\item Secreting digestive enzymes to loosen the basal lamina.
+		\item Migrate from the disrupted membrane, up in the TAF gradient towards the tumor (i.e. chemotaxis). Look at the citations in the corresponding section)
+		\item Small, finger-like capillary sprouts are formed by accumulation of endothelial cells which are recruited from the parent vessel.
+		\item The sprout grow in length and continue to move toward the tumor directed by the motion of the leading endothelial cells at the sprout tip.
+	\end{itemize} 
+	\item Further sprout extension occurs when some of the endothelial cells of the sprout wall begin to proliferate.
+Cell division is largely confined to a region just behind the cluster of mitotically inactive endothelial cells that constitute the sprout time.
+Note that \cite{Sholley1984} demonstrated that in the absence of endothelial cell proliferation, a restricted capillary network, which stops after a few days and never reaches the tumor, is formed.
+Thus, unless the endothelial cells undergo mitosis, the capillary sprouts cannot complete vascularization completely.
+	\item This process of sprout tip migration and proliferation of sprout-wall cells forms solid strands of endothelial cells amongst the extracellular matrix.
+	\item The tip cells find their way through the extracellular matrix that consist of itersitial tissue, collagen fiber and fibronectin as well as other components.
+	\item Interactions between the endothelial cells and the extracellular matrix, in particular fibronectin.
+	\begin{itemize}
+		\item Cultured endothelial cells are known to synthesize and secrete cellular fibronectin (i.e.
+the insoluble form) \cite{Birdwell1978,Birdwell1980,1978,Macarak1978,Nerlich1991}.
+Note that the insoluble form is first secreted in the soluble form and then assembles into insoluble matrix.
+		\item The expression of secreted fibronectin in cultured endothelial cells closely reflects the distribution of pre-existing fibronectin  observed in matrices \emph{in vivo} \cite{Vlodavsky1979,Hynes1989-je}
+		\item The fibronectin secreted by the endothelial cells do not diffuse and bound the to extracellular matrix \cite{Birdwell1980,Hynes1989-je}.
+		\item Fibronectin secreted by the endothelial cells is a major ligand between cells and matrix material in many situations.
+		\item Endothelial cells use fibronectin for attachment to the matrix via integrins, a family of cell surface receptors \cite{Johansson1987,Hynes1989-je,Alberts2002a}.
+	\end{itemize}
+	\item Haptotaxis of the endothelial cells: migration of the endothelial cells up the gradient of adhesiveness \cite{Everitt1996,McCarthy1984,Carter1965}.
+	\item Thus in addition to the chemotaxis of endothelial cells following the chemical gradient of TAF, there is a complementary haptotactic response to the fibronectin present within the extracellular matrix \cite{Bowersox1982}.
+	\item Modes of the spatial growth:
+	\begin{itemize}
+		\item Initially, the sprouts arising from the parent vessel are essentially parallel to each other. 
+		\item Tend to incline towards each other when the finger-like capillary sprouts have reached a certain distance from the parent vessel \cite{Paweletz1989} leading to numerous tip-to-tip and tip-to-sprout fusions known as anastomoses.
+This anastomoses leads to a network of poorly perfused loops and arcades.
+		\item Form this process of anastomoses, the first signs of circulation can be recognized and from the primary loops, new buds and sprouts emerge repeating the angiogenic sequence of events.
+		\item The process of emerging new capillary sprouts from the existing sprout tips is often called \emph{sprout branching}.
+		\item As the sprouts approach the tumor, their branching dramatically increases until the tumor is eventually penetrated.
+	\end{itemize}
+	\item The repeated steps of angiogenesis: endothelial cell migration, sprout extension, cell proliferation and loop formation.
+\end{enumerate}
+
+\subsection{A Review of previous mathematical models}
+\begin{itemize}
+	\item Continuum and deterministic models for one dimension developed by \cite{Liotta1977,Balding1985,Chaplain1993,Byrne1995a,Orme1996}.
+These models were able to capture some features like average sprout density and network expansion rates.
+But was not able to capture more detailed information concerning the actual structure and morphology of the capillary network.
+	\item More realistic 2D continuum models by \cite{Chaplain1995,Orme1997}: These models allowed a more detailed qualitative comparison with \emph{en vivo} observations about the spatial morphology of the networks.
+However, even these models were not able to capture the events like repeated sprout branching and hence overall dendrictic structure of the network.
+	\item {\color{orange} Important: }More general continuum branching models \cite{Meinhardt1976,Meinhardt1982,Ermentrout1993}.
+	\item Discrete probabilistic framework by \cite{Stokes1991} using stochastic differential equations.
+This model incorporated chemotaxis but not the Haptotaxis.
+Also, was not able to capture the fact that more branched appear as the network gets closer to the tumor.
+This effect is called the \textbf{brush border} effect\cite{Gimbrone1974,Ausprunk1977,Zawicki1981,Muthukkaruppan1982,Sholley1984}.
+\end{itemize}
+
+\textbf{Some notes about the fibronectin:} Fibronectin is a high molecular weight (about 500 kD) Glycoprotein (i.e.
+protein  + an Oligosaccharide) that binds to the membrane-spanning receptor protein called Integrin.
+On the other hand, they also bind to other extracellular matrix proteins like collage, fibrin, etc.
+
+Fibronectin in blood plasma, is in soluble form (made in Liver), however, in the extracellular matrix, the Fibronectin is first secreted in soluble form (mostly by the fibroblast cells) and then assembles into an insoluble matrix in a complex cell mediated process.
+In the terminology above, the soluble fibronectin is called \emph{plasma fibronectin} and the insoluble form is called \emph{cellular fibronectin}.
+
+\subsection{Details of the model}
+Here in this section I will follow the model described in \cite{Anderson1998}.
+This model is based on the experimental data presented in \cite{Gimbrone1974} and \cite{Muthukkaruppan1982}.
+The variables of the model are as the following.
+
+\begin{itemize}
+	\item $ n = n(X,t): \Omega \times \R \to \R $: the endothelial-cell density (per unit area).
+	\item $ c = c(X,t): \Omega \times \R \to \R $: the tumor angiogenic factor (TAF) concentration (nmol per unit area).
+	\item $ f = f(X,t): \Omega \times R \to \R $: the fibronectin concentration (nmol per unit area).
+\end{itemize}
+
+\subsection{Some Thoughts}
+\subsubsection{Discrete vs. continuous model}
+For me, the distinction between discrete model and the continuous model discussed above is just the difference between solving a SDE (which will give us some trajectories) vs.
+solving the corresponding Fokker-Plank equation.
+However observe that, in SDE trajectories, we can measure some quantities like the number of branches, number of loops, the tortuosity, distribution of the length segments, etc.
+On the other hand, the corresponding Fokker-Plank equation (the advection-diffusion PDE) gives us the time evolution of the tip cell density.
+So the following is my question
+
+\begin{openQuestion}
+	Given an SDE, is it possible to write down a differential equation (like the Fokker-Plank equation) that can capture a measurement of interest?
+Like the time evolution of the average number of collisions between $ n $ trajectories of the SDE.
+\end{openQuestion}
+Here are some possible steps to find the answer to the question above.
+\begin{observation}[Steps to solve the question above]
+	The overall theme of solving the question above is to generalize the Fokker-Plank equation of an SDE such that we can write down a similar PDE for any measurement of interest. 
+	\begin{enumerate}[(I)]
+		\item First, try to think about the Fokker-Plank equation in this framework.
+I.e.
+try to observe what kind of measurement is the variable $ \rho $ whose time evolution is captured by the Fokker-Plank equation.
+		\item Try to abstract the notion of $ \rho $. I.e. What are the properties of $ \rho $ that help in developing the Fokker-Plank equation?
+	\end{enumerate}
+\end{observation}
+
+\subsection{Important citations of this paper}
+\begin{itemize}[noitemsep]
+	\item biology of the angiogenesis
+	\begin{itemize}[noitemsep]
+		\item Embryo angiogenesis \cite{Graham1992}.
+		\item Angiogenesis in tissue repair \cite{Arnold1991,Pettet1996}.
+		\item Uncontrolled angiogenesis in tumors, arthritis, abnormal neovascularization of the eye, duodenal ulcers, and following myocardial infarction \cite{Folkman1985,Folkman1995,Folkman1987}.
+		\item Pathological examples of angiogenesis \cite{Muthukkaruppan1982,Ribatti2008}.
+	\end{itemize}
+	\item Angiogenic factors:
+	\begin{itemize}
+		\item Several angiogenic factors like vascular endothelial growth factor (VEGF), acidic and basic fibroblast growth factor (aFGF, bFGF), angiogenin, etc has been isolated \cite{Folkman1987,Relf1997}.
+		\item Endothelial cell receptors for these angiogenic factors \cite{Millauer1993,Hatva1995,Mandriota1995,Fong1995,Hewett1996,Patterson1996,Kappel1999,0deab5d12daa4d32b398e5c5f589f651,Hanahan1997}
+		\item Effect of disrupting the angiogenic factors on the final structure of the capillary network \cite{Dumont1994,Fong1995,Sato1995,Hanahan1997}.
+	\end{itemize}
+	\item Initial response of the endothelial cells to the angiogenic factors is a chemotatic one \cite{Sholley1984,Terranova1985,Paweletz1989,Stokes1990}.
+	\item New finger-like capillary sprout formation \cite{Cliff1963,Warren1966,Ausprunk1977,Sholley1984}
+	\item New tip cells finding their way through the extracellular matrix \cite{Liotta1983,Paweletz1989}.
+\end{itemize}
+
 \newpage
-
-\section{Some Ideas to Try}
+\section{Idea Bin!}
 This section might have very simple, basic and sometimes silly ideas that came into my mind during developing some models and I thought they might worth trying

### 2024-06-14 Findings (Shared Folder)/Chapters/modelAttempts.tex
new notes added for angiogenesis

@@
 
+\newpage
+
 \section{Anderson Chaplain Model of Angiogenesis}
 
-\subsection{Biology Facts and Basics}
+\subsection{Biological Facts and Basics}
+\label{subsection:AndersonChaplainBiologyFacts}
 The sequences of angiogenesis is assumed to be
@@
 		\item Migrate from the disrupted membrane, up in the TAF gradient towards the tumor (i.e. chemotaxis). Look at the citations in the corresponding section)
-		\item Small, finger-like capillary sprouts are formed by accumulation of endothelial cells which are recruited from the parent vessel.
+		\item \label{item:InitialDistributionOfTipCells} The endothelial cells form into a few cell clusters \cite{Muthukkaruppan1982,Orme1996} which eventually become sprouts.
+Small, finger-like capillary sprouts are formed by accumulation of endothelial cells which are recruited from the parent vessel.
 		\item The sprout grow in length and continue to move toward the tumor directed by the motion of the leading endothelial cells at the sprout tip.
@@
 Thus, unless the endothelial cells undergo mitosis, the capillary sprouts cannot complete vascularization completely.
+\label{item:tipCeillsAreDividing}
 	\item This process of sprout tip migration and proliferation of sprout-wall cells forms solid strands of endothelial cells amongst the extracellular matrix.
@@
 	\begin{itemize}
+		\item \label{item:initialPhaseOfFibroNectin} \textbf{Initial distribution of fibronectin:} At the very early stages of angiogenesis, when the endothelial cells lining the parent vessels, are stimulated by TAF, they secrete enzymes to digest through the basal lamina of the parent vessel.
+This initial damage results in an increased vessel permeability \cite{Clark1981-pb}.
+This allows the plasma fibronectin from the blood to leak from the parent vessel and diffuse through the domain \cite{Hynes1989-je}.
+The diffusion coefficient of fibronectin has been measured to be $ 2\times10^{-7} \text{cm}^2\text{s}^{-1} $ \cite{Williams1982,Rocco1987}.
+This plasma fibronectin eventually binds to the extracellular matrix \cite{Oh1981,Deno1983,Clark1983}, thus prior to the start of angiogenesis we will have a gradient of fibronectin that is high near the parent vessel and gets lower away from the parent vessel.
+This has been experimentally observed by \cite{Paku1991} and \cite{Clark1982,Clark1981-pb,Clark1983}.
+It has also been experimentally observed that a higher concentration of laminin (another matrix macromolecule with similar adhesive properties to fibronectin) are initially found around the parent vessel \cite{Hynes1989-je,Paku1991}
 		\item Cultured endothelial cells are known to synthesize and secrete cellular fibronectin (i.e.
@@
 	\item $ f = f(X,t): \Omega \times R \to \R $: the fibronectin concentration (nmol per unit area).
+
+\begin{strangeObs}[Ambiguity in the definition of one of the parameters]
+	\label{strange:AmbiguityOfTipCell}
+	As we read in \cite{Anderson1998}, $ n $ is introduced to be the ``endothelial-cell density''.
+However, I feel that it is not true and this variable should represent the ``endothelial \emph{tip} cell density''.
+I feel this way because it is the tip cells that are doing a random walk (thus diffusion) and reacting to the TAF and fibronectin gradients.
+\end{strangeObs}
 \end{itemize}
+
+\subsubsection{Equations for endothelial cells}
+The endothelial cells are moving in the space due to (1) their random motion (i.e.
+diffusion), (2) chemotaxis (moving up in the gradient of TAF), and (3) haptotaxis (moving up in the gradient of fibronectin).
+Combining these factors we will come up with the following flux for the tip cells
+\[ J_n = -D_n \nabla n + \chi n \nabla c + \rho_0 n \nabla f. \]
+where $ D_n $ is the diffusion coefficient of the endothelial cells, and $ \chi, \rho_0 $ are the sensitivity of chemotaxis and haptotaxis of endothelial cells due to the TAF and fibronectin gradients respectively.
+Putting this in the continuoity equation, we will get
+
+\[ \frac{\partial n}{\partial t} = -\nabla\cdot (-D_n \nabla n + \xi n \nabla c + \rho_0 n \nabla f) = D_n\nabla^2 n  - \nabla\cdot(\chi n\nabla c) - \nabla\cdot(\rho_0 n \nabla f).    \]
+
+\begin{strangeObs}[Growth of the tip-cells are not considered in the model]
+	As we discussed in \autoref{subsection:AndersonChaplainBiologyFacts} in \autoref{item:tipCeillsAreDividing}, the endothelial tip cells are dividing (otherwise the vascularization will not be completed).
+But in the PDE above, we didn't have any term for the tip cells proliferating or degrading.
+\end{strangeObs}
+
+\begin{strangeObs}[A problem with the diffusion coefficient of endothelial cells]
+	As explained above, we are modeling the random mobility of endothelial tip cells by diffusion.
+However, according to the discussion (page 865 in \cite{Anderson1998}), it has been observed \cite{Rupnick1988} that when the migration of the endothelial cells are constrained by the surrounding cells, then their estimate for the for the diffusion coefficient the tip cells was too large and did not agree with experimental results.
+
+	Because of this observations, authors of \cite{Anderson1998} used a diffusion coefficient that was about 100 times smaller!
+(they cite to \cite{Bray2000} for this decision).
+I feel like this is not a good approach!
+The biological observation above is a signal that capturing the random mobility of the cell simply by diffusion is not a correct framework to use.
+
+	Also, according to the discussion on page 886 of \cite{Anderson1998}), there are some in vivo experiments that indicate that there appear so the very little random mobility of endothelial cells at the capillary sprout tops \cite{Paweletz1989,Paku1991}.
+\end{strangeObs}
+In the PDE for tip cell density, the parameter $ \chi $ can depend on the concentration of TAF. We can model this interaction as (see the strange observation box below).
+
+\[ \chi(c) = \chi_0 \frac{k_1}{k_1+c}. \]
+
+\begin{strangeObs}[The chemotaxis sensitivity seems not to be correct]
+	In the model above by \cite{Anderson1998}, the authors are assuming that the functional form of $ \chi $ is given as
+	\[ \chi(c) = \chi_0 \frac{k_1}{k_1+c}. \] 
+	They say that the more realistic assumption is that the chemotactic sensitivity decreases with increased TAF concentration.
+They present the following references for this claim \cite{Lapidus1976,Lauffenburger1984,Sherratt1994,Woodward1995,Olsen1997}.
+This does not make sense for me.
+I feel like since the cell feels the chemicals by cell surface receptor binding, then the more chemicals available, the more strong is the signal for chemotaxis.
+So I feel that the appropriate model for this will be
+	\[ \chi(c) = \chi_0\frac{c}{k_1 + c}. \]
+\end{strangeObs}
+
+\subsubsection{Equations for TAF}
+To derive the equation for TAF concentration, first observe that TAF is secreted by the tumor cells where it starts diffusing into the space. I.e. the governing equation for TAF will be
+\[ \frac{\partial c}{\partial t} = D_c\nabla^2 c - \theta c, \eqTag\label{eq:initialPDEForTAF} \]
+where $ D_c $ is the TAF diffusion constant and $ \theta $ is the decay rate.
+We assume that during the initial stage of tumor secreting TAF, its distribution in the environment reaches a steady state which gives us the initial condition for the concentration of $ c $ in the domain.
+
+In the stage that the angiogenesis begins (when we have an established steady state for $ c $), while the tip cells move up the gradient, TAF is used by uptake and binding to the cells \cite{Ausprunk1977,Hanahan1997}.
+This interaction can be modeled by
+\[ \frac{\partial c}{\partial t} = -\lambda n c, \]
+where $ \lambda $ is a constant.
+
+\begin{strangeObs}[TAF interaction with cells does not seem to be correct]
+	In the model above (from \cite{Anderson1998}), the uptake and binding of TAF to the tip cells is modeled in a mass action kinetic way.
+I have a feeling that this might not be true and it should be modeled by Michaelis-Menten equation.
+Also, this model assumes that the used TAF is not get replaced, but I feel that as long as tumor is under oxygen stress, the secreting mechanism of TAF in tumor will still be active.
+\end{strangeObs}
+
+\begin{beCareful}[To much intuition is not good]
+	The way that the model above is presented is not appealing for me at all.
+This is the way that \cite{Anderson1998} does, and I think there are more than enough common sense floating around!
+For instance, they say that in the initial stage, the tumor secretes the TAF and it starts diffusing in the environment while degrading and eventually reaches steady state (which is fair up to now).
+Then, they say that during the angiogenesis stage the differential equation governing the concentration of TAF is different (Which also makes sense).
+My problem is that I want to see both of the terms in a single PDE and then by appropriate mathematical treatment (slow-fast dynamics, perturbation theory, etc) arrive at the conclusions that they stated using the commons sense.
+I need to do this for my thesis.
+\end{beCareful}
+
+\subsubsection{Equations for fibronectin}
+As we discussed in the biology background subsection, the endothelial tip cells tend to move up in the gradient of fibronectin.
+Fibronectin is present in the domain before angiogenesis.
+As the tip cells do haptotaxis (moving up {\color{red} (up or down?)} in the gradient of fibronectin) they use some fibronectin to connect to the ECM.
+Tip cells also secrete fibronectin which dose not diffuse.
+Thus the corresponding differential equation will be
+\[ \frac{\partial f}{\partial t} = \omega n - \mu n f, \]
+where $ \omega $ is the production rate of $ f $ by the tip cells and $ \mu $ is the rate at which the tip cells use fibronectin to bind to the ECM.
+
+\begin{strangeObs}[The model for fibronectin dynamics is too simple]
+	The model above for the time evolution of fibronectin is given by \cite{Anderson1998}.
+According to our detailed discussion in \autoref{item:initialPhaseOfFibroNectin}, I feel like this model is too simple to capture the time evolution of the fibronectin.
+We are assuming that the fibronectin is used and produced only by the tip cells (if by \autoref{strange:AmbiguityOfTipCell} we come to the conclusion that $ n $ represents the tip cells).
+This implies that the stalk cells are not producing fibronectin which might not be true.
+Also, as assume that as soon as the process of angiogenesis starts, the leakage of fibronectin from the parent vessel into the domain stops.
+According to the parameters in \autoref{tab:AndersonChaplainModelRealisticParameters}, the time scale happens to be $ \tau = L^2/D_c \approx 2\ \text{days}  $.
+Since in \autoref{item:initialPhaseOfFibroNectin} we discussed that the reason of the initial distribution of fibronectin is the damage to the basal lamina of the parent vessels, then I feel like in the span of several days that dame might still persist so we might still have the leakage of fibronectin to the domain.
+\end{strangeObs}
+
+\begin{strangeObs}[The initial condition for fibronectin is too intuitive]
+	\label{strange:FibronectinInitialConcentration}
+	As we will see in {\color{red} [add the cross reference for the fibronectin initial condition here]}, the initial condition for fibronectin concentration, although makes sense, but it is not appropriately modeled compared to what we did for TAF (see Equation (\autoref{eq:initialPDEForTAF}))
+\end{strangeObs}
+
+\begin{summary}
+	The complete set of equations will be give as
+	\begin{align*}
+		&\frac{\partial n}{\partial t} =  D_n\nabla^2 n  - \nabla\cdot(\chi n\nabla c) - \nabla\cdot(\rho n \nabla f), \\
+		&\frac{\partial c}{\partial t} = -\lambda n c, \\
+		&\frac{\partial f}{\partial t} = \omega n - \mu n f,
+	\end{align*}
+\end{summary}
+\begin{remark}
+	We assume that non of species, i.e. tip cells, TAF and fibronectin leave the domain of simulation. Thus the boundary condition is give as 
+	\[ \hat{\xi} \cdot J_n = 0. \]
+\end{remark}
+
+\subsection{Non dimensionalization of the system}
+For an easier analysis of the system we first need to write the system in non dimensionalized way. Consider the following change of variables
+\begin{align*}
+	& n = N_0 \tilde{n}, && c = C_0 \tilde{c}, &&& f = F_0\tilde{f}, \\
+	& t = \underbrace{\tau}_{\tau = L^2/D_c} \tilde{t}, && x = L \tilde{x}.
+\end{align*}
+Substituting these change of variables in the model we will get the following non dimensionalized form of the model. Note that we remove the tilde symbol for more clear look.
+\begin{align*}
+	&\frac{\partial n}{\partial t} = D \nabla^2 n - \nabla\cdot(\frac{\chi}{1+\alpha c} n\nabla c) - \nabla\cdot(\rho n \nabla f), \\
+	&\frac{\partial c}{\partial t} = -\eta n c, \\
+	&\frac{\partial f}{\partial t} = \beta n - \gamma n f,
+\end{align*}
+where 
+\begin{align*}
+	& D = \frac{D_n}{D_c}, && \chi = \frac{C_0\chi_0}{D_c},\ \alpha = \frac{C_0}{k_1} && \rho = \frac{f_0\rho_0}{D_c},\\
+	& \eta = \frac{L^2\lambda N_0}{D_c}, && \beta = \frac{L^2\omega N_0}{F_0 D_c}, && \gamma = \frac{L^2\mu N_0}{D_c}.
+\end{align*}
+
+\subsection{Some discussions on the parameters used}
+The following table summarizes the realistic values for the model parameters discussed in \cite{Anderson1998}.
+
+% Please add the following required packages to your document preamble:
+% \usepackage[table,xcdraw]{xcolor}
+% Beamer presentation requires \usepackage{colortbl} instead of \usepackage[table,xcdraw]{xcolor}
+% Please add the following required packages to your document preamble:
+% \usepackage[table,xcdraw]{xcolor}
+% Beamer presentation requires \usepackage{colortbl} instead of \usepackage[table,xcdraw]{xcolor}
+% Please add the following required packages to your document preamble:
+% \usepackage[table,xcdraw]{xcolor}
+% Beamer presentation requires \usepackage{colortbl} instead of \usepackage[table,xcdraw]{xcolor}
+\begin{table}[!ht]
+	\centering
+	\begin{tabular}{|c|c|c|}
+		\hline
+		\textbf{Parameter}       & \textbf{Value {[}units{]}}                       & \textbf{Reference} \\ \hline
+		$D_n$                    & $10^{-10}\ [\text{cm}^2\text{s}^{-1}]$           &  \cite{Bray2000}   \\ \hline
+		$ D_c $                  & $2.9 \times 10^{-7}\ [\text{cm}^2\text{s}^{-1}]$  & \cite{Bray2000,Sherratt1990} \\ \hline
+		$\chi_0$                 & $ 2600\ [\text{cm}^2\text{s}^{-1}\text{M}^{-1}] $ &  \cite{Stokes1990}  \\ \hline
+		$ \rho_0 $               & same as $\chi_0$ but $ \chi_0  > \rho_0 $        & \cite{Anderson1998}   \\ \hline
+		$ c_0 $                  & $10^{-10}\ [M]$                               &   \cite{Stokes1990} \\ \hline
+		$ f_0 $                  & $10^{-10}\ [M]$                               &  \cite{Terranova1985}  \\ \hline
+		$ \lambda, \omega, \mu $ & no available estimates.                          &  \cite{Hynes1989-je}  \\ \hline
+		$ L $                    &   $ 2\ [mm] $       &     \cite{Gimbrone1974}     \\ \hline
+	\end{tabular}
+	\caption{The biologically realistic parameter values for the Anderson-Chaplain model.}
+	\label{tab:AndersonChaplainModelRealisticParameters}
+\end{table}
+\FloatBarrier
+
+\subsection{Simulating the model}
+
+\subsection{Simulation Domain}
+to be expanded: square $ [0,1]\times [0,1] $. With the tumor located at $\set{1} \times [0,1] $ and the parent vessel located at $ \set{0} \times [0,1] $.
+
+\subsubsection{Initial condition for tip cell density}
+As we discussed in \autoref{item:InitialDistributionOfTipCells} the tip cells, after getting stimulated by TAF form small clusters.
+Thus we assume that the initial concentration of the tip cells are give as
+\[ n(x,y,0) = e^{-x^2/\xi_3} \sin^2(3\pi y). \]
+\begin{figure}[!ht]
+	\centering
+	\includegraphics[width=0.4\linewidth]{images/tipCellInitialDensity}
+	\caption{Tip cell initial density.}
+	\label{fig:tipcellinitialdensity}
+\end{figure}
+
+\subsubsection{Initial condition for TAF}
+Assuming Equation \autoref{eq:initialPDEForTAF} models the initial phase of the time evolution of TAF, then we can assume that the steady state of that equation is the initial condition for the phase that the angiogenesis starts happening.
+Depending on the initial configuration of the tumor in the domain, \cite{Anderson1998} came up with two biologically interesting initial condition.
+
+\textbf{Case 1: A column of tumor cells or larger circular implant.} For this case the initial concentration of TAF will be
+\[ c(x,y,0) = e^{-(1-x)^2/\xi_1}, \qquad (x,y) \in [0,1] \times [0,1]. \]
+
+\textbf{Case 2: A small circular tumor centered at (1,1/2)}.
+In this case the initial concentration of TAF is given by (\cite{Anderson1998} and they reference \cite{Chaplain1995,Chaplain1996} for the equation)
+\[ c(x,y,0) = \begin{cases}
+	1, &\qquad 0\leq r\leq r_0, \\
+	\frac{(\nu - r)^2}{\nu - r_0}, &\qquad r_0 \leq r \leq 1,
+\end{cases} \]
+where $ \nu $ is a positive constant and $ r $ is given by
+\[ r = \sqrt{(x-1)^{2} + (y-1/2)^2}, \]
+assuming that the tumor is centered at $ (1,1/2) $ with a radius of $ r_0 $.
+Note that $ \nu $ is in fact showing the concentration right at the edge of the tumor.
+To make the transition between concentration $ c $ inside the tumor (i.e.
+$ 0 \leq r \leq r_0 $) and outside the tumor (i.e.
+$ r_0 \leq r \leq 1 $) continuous, the value of $ \nu $ should be $ \boxed{\nu = r_0 + 1} $.
+We can see this by simply trying to force the function $ c(x,y,0) $ to be continuous.
+The following figure illustrates the initial concentration of TAF for $ r_0 = 0.2 $ (thus $ \nu = 1.2 $).
+\begin{figure}[!ht]
+	\centering
+	\includegraphics[width=0.5\linewidth]{images/TAFInitialConcentrationSmallTumor}
+	\caption{Initial concentration of TAF for small tumor with radius $ r_0 = 0.2 $ centered at $ (1,1/2) $. The green region shows the tumor inside which the concentration of TAF is assumed to be 1.}
+	\label{fig:initialConcentrationOfTAFSmallTumor}
+\end{figure}
+
+\begin{beCareful}[I am not sure if the formulas above are really the steady state solution of TAF PDE]
+	Note that the initial concentration of TAF should be the steady state of Equation \autoref{eq:initialPDEForTAF}.
+\cite{Anderson1998} claims that the two cases above are the steady state solutions for the early stage TAF PDE.
+I have not checked this yet and I need to evaluate if this is really true.
+\end{beCareful}
+
+\subsubsection{Initial condition for Fibronectin}
+The initial condition for fibronectin has been a more phenomenological one in \cite{Anderson1998}. They Assumed the initial concentration of fibronectin is given by
+\[ f(x,y,0) = k e^{-x^2/\xi_2}, \qquad (x,y) \in [0,1] \times [0,1], \] 
+where $ 0<k<1 $ and $ \xi_2 > 0 $.
+As discussed in \autoref{item:initialPhaseOfFibroNectin} this choice of initial conditions is justified up to some degree.
+However look at \autoref{strange:FibronectinInitialConcentration} for a discussion about possible alternative approaches to find a more realistic initial condition.
+
+\subsection{Make it a Combo! Adding Stochastic Model to Generate Vascular Networks}
 

### 2024-06-14 Findings (Shared Folder)/Chapters/modelAttempts.tex
new results

@@
 
-\subsection{Simulation Domain}
-to be expanded: square $ [0,1]\times [0,1] $. With the tumor located at $\set{1} \times [0,1] $ and the parent vessel located at $ \set{0} \times [0,1] $.
+\subsubsection{Simulation Domain}
+The simulation domain is the square $ [0,1]\times [0,1] $. With the tumor located at $\set{1} \times [0,1] $ and the parent vessel located at $ \set{0} \times [0,1] $.
+
+I implemented this model in the Visual PDE online solver, with the parameters that both their result look good and also the default solvers of this tool are stable.
+You can find the model \href{https://visualpde.com/sim/?options=N4IgRgqiBcIHYFMCuBbAhnOIA04BqM8y6mO4A6oYqhlrgMaEBuAlgE4sAmLAzmfQBEYAFzZIEuThCiwhkiAVgBaAAQANAPQAKAIwBqNACp6ASkN0QUysrbmynPIpAAGe+Wsv7fWDud-7wjDOAHR+AKy4AGYAkiJiEiD0MiAIAB4ADlpKIX46hqkAegBMZlo8LHBaAMyG6SyGWgCeGgAyAPqNJl3F-E5pmdmhfn41Wqkqqu2pJsUm-B79WTnDeWMzJWQA1oQCKgC8KjoA3Gr7KgBsfkdoZ8tHbGdhzkfpZ8dgt0NXAOafzmFHMjoQghKoAdiBLBBZCwsCqMIAKjAns5cOkEGxIgB7NgkegIACyWM4CDi4jRhAAwkgeMIsSgyOk2FiwAgAGJIWHwRnM1ktBBwb7CAAWMCqw1wbGSrklTiU6XMxjIbA8YEMkQmKm+isiZB46UIvj1XLgKnoKl1uB4AEcyQlhJw7bgmDAANqgOBoFCknxkLFMDEAGzQjQAohk2IaeSz2ZyqNHWWpoWjeQgAJrJkA8JCYtD4jlcmUgADuwrQwgRWIACoGsYFYFgAL7YUBoNjM4uUrG1nMwc5hMJVCIgNsdgQC8rCRpBYLD0dY4v8wUigloVJR3Dz4sAZXoaEDPpHSDpZC3SdgUw0RU16837YXGYvHSvmunm+PWO3CGEXZ7bAAShg3w+pE+48Ake7emwaAIsK35oGKqKJF6GJoFWwpQtA4oMCh0EAFpYvSM4ABwMN2WI5ugBqwKwHDcHwZF-mAbYwKBgbgWRcB0jm3hsRxiRYlxFFsL+wlBJx3FsKG6TlLWsI5DoEnCQAcqgyIMDSdIoNuOagfirFgQkCAoGAWI8Lxhm4MZpnmQAgiZLACvWITDtZZk8AILCRJENI+iE5xWSZ7noZhOizmCzikSkQXmduGFwNO0C+IFNk8NuKCESKiDmTORQpe527ovQSDBpG0AhHl0WpbB8EzmCxFhAALFEgYsOk6KcKJlFoNRfEJK1iDkFwK4lTAikgOgqRdWweD7uIILBOCkJwNNs2BvNsBFv6QYhgZ7EJNtbDBo003iSAh3HdJsmCTO4S4BdIbhkyG7nQGR0hi0FQIENnAjYGY1orWFaNOihDpMGiAJhU3x7fxTIxgW8YpjGS5CqKWESiA8OsgiIOHhUwgIN80H-cjiaZtj6aZtmub5nGm1kCKbAIAgAgABIICw3zCsIu77n5wRgs6CD0HSbBso5gaOtAfW4KW5aVjWdaEIwuCet6MBFM2rb3p25G9tA-aDnOuvjnAk6JS5d4dqjK5ri9W58wehBoB+p66+eICXteqi3iOuuPl7z4+yob5HnSX4-vrAFASBlnIVBMFwcICEYzhichYh6eoQRRHlcEUX0NHVHMOwXC8Pw0fMWVssCUJPGwxBgmSad5VKTmV0sHJt3jUX9dsKpDLQMOxW0vSOm03H+35TlMvx25dkOU5M6uTFHleT54EzgFVXBRhY3hZFM9pfFiXJbvsUZXWwrZd4FXH4VIslSx+eVQvPA1SndUNc1ICRK17UECdWLj1RuuABrfWGsKAko0kq4EmqtOah5QQQngRURB61kF+jesdMBr0donWjmdB6YYZJdxuvnO6+D3phgjC9Ehn1BpQJgf9OBWMga41BrAcGGBSRk2hngymiMGwJgQLbdG4o-BkwQJw-GXEiYk1EZ7IslNA5FhpnpWMhZGbCmZqzDmXMeZOwFkLEAAZRY4glggKWeD5YVmrEDQglp4AoTFI2AAunLaW3I5bUR8VmFgKAEQsGEM7WAtlBQsCxMBbKvBNQRJJDwQSKhKRlh4RUEAjYgA}{here}.
 

### 2024-06-14 Findings (Shared Folder)/Chapters/randomNotes.tex
Updates from Overleaf

@@
 \subsubsection{Spiraling in to the Origin}
-We can start with the vector field $ F(x,y) = (x,y) $ and this time, rather time rotating it with a rotation matrix, we can rotate that using a matrix corresponding to a complex number $ z = \mu + i \omega $, i.e.
+We can start with the vector field $ F(x,y) = (x,y) $ and this time, instead of rotating it with a rotation matrix, we can rotate that using a matrix corresponding to a complex number $ z = \mu + i \omega $, i.e.
 \[ F_\text{new}(x,y) = \matt{\mu}{-\omega}{\omega}{\mu}F(x,y). \]

### 2024-07-06 Findings (Shared Folder)/Chapters/theoriticalInvestiagtion.tex
new notes added

\chapter{A Theoretical Investigation}

In this chapter I will be trying to appropriately formalize the problem in hand while being very careful about the hidden structures in the problem.

\section{The Set of All Vascular Structures. Is That a Manifold?}

I start with my formal definition of a valid vascular structure.
\begin{definition}
	Consider $ \Omega = [0,1] \times [0,1] $ as a closed subset of $ \R^2 $. A valid vasculature, is a network in the domain, where there is at least one path that starts from $ \set{0}\times (0,1) $ and reaches $ \set{1}\times (0,1) $.
\end{definition}
\begin{remark}
	The definition above needs to further formalized. For instance, what is a network? if I mean a graph theoretical concept, then I need to formally define it.
\end{remark}

Let $ \mathcal{N} $ denote the set of all valid vascular systems. Introducing the notion of the set of all valid vascular structures, if we have no information about the topology or algebraic properties of that set is no more useful than re-naming an object.

We can consider a particular vascular system as a mapping $ s: \Omega \to \R $, where a given vascular network will be just a domain coloring of the map $ s $. Then the set of all vascular systems $ \mathcal{N} $ will be the same as the function space of all functions mapping $ \Omega $ to $ \R $. This set is far too complex than being useful. So I will follow an alternative approach.We will use the notion of graphs to represent a particular vascular system.

\begin{observation}[Vascular System as a Graph]
	A weighted multi graph (where we allow to have multiple weights on an edge, and also allow assigning weights to nodes) can capture any necessary information of a vascular system that determines the flow of blood through it. For instance, we can have the following correspondence between a given vascular system and  a multi graph.
	\begin{itemize}[noitemsep]
		\item Denoting the branching points as nodes.
		\item Denoting the vessel segments as edges.
		\item Assigning a weight $ w\in \R^n $ to each edge, where $ n $ is the number of important characters of the vessel segment that affects the dynamic of blood flow. These weights can be the length of segment, the radius of segment, etc.
		\item Assigning a weight to $ v \in \R^m $ to each node, where $ m $ is the number of important parameters at each node affecting the fluid dynamics of blood in the vessels. For instance, the hematocrit splitting ration depends on the number of branches in a given node (which is reflected by the degree of a given node), and also the angle between branches. The angle between branches, etc can all be considered in the weight assigned to each node.
	\end{itemize}
\end{observation}

The observation above is just a general blue print. We can obviously make some simplifications and reduce the number of weights at each edge.

\begin{observation}[The Set of All Vascular Systems is a Manifold]
	For now, let's consider that for each vessel segment we assign one weight that can be the radius of the vessel segment, or a combination of parameters, like flow as given by the Poiseuille's law. For each node we assign no weights (for instance we assume that the hematicrit splitting is not significant). Then we can identify each vascular network with an adjacency matrix of the associated graph.  Thus the set of all vascular systems with $ n $ nodes will be the same as the set of all $ n\times n $ matrices. 
	\begin{proposition}
		The set of all $ n\times n $ matrices topologized with the subspace topology of $ \R^{n\times n} $, is a smooth manifold.
	\end{proposition}
	Thus the set of all vascular systems with $ n $ nodes is a smooth manifold.
\end{observation}

\begin{remark}
	In the observation box above we stated that the set of all manifolds with $ n $ nodes is a the same as the set of all $ n\times n $ matrices. But can we have any statements about the set of all vascular structures with any number of nodes? For this case I think I need to design a graded structures (as in the case of graded algebra) that can capture the whole set of all vascular structures.
\end{remark}

Intuitively, we tend to identify the vascular systems that have the same total flow across the network. To put this in a useful theoretical ground we have the following observation box.

\begin{observation}[Vascular Systems Identification Based on the Total Flow in the Network]
	Let $ \mathcal{N}_n $ denote the set of all vascular networks that has $ n $ nodes. Define the map
	\[ F: \mathcal{N}_n \to \R \]
	which is the total flow of the network. Then we define the following equivalence relation. For $ x,y \in \mathcal{N}_n $
	\[ x \sim y \quad \Leftrightarrow \quad x,y \text{ are on the same level set of $ F $ }. \]
\end{observation}
As we will see later, this map is very important and can be computed by considering the rheology of blood and the geometry of the network. The observation box below shows that any intervention induces a vector field on the manifold $ \mathcal{N} $ and as we will see later, the interplay between this vector field and the gradient of the map $ F $ turns out to be very important.

\begin{observation}[Any Intervention Induces a Vector Field on $ \mathcal{F} $]
	First, I need to mention that since I have not yet developed a graded structure to handle different vascular systems with different total number of nodes as a graded manifold (or some similar construct) I will assume that the interventions do not change the number of nodes. But this is just a bureaucracy and I will resolve it as soon as I develop rich structures.

	Every intervention induces a vector field on the manifold $ \mathcal{N} $ denoted by $ \mathcal{I} $. So the time evolution of a given vascular system due to particular intervention (like radiotherapy or chemotherapy) will be given be a one-parameter group of diffeomorphisms of $ \mathcal{N} $. In other words, the time evolution of a given vascular system $ p \in \mathcal{N} $ will be the integral curve of $ \mathcal{I} $ starting at $ p $.
\end{observation}
\begin{remark}
	According to the remark above, we can model the angiogenesis process as a background vector field on the space of all vascular systems. Any other background biological mechanisms can also be lumped in this vector field. We denote this vector field as $ \mathcal{G} $.
\end{remark}
\begin{remark}
	The vector space $ \mathcal{I} $ due to intervention can be possibly learned by an appropriate machine learning system.
\end{remark}

\begin{openQuestion}[Can a Machine Learning System Learn These Vector Fields?]
	Our ultimate goal in this project is to determine the mapping $ F $ as well as the intervention vector field $ \mathcal{I}_i $, along with any background vector field $ \mathcal{G} $. Denote the intervention and the background vector fields as 
	\[ \mathcal{I} = \mathcal{I}_i + \mathcal{G}. \]
	One possible parallel approach to determine these vector fields is to use appropriate machine learning system along with a suitable simulator the learn these vector fields.
\end{openQuestion}

\begin{observation}[Effectiveness of the Intervention is the Lie Derivative of $ \nabla F $ with respect to $ \mathcal{I} $]
	Formalizing the problem in the structure above is useful from many points of view and can be a unifying theme.

	For instance, let $ F:\mathcal{N}\to \R $ denote the total flow mapping and $ \mathcal{I} $ be the induced vector field due to intervention (and possibly plus any background process like angiogenesis). Then if $ \mathcal{I} $ align with $ \nabla F $ we will have the maximum effect due to therapy, and if $ \mathcal{I} $ gets perpendicular to $ \nabla{F} $ we will have no effect (i.e. the intervention will have no effect on the flow). Thus we can introduce a notion of effectiveness of intervention which is the Lie derivative of $ \nabla F $ with respect to $ \mathcal{I} $. In other words
	\[ E = \mathcal{L}_{\mathcal{I}} \nabla F = [\mathcal{I},\nabla F]. \]
	Alternatively, we can define other measures. For instance, assuming that the manifold $ \mathcal{N} $ is a Riemannian manifold, hence we have the notion of inner product at a given tangent space, we can define
	\[ E = \langle \mathcal{I}, \nabla F \rangle . \]
\end{observation}

\section{Statistical Properties of Vascular System}
For me, the phenomena of angiogenesis if very similar to the phenomena of percolation. I am not claiming that it is the same phenomena, but I am emphasizing the similarities. For instance, some fluid is percolating, then instead of a uniform flow of fluid through the porous media, individual channels form (I think due to the surface tension of water, the water molecules tend to follow the pre-existing water channels rather than making a new one through wetting the surface).

I think at some step, I would need to consider percolation models and their statistical signatures to see if I can use somewhat modified versions of that.

### 2024-07-06 Findings (Shared Folder)/Chapters/theoriticalInvestiagtion.tex
new notes

@@
 \end{remark}
-\begin{remark}
-	The vector space $ \mathcal{I} $ due to intervention can be possibly learned by an appropriate machine learning system.
-\end{remark}
 
-\begin{openQuestion}[Can a Machine Learning System Learn These Vector Fields?]
+\begin{openQuestion}[How to Determine These Vector Fields?]
 	Our ultimate goal in this project is to determine the mapping $ F $ as well as the intervention vector field $ \mathcal{I}_i $, along with any background vector field $ \mathcal{G} $.
@@
 	\[ \mathcal{I} = \mathcal{I}_i + \mathcal{G}. \]
+	I am still thinking actively how my modeling attempts (like agent based model, or the PDE model) manifest themselves in this general framework?
+How these models can ultimately help to determine these vector fields, at least locally.
+\end{openQuestion}
+
+\begin{openQuestion}[Can a Machine Learning System Learn These Vector Fields?]
 	One possible parallel approach to determine these vector fields is to use appropriate machine learning system along with a suitable simulator the learn these vector fields.
@@
 
+\begin{observation}[Homogenization and Darcy's Law]
 	I think at some step, I would need to consider percolation models and their statistical signatures to see if I can use somewhat modified versions of that.
+After making this connection.
+I can apply the idea of homogenization and possibly Darcy's law suggested by Leah.
+\end{observation}

### 2024-07-11 Findings (Shared Folder)/Chapters/PapersReviewed.tex
new notes added

@@
 \end{itemize}
+
+\section{Structural Features of Microvascular Networks Trigger Blood Flow Oscillations}{Y. Ben-Ami}{2022, Bulletin of Mathematical Biology}
+
+\subsection{Introduction}
+
+\subsection{Methods}
+
+\subsection{Useful facts}
+\begin{itemize}[noitemsep]
+	\item Some structural properties of the vascular network can trigger self-sustained oscillations in the blood flow through the network. This has also been shown experimentally \cite{Kimura1996-od}
+\end{itemize}
+
+\subsection{I Need to Think More ...}
+
+\subsection{Useful Papers Cited}
 

### 2024-07-22 Findings (Shared Folder)/Chapters/modelAttempts.tex
new notes to the lab report
new updates

@@
 \end{itemize}
+
+\subsection{Optimal Transport}
+I just came across this \href{https://www.math.ucdavis.edu/~qlxia/summary/transport.html}{website} about the computational approaches to the optimal transport.
+I was shocked how similar their plots was to the vascular structures.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.7\linewidth]{images/optimalTransportAndAngiogenesis}
+	\label{fig:optimaltransportandangiogenesis}
+\end{figure}
+\FloatBarrier
+After this observation and looking at some similar plots it just came to me that the angiogenesis is really an optimal transport problem where we want to deliver nutrients and oxygen in an efficient way to every single cell of the body.
+The work by \cite{Pegon2019-ww} triggered this idea to see if I can use some of the theoretical tools developed in the optimal transport literature to my benefit.

### 2024-08-11 Findings (Shared Folder)/Chapters/branchingMorphogenesis.tex
new notes added

\chapter{Branching Morphogenesis}

\begin{itemize}[nosep]
	\item Fisher-KPP model for branching morphogenesis in VisualPDE: \href{https://visualpde.com/sim/?options=N4IgxiBcIG4JYCc4BM4GcQBoTIKq6hABEscA1MwgBlOQHU7raNoBGKj0sA6G7Aa0JEABAF5hrAHRVhcAHbCA2q0zsAugG4EAfQBGY4VWntZCxVVWa52meIDM000osBWTToCmBo6xdPzlhoANt7SMvLOmAAsmgC2oRz+FjEapLEAhtSSdgDsaXBM2AAOhADCAK5oAC4A9rGkRQg1uh4AYuVyhOUNTS0AMh5yAOZVABZQdhwWIAg8M3oAVOULABSsALQr5QDUMACUAPQr1lR7e8LrwrFLpAiU0J5LwtvCKzq6Byd7S6s7+xdBBYwW6MXikNAlaB2UjAyCKUDpBBNADupRqQRq5QQUAAbC4XHYXNhESiiIM0HAqgBPKBGIkgEk1ZEDYZjACy6QAHoRWKRGciAMpgdJBDyEdLlWp8pFMgAahD62k5BwATBdhNziTLkQBNBXaKmq9U04mSmoCjxVNEYrEAJXSwzFkAAZiK0B5sMLYh4EOkACqjS2ZSCTT3pb2+gAKowKIemXp96QAWjU6rTJAAOT3ozEIDKQ2CIFDoLg5rG6RFQV1Bd3ZuS1LEsau18A1eu56252l1hsIACiRQpGM6kCMVBUrfbWIAcuV6pB6WBKrVYgKsa6wE7mx6QB5Yroamgm26d3uD0eAIL7uCDKrp+lnw9oIhwZ3OypOow47CPo-R2NSC4ORUFmu77k+AoxnINKQOwP7gUeAqxKmYxyB4R7piq8HnmgApFB4S5BJWo6SFhYE4QGQbpjkGYuFE2DOkEcBFPhyCdli+ZVie2BMWhdAoOy5RBFAE4ZJy7F3CK5ROiARi5PkcgSWQUkyXwIA1DAPpETB27YBpWnpFSEndupmkINpA5Dm26YcPS+nmYZfaco0PKkPZ2l9PIHj8cggnCbBxQYlUfpUvhhBFERaE9M08hDFxNY7o0zRtB0XTRf0gwjOMcYcMUvQeCFYXQPIVQeEMvrCXlyXymCVUtHqtUgGg67pJu7QjrJpBjAgHgeEQAASHhwEMoxVEKIqfpIOTYJpYC1AgrQ3kEyDxS2yKjOkwU1JGQVpdgcjhjJvIAL6YAi2rGZAeIEvS-JknIFLUveWooiyWUctybDSii42iuKZrfXK+rKmqlyagy2oNSAiqGqDwgmgyZoWlaZYIPajqrTuCa+pRVTBqG4CHVGMYTPGRPJqm85GKBYCo5x0DwEgqAYNmNoIBW2Iutxk69seCU9h2qMmbTU79oOcDDjZE4i72s7zouy51GuCAblu3O-nzLYa1eug3vWz3kU+L5vh+6bfobf4k7BkhASB2EQVBMFwRbuHITUqHoSwRhkRreEEUJxHe-bR648GRg0XRDFMSxHhsXT6SQrpIC8d5AmjGyQkidgYlKSpWTydn8i50E0mFKZBk6dz7mGZd0zV1SlkS9ZJG2XpZkWc5nMgLybcV55fFpxn-kThF7uFTJEUOmKdWxZjdUpR13Tz29Yyk7lIBJS04+ECVZUVelHg1Z189Q2pzUq61C9l91vUDUNI1jcKf0kdNsAEfNi0eMtc8gOtm1+ttXaDNSAHW9FAFUx01DYGRCtaAS9f4FngRSWIfpKTPxAItNAgYEDrAANKRkjMINkNRkBf2EM6GoCBhAACFfRyDAFBIYRDKFFFGDUIYgx0IlmOkAA}{Link}.
\end{itemize}

### 2024-08-12 Findings (Shared Folder)/Chapters/branchingMorphogenesis.tex
new notes added

@@
 	\item Fisher-KPP model for branching morphogenesis in VisualPDE: \href{https://visualpde.com/sim/?options=N4IgxiBcIG4JYCc4BM4GcQBoTIKq6hABEscA1MwgBlOQHU7raNoBGKj0sA6G7Aa0JEABAF5hrAHRVhcAHbCA2q0zsAugG4EAfQBGY4VWntZCxVVWa52meIDM000osBWTToCmBo6xdPzlhoANt7SMvLOmAAsmgC2oRz+FjEapLEAhtSSdgDsaXBM2AAOhADCAK5oAC4A9rGkRQg1uh4AYuVyhOUNTS0AMh5yAOZVABZQdhwWIAg8M3oAVOULABSsALQr5QDUMACUAPQr1lR7e8LrwrFLpAiU0J5LwtvCKzq6Byd7S6s7+xdBBYwW6MXikNAlaB2UjAyCKUDpBBNADupRqQRq5QQUAAbC4XHYXNhESiiIM0HAqgBPKBGIkgEk1ZEDYZjACy6QAHoRWKRGciAMpgdJBDyEdLlWp8pFMgAahD62k5BwATBdhNziTLkQBNBXaKmq9U04mSmoCjxVNEYrEAJXSwzFkAAZiK0B5sMLYh4EOkACqjS2ZSCTT3pb2+gAKowKIemXp96QAWjU6rTJAAOT3ozEIDKQ2CIFDoLg5rG6RFQV1Bd3ZuS1LEsau18A1eu56252l1hsIACiRQpGM6kCMVBUrfbWIAcuV6pB6WBKrVYgKsa6wE7mx6QB5Yroamgm26d3uD0eAIL7uCDKrp+lnw9oIhwZ3OypOow47CPo-R2NSC4ORUFmu77k+AoxnINKQOwP7gUeAqxKmYxyB4R7piq8HnmgApFB4S5BJWo6SFhYE4QGQbpjkGYuFE2DOkEcBFPhyCdli+ZVie2BMWhdAoOy5RBFAE4ZJy7F3CK5ROiARi5PkcgSWQUkyXwIA1DAPpETB27YBpWnpFSEndupmkINpA5Dm26YcPS+nmYZfaco0PKkPZ2l9PIHj8cggnCbBxQYlUfpUvhhBFERaE9M08hDFxNY7o0zRtB0XTRf0gwjOMcYcMUvQeCFYXQPIVQeEMvrCXlyXymCVUtHqtUgGg67pJu7QjrJpBjAgHgeEQAASHhwEMoxVEKIqfpIOTYJpYC1AgrQ3kEyDxS2yKjOkwU1JGQVpdgcjhjJvIAL6YAi2rGZAeIEvS-JknIFLUveWooiyWUctybDSii42iuKZrfXK+rKmqlyagy2oNSAiqGqDwgmgyZoWlaZYIPajqrTuCa+pRVTBqG4CHVGMYTPGRPJqm85GKBYCo5x0DwEgqAYNmNoIBW2Iutxk69seCU9h2qMmbTU79oOcDDjZE4i72s7zouy51GuCAblu3O-nzLYa1eug3vWz3kU+L5vh+6bfobf4k7BkhASB2EQVBMFwRbuHITUqHoSwRhkRreEEUJxHe-bR648GRg0XRDFMSxHhsXT6SQrpIC8d5AmjGyQkidgYlKSpWTydn8i50E0mFKZBk6dz7mGZd0zV1SlkS9ZJG2XpZkWc5nMgLybcV55fFpxn-kThF7uFTJEUOmKdWxZjdUpR13Tz29Yyk7lIBJS04+ECVZUVelHg1Z189Q2pzUq61C9l91vUDUNI1jcKf0kdNsAEfNi0eMtc8gOtm1+ttXaDNSAHW9FAFUx01DYGRCtaAS9f4FngRSWIfpKTPxAItNAgYEDrAANKRkjMINkNRkBf2EM6GoCBhAACFfRyDAFBIYRDKFFFGDUIYgx0IlmOkAA}{Link}.
+	\item Guo pattern formation model for side branching: \href{https://visualpde.com/sim/?options=N4IgRgTiBcIAwDoCsIA0IDGMQDcCWEeAJngM5ohECqV2AIgPoCCFRAam-QwJKsDqfLgGVWAFxiI4ARnQZasAEpMAcnQBU3BkLVS4FDJ1hSKAay5MABAF4LkgEwW8AOwsBtOKikBdANyNu1rYIdgBsji7unr6MQoGSYc5uHt4+GHEIcHAOiZEpALYArulSCRHJvhAAFgD2zOlwAMzhSVE+TkU2kgAszbkVNTz1mY295akMcPXZZa0A5gCGeXnzU6OtAKYADqR4ADbVLp0IABxIaynrDFLFkzljREM9d60AZoG6574Uy9iIDQDs3zwvwoTmwXQom2wAGECqRRNU8pCINUwOsAGIFMGwFjoTYotEAGXWTlmokqMAaw3QEHkmDUTAAenY1EIAPTcAC0hQZAGoqrUmGoAJoUCCGelMlmxTkWdoaCz8gbcEViwSwDATCyyhZLeas7UWLY7fZOEWssUARWwRAZhsuUjUFmFiosAAphcyAJRst1SXkvEXeiikbEgSwBWKi9CkACOMFEEAK63QOBgrlA8wgKIA7tDqvsClBoCEkEgGkh0Fnc3QSTtRABPCTIKvZ6o54mk8kAWXmAA9sMZW7mhBh5rt1th5gUERRq+2ABrYQkMPtshyygfD9ui2Arhvrw1Nqsz6pCdaifOFiAKeakyfQF7j0gpzCLdYQeYAFUqF-mlI8N88g-eYAAVKmBaAqVkd9PwALWqRFm2OWQC2qItlihWB8EIEhyFQ68wCzGAn12F9UKcBEi3IR9n1fDADioiAr3Q4tAIYyjWIAUW2PYDmbaQKKY5QCiRaBK0wOEETyIQiyfDAH1I8iQHWPIwGqUgaKU19VPUzSmDUvASXEaBEAk3SNNIOg8BeF44QfRAQnQCzNPAyCpGQf44BQlS1MsoQIKcJtoF0Zy-M0oQ8kQ8knHWTTmzsMK9NIIRNnWDACl2YjTOCJLLJ-P9m3+U4unQF5djwTY0qIFiMPmLDtPQCrYr4Yge0ymAZBAZY+1q8Vx2TX4EABIEnD6tgBofeAKGqHAPyy4LGpAWb5vmBs+okdAVogBaeJNficsyCTtt2vt8UHGa5p2tbCWcdZWqIdrdk6vF9lEL8GzS7BNiy2LkVRZxZhIui8QJDEsWwXEQHxVF1k7MkKSg6lobBj6vtgZxRHWWZP2e0HYaXWA9HxtFd2mmM5PmBTMTDYmQHJCB1nWOgAAl1jwWZKlEUdxwchB-lTdKEQgdEjN2IhgbI18c0qeZ3uqUC3sh0F3wugBfLx0BzCWcQoHMsPDEM8DyL88FECdsAAcQKaoLG7aoiHWXYLCEYh1gsAAhT8nAwQKgbVoA}{Link}
 \end{itemize}

### 2024-08-13 Findings (Shared Folder)/Chapters/branchingMorphogenesis.tex
new notes added

@@
 	\item Guo pattern formation model for side branching: \href{https://visualpde.com/sim/?options=N4IgRgTiBcIAwDoCsIA0IDGMQDcCWEeAJngM5ohECqV2AIgPoCCFRAam-QwJKsDqfLgGVWAFxiI4ARnQZasAEpMAcnQBU3BkLVS4FDJ1hSKAay5MABAF4LkgEwW8AOwsBtOKikBdANyNu1rYIdgBsji7unr6MQoGSYc5uHt4+GHEIcHAOiZEpALYArulSCRHJvhAAFgD2zOlwAMzhSVE+TkU2kgAszbkVNTz1mY295akMcPXZZa0A5gCGeXnzU6OtAKYADqR4ADbVLp0IABxIaynrDFLFkzljREM9d60AZoG6574Uy9iIDQDs3zwvwoTmwXQom2wAGECqRRNU8pCINUwOsAGIFMGwFjoTYotEAGXWTlmokqMAaw3QEHkmDUTAAenY1EIAPTcAC0hQZAGoqrUmGoAJoUCCGelMlmxTkWdoaCz8gbcEViwSwDATCyyhZLeas7UWLY7fZOEWssUARWwRAZhsuUjUFmFiosAAphcyAJRst1SXkvEXeiikbEgSwBWKi9CkACOMFEEAK63QOBgrlA8wgKIA7tDqvsClBoCEkEgGkh0Fnc3QSTtRABPCTIKvZ6o54mk8kAWXmAA9sMZW7mhBh5rt1th5gUERRq+2ABrYQkMPtshyygfD9ui2Arhvrw1Nqsz6pCdaifOFiAKeakyfQF7j0gpzCLdYQeYAFUqF-mlI8N88g-eYAAVKmBaAqVkd9PwALWqRFm2OWQC2qItlihWB8EIEhyFQ68wCzGAn12F9UKcBEi3IR9n1fDADioiAr3Q4tAIYyjWIAUW2PYDmbaQKKY5QCiRaBK0wOEETyIQiyfDAH1I8iQHWPIwGqUgaKU19VPUzSmDUvASXEaBEAk3SNNIOg8BeF44QfRAQnQCzNPAyCpGQf44BQlS1MsoQIKcJtoF0Zy-M0oQ8kQ8knHWTTmzsMK9NIIRNnWDACl2YjTOCJLLJ-P9m3+U4unQF5djwTY0qIFiMPmLDtPQCrYr4Yge0ymAZBAZY+1q8Vx2TX4EABIEnD6tgBofeAKGqHAPyy4LGpAWb5vmBs+okdAVogBaeJNficsyCTtt2vt8UHGa5p2tbCWcdZWqIdrdk6vF9lEL8GzS7BNiy2LkVRZxZhIui8QJDEsWwXEQHxVF1k7MkKSg6lobBj6vtgZxRHWWZP2e0HYaXWA9HxtFd2mmM5PmBTMTDYmQHJCB1nWOgAAl1jwWZKlEUdxwchB-lTdKEQgdEjN2IhgbI18c0qeZ3uqUC3sh0F3wugBfLx0BzCWcQoHMsPDEM8DyL88FECdsAAcQKaoLG7aoiHWXYLCEYh1gsAAhT8nAwQKgbVoA}{Link}
+	\item Guo pattern formation with better parameters \href{https://visualpde.com/sim/?options=N4IgRgqiBcIHYFMCuBbAhnOIA04BqM8y6mO4A6oQCYCWATjQMYAWANggC5lgCK19TNp251CABgB0AVjJUIUWABEA+gEFZeAkuUBJWeUraAyrK7RJYgIy5GCkACVVAOUUAqHcqOvLY12LKMWiCWZADWhCqqAARRALwxFgBMMTRwMQDaYthRlgC6ANwqOjHxUZKJYilpUZnZeYWeJQkSAJwAbFUZWTkFjE1lEmJiyVGpXXUFKEj9kpYdo9W1Pfl0zAD2ajODAMydNd31cNNxzWIALHtL9asbxScDQ0O7C+PLjMqV90mXBwUA5mgUOgtsMfhN8ggAA4AZxorDW1VKkgAHFIwcsEMpLFtsS99uCqCDzuj6gAzfo+EkFMjocQSbYAdhpNHEZCw0DOuDgABUYGchrhIYQAMJIaEcNYoMiQuhrMAIABiSHZIHUgtl8oAMgg4H8OMwYNtHrgGDAOHQkAgTXZGK5VAA9RKuIwAeh0AFopnaANQ3NSuACaZDoQVtDqdRii7qiR3cUV9610geDhhA70q0YBQLQzqjUShsPhcEDzuDfFgVDtecxllcUQD8aiAAoA46AJQupuWb2kwPtsjQlXRYqRoO4aEARzNFqtIAAbjB0qA0HRZQB3YVreFIUTQNpSKTbKS4FfrxQ62EcACeMEkx5Ap7Wa+1uv1AFk0AAPQghE+rp9GIwaDsIQaBIBKZCPmuAAahCason4usk0bfn+65BrA8FXkheY3ie4FrEYnCbtudD2BgfwIDApLAdCs5ASgCB0Gg3LMJwaCGlkaaAkxaAAArMCy0BGjYPHMQAWmskq3hIyI2Fuaw7ugQqwBwO5gGsAQKepK7UbR9EIhKO7QnprB0fJcBGXQJGKbuXGMIZtkAKIwnCCIyVYFlWU4qAwPejBihKKBGDuNGMFR0A0WZs4ICgGnQiZkX6bgsXxdCqhxTQOpmHeKVxWsCWKDQpKkmKEWSG0eVpQJQmWNIDJiHJICpQV0JGIJcA3tAPhVa1RgoFJ+qIAlMmJL1CVGJCCABawunmBIY3NflCWsexMkMqinIgKSrA0JCU1UDZSloCpUXmSAu2IOQNBUO+SCsDA1ggOgn5HSGwGWnSjLMnAb14B9EUgP4uBrHOTGzV1Z2zqD4NoFeb23iDYN0BDLmFu581DPeMMo3DTmfjKP5kDjEOaqkCDXbdzBvvdj2CvCHDcleU2EJCs2INKGqpH8pnnTKcqKsqhBqiA-NajqeoGsJxqixqCBMyzsCpBwCB-MxD3qgLsGwMDssCxhQMDqFaDhUqKq6-qdAIAgigABIIDQfzMBwgHAeVEgMrgYOMBKdAKllrBULzs5rswaCM2sfEM8LbI8UTAC+uS4GuQewCLa4qaqA40Cg3I0BwIGwAA4kgaxRG+axUAgrBREYN0IFEABCzFwCw3MgPHQA}{Link}
+	\item Guo mostly side branching: \href{https://visualpde.com/sim/?options=N4IgRgqiBcIHYFMCuBbAhnOIA04BqM8y6mO4A6oQCYCWATjQMYAWANggC5lgCK19TNp251CABgB0AVjJUIUWABEA+gEFZeAkuUBJWeUraAyrK7RJYgIy5GCkACVVAOUUAqHcqOvLY12LKMWiCWZADWhCqqAARRALwxFgBMljE0cDEA2mLYUZYAugDcKjox8VGSllKp6VFZOflFnqUJEgDMAJwAbNWZ2bmFjM3lEmJiiT21fQ0oSEMV3VFpvfWFdMwA9mpzI60Tdf0FcLNxLWIALHtTqxu626Oju4s1+w2MymJ340-LBwDmaCh0J9LisCggAA4AZxorHWNTKkgAHFVvpNQQhlCkTsMsUs0QcqHdziCDgAzIY+EkNMjocRtADsNJo4jIWGgZ1wcAAKjAzqNcODCABhJCQjjrFBkcF0dZgBAAMSQbJA6gFMrlABkEHBfhxmDBWvdcAwYBw6EgEMa7IxXKoAHqJVxGAD0OgAtDNbQBqNabVSuACaZDoQRt9sdRiibqiR3cUR9Nx0geDhhAbw+0f+gLQTqjUQh0NhcEDTuDfFgVFteYxllcUQD8aiAAoAw6AJTOpuWL2kwPtsiQwWwSStAfK6IlSNB3CQgCOpvNlpAADcYBlQGg6DKAO5C9awpCiaCdKRSVpSXCbneKbXQjgATxgkgvICv623Wp1eoAsmgAB6ECEl5bu+RiMGg7CEGgSDimQb7bgAGoQGrKH+zrjNGAHATuQawCh97oXmj6XjB6xGJwe4HnQ9gYL8CAwKSEGQku4EoAgdBoFyzCcGgBrZGmALsWgAAKzDMtAho2IJHEAFrrBKT4SIiNj7ush7oEOIAcIeYDrAEqk6ZuDFMSxcLioekLGawzEqXA5l0JRalHvxjBmU5ACiUIwnCilWLZ9lOKgMAvoworiigRiHoxjD0dAjHWUuCAoLpkKWXFJm4ElKWQqoyU0NqZjPplyXrKlig0KSpKirFkidMV2WieJljSPSYjKSAWWlZCRhiXAj7QD49VdUYKDyXqiCpYpiRDalRjgggoWsEZ5gSNNHUlalXE8Yp9LIhyICkqwNDgvNVCOepaBDvFNkgEdiDkDQVA-kgrAwNYIDoH+50hhBFp0q0jK4CgaTfXgv2xSA-i4Osy7sUt-XXUuMNw2g97fU+0Ow3Q8OeYWPkraML7I9jqPuX+0qAWQxPwxqaQIA9T3MN+L1vQKsIcFy97zYQ4JLYgUrqmkvxWTd0qygqSqEKqIBi5q2q6vqElGjL6oIJz3OwGkHAIL8HGvWq4tIcOAvi7hkMDlFaAxYqypQ1pzB0AgCCKAAEggNC-MwHBgRBNUSPSuCw4w4p0PK+WsFQItLtuzBoBz6zCezUusoJlMAL55Lg26R7A04gNumnS9CKBcjQHCQbAADiSDrFE37rFQCCsFERiPQgUQAEIcXALBCyAadAA}{Link}.
+	\item Guo symmetric side branch \href{https://visualpde.com/sim/?options=N4IgRgqiBcIHYFMCuBbAhnOIA04BqM8y6mO4A6oQCYCWATjQMYAWANggC5lgCK19TNp251CABgB0AVjJUIUWABEA+gEFZeAkuUBJWeUraAyrK7RJYgIy5GCkACVVAOUUAqHcqOvLY12LKMWiCSlgGGIKG4ANaEKqoABPEAvIkWAEyJNHCJANpi2PGWALoA3Co6iSnxkmkAbJnZ8XkFxWWelakSYvXxWbn5haWMHdVdYhm9jc2DJShIIyE9fU0DrXTMAPZqC10AzA39LaVw88mdYgAsBytHJetbFWejYi+W19OtjMpiO+Pvq6UAOZoFDoX4TZYfUoIAAOAGcaKwNo0qjV-rcEMo3k8Qj9JocZlRfpd0TMAGYjHyk1pkdDiCS7ADstJo4jIWGgF1wcAAKjALi9cDDCABhJBwjgbFBkGF0DZgBAAMSQHJA6iFcoVABkEHBARxmDBdi98iAGDAOHQkAhcHQ7IxXKoAHppVxGAD0OgAtHNHQBqe5qVwATTIdCCDudrqM8S98RO7niAc2uhDYfCXx+ceBoLQbtj8VhCKRcBDbrDfFgVEdBcxllc8WDSfiAApgy6AJTuluWP1kkOdshw1UJCox0O4OEARwtVptIAAbjAcqA0HQ5QB3EUbJFIUTQWpSKS7KS4NebxS6hEcACeMEkp5A542G51eoNAFk0AAPQiRJ-ri+RiMGg7CEGgSCSmQz4bgAGoQWrKN+7oZHGv5noBG6hrAiE3ihBZ3mekEbEYnDbrudD2BggIIDAZKgXC84gSgCB0GgPLMJwaBGqazGsWgAAKzCstAxo2CC-EAFobFK94SAAHDYO4bHu6DCrAHB7mAGwBMpWlrnRDFMcikp7nChmsIxSlwKZdDkSp+68SZDkAKLwoiyJyVY1m2U4qAwI+jDipKKBGHu9GMLR0D0ZZ84ICg2lwuZ0VGbg8WJXCqgJTQupmA+aUJRsSWKDQZJkuKUWSLUBUZUJImWNIjJiIpIDpUVcJGMJcB3tAPg1e1RgoDJBqIElclpP1SVGDCCBBawBnmBIE2tYVSUcVxcmMvJUhciAZKsDQMIzVQ9mqWg6kxVZIAHYg5A0FQn5IKwMDWCA6Dfqd4agda9JMiycCfXg31RcEZAbAurHzT1l3zuDkNoDen33rgcN0FDbnFp5i0vI+qPo9+sp-mDENowjWpZAgd0PcwH5PS9QpIhwPI3jNhAwvNiAypqWSAhZV2yvKSoqoQ6ogAL2q6vqhqiSaGqC8zrOwFkHAIICbHPXLCrwbA-iawg2Gg5O4VoJFyqqrrIAGnQCAIIoAASCA0ICzAcMBoGVRIjK4BDjCSnQio5awVB8-OG7MGgTMbAJjMi+yElEwAvkUuAbsHsATiAG7qWqQ40CgPI0BwYGwAA4kgGzxB+GxUAgrDxEY90IPEABCbFwCwPMgAnQA}{Link}.
+	\item Guo pure tip splitting \href{https://visualpde.com/sim/?options=N4IgRgqiBcIHYFMCuBbAhnOIA04BqM8y6mO4A6oYqhlrmAIpXG1lgBOhADAHQCsZACYQosACIB9AIJC8BcRICSQ8pQUBlIQBcYvLgEZcAY1Eg9BskflmeF42pD6yAa0KSpAAg8BeL3oBMXgCWcF4A2lzYHvoAugDckopevh68-gBswaEeEVGxCRLqyX62mR4h4ZHR8UbFqbZcgeXZudVxKEh1vPplFTlV+ewAFgD20l22AMxZlXnxcJ0+JVwALDP9c3HDY0lL9VwH+uut+UYSXBONxwPxAOZoKOiXTX0n8QgADgDOQQA2I9kUvpbNdNggJEc9t0Ls1Zm1BJcuAAOUFtABmdX0MNeNziZHQ3B4kwA7PigtwyFhoCtcHAACowFYHXAfQgAYSQXy0IxQZA+7BGYAQADEkFSQDIWQKhQAZBBwW5aIYwSYHSIgdjk6BadhIBC4dimIwAKikAD1-Mb1AB6RQAWg6poA1NtpMaAJpkdjWE3my1FO0eBbGpIu0ZKD1ehxnC6B+6PNBWjyBz4-f5wD1Wr1MWCCU3Jjzg-TGjzujxOjwACndFoAlNbK-onWiPXWyF9WbBeCt2+LPEkip7cF8AI4wHV63AANxgYVAaHYAoA7myRv8kJxoOk+HxJnxcAvl2J5T8tABPXT8A+LkZLuUKpUAWTQAA9CE5r8v1EY0L8EIQ0CQbkyEPW8AA1CBlCQX2tQJAzfT9b09WAoLPWCCwvA8gJGdQEC0Vd13YAAlDBbn-aA0V-L59RAH8UAQdg0DpIY8LQFV1Tohi0AABSGLVVWMB4uIALRGHlLyRYw1xGDd0E7EAtA3MARksaSlIXGBKN+aipLgbkNy+TSqJoowAX09gCJkzcOLMqyAFFvj+AFLwMXTzIAOVQGB91ozluRQdQN0ooxyK0nSQAQFBlK+QyKOM3BIuir4pCioJ5R0aBeB8xKRhisQgjRNFOXI3h0gSqLcq+XitWBPhiWRcqkvUPi4AvaAsUayr1BQMSlUQGLL38TqYvUD4ECMJBfg0zKeCGiKKpi5jWMvYkkT4GkQDRX4gg+MbBEs2S0E7MKaO2xByCCQQn0mmBDBAdAXwO71fz1QkSTJOAnrwF7yLMMgRinBiprak7cABoG0DPJ7dDBwH2GBhy02cmaDh88H4chuyX35d9-rh4GZRCBALquoZHxu9qWX+LQ6TPMbCA+KbED5aUQluIztJo-lBRFMVCElEBudleVFWVaBVWZQXpQQWn6dgEItAQW5GN+FmeYgrs1aFZC-uHIK0BC0VxS4MglXYBAEDEAAJBAgluIYtG-X8Sp4Ylp3G7l2GFNLfkEDnwqXIY0BpkZuOp-nKSE3GAF8YlwJc-dgIcQCXeSBZ+FA6SCLQ-0IABxJARg8R8RkEBBfg8dRLoQDwACFGLgIwWvZ6OgA}{Link}.
 \end{itemize}

### 2024-08-21 Findings (Shared Folder)/Chapters/Comments.tex
new models added

--- 2024-01-28 Findings (Shared Folder)/Findings.tex
@@
-%\documentclass[12pt,a4paper]{book}
-%\usepackage{commands}
-%
-%\begin{document}
-%
-%	
-%\end{document}
-\documentclass{labdiary}
-\usepackage{commands}
-\usepackage[numbers]{natbib}
-
-\title{My Findings}
-\author{Ali Fele Paranj}
-\date{\today}
-
-\begin{document}
-
-	\maketitle
-
-	\tableofcontents
-
-	\include{Chapters/StatisticalPhysicsForBiology.tex}
-	\include{Chapters/PapersReviewed.tex}
+\chapter{Leah's Chapter}
 
@@
 
-	%\backmatter
-	% Appendices, bibliography, etc.
+\section{Distribution Model for Branching}
 
-	% Bibliography
-	\bibliographystyle{unsrtnat} % The bibliography style
-	\bibliography{References/references} % The filename of the .bib file (without the extension)
+%	
+\subsection{Introduction}
 
-\end{document}
+The purpose of the model below is to derive a continuum description of vessel density so that it will be possible to compute mean vessel radius, as well as other aggregate properties that can, in turn, help to approximate the flow of blood through tissue.
+
+Some questions that this model could help answer are: how do tip branching and fusing combine to determine the properties of the vessels in some average way.
+How is mean radius dependent on assumptions about branching and fusing?
+If drugs or treatments affect the mortality of vessels, or the branching, how would the aggregate properties change?
+
+\subsection{Derivation of the model}
+
+Let $n(r,t), \rho(r,t)$ be the density of tips of blood vessels and of vessels that have radius $r$ at time $t$. That is,
+\[
+\int_a^b n(r,t) \, dr = \text{number of tips of vessels whose radius is in } a \le r \le b
+\]
+and similarly for $\rho$.
+Note that the total number of tips and the total length of vessels in the ``unit volume'' would then be
+\[
+N=\int_0^\infty n(r,t) \, dr, \quad  P=\int_0^\infty \rho(r,t) \, dr
+\]
+and the mean radius of the vessels in the unit volume would be
+\[
+r_\text{mean}=\int_0^\infty r \, \rho(r,t) \, dr
+\]
+We can similarly get higher moments of the distribution. In practice, there is some maximal radius, which is the upper limit of the integral.
+
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.5\linewidth]{Figures/SimpleTubesVesselModel.png}
+%	\caption{The simple model for radial distribution of vessels can provide the most basic first idea of a block of tissue as a porous medium through which blood can flow.
+The model can inform us about the distribution of ``pores'' (radial vessels of various sizes).
+Left: the model keeps track of vessels of various radii in a unit volume of tissue.
+Right: view from another side of the block is basically that of a lot of ``holes'', like in porous medium.
+}
+%	\label{fig:SimpleTubesVesselModel}
+%\end{figure}
+
+We consider the following processes ad rates
+\begin{align*}
+	\beta(r)&= \text{branching rate,}\\
+	K(r,R) &= \text{probability that tip of a vessel of radius } R \text{ produces a daughter of radius } r, \\
+	\phi &= \text {average number of daughters produced per branching event},\\
+	A(r,R)&= \text{probability of fusing of tip of radius }r \text{ with vessel of radius }R,\\
+	\psi(r)&= \text{anastomosis rate},\\
+	v(r)&= \text{rate of tip extension},\\
+	\gamma(r) &= \text{rate of vessel mortality}.    
+\end{align*}
+
+Then a model for $n,\rho$ would be
+\begin{align}
+	\frac{\partial n}{\partial t} &= -\beta(r) n + \phi \beta(r) \int K(r,R) n(R,t) \, dR - \psi n(r,t) \int A(R,r)\rho(R,t)\, dR\\
+	\frac{\partial \rho}{\partial t} &= v(r) n(r,t) - \gamma(r) \rho(r,t)  .
+\end{align}
+
+In the first equation, the terms are (a) loss of a tip of radius $r$ when it splits into daughters of other radii, (b) creation of new tip of radius $r$ from any other parents of radius $R$.
+The integral sums over all such possible parents, with a kernel $K(R,r)$ that weights the contribution from $R$-radius parents to $r$-radius daughters.
+($K$ is assumed to be normalized; more will be said about this kernel later.) A branching event produces 1 or more daughters, in general, and $\phi$ is that average number of daughters.
+($\phi \approx 2$ typically.) (c) The third term is loss of tip of radius $r$ due to fusing with any other branches of radius $R$.
+The kernel $A$ (``anastomosis''), also normalized, gives the probability of this event, and $\psi>0$ is some rate.
+
+In the second equation, density of $r$-radius vessels increases when their tips grow, and we assume a low rate of decay to allow for some steady state.
+The vessels grow in length at their tips (which deposits branch length behind them).
+But it is assumed that they do not widen radially.
+(If this does happen, there would be a term of the form $-\partial (n \omega)/\partial r$ where $\omega$ is the rate of radial growth on the RHS of the first equation to take that into account.)
+
+The above model is nonlocal in that the dynamics of vessels at one radius (say $r$) depend on vessels at far larger (or smaller) radii.
+The kernels in such nonlocal equations will be important in the predicted behaviour of the system.
+However, the model is currently not spatial.
+It could be extended in that was, similar to \cite{edelstein1989models}, but we first consider this nonspatial version.
+For models using nonlocal equations (PDEs with integral terms) we refer the reader to papers on angular distributions such as \cite{edelstein1990contact,edelstein1990models,mogilner1995selecting,mogilner1995modelling,mogilner1999non,edelstein2005mathematical}
+
+In principle, many of the coefficients may be dependent on vessel size.
+That dependence will greatly influence the outcome of the model.
+However, most of the analysis will be restricted to simple special cases where those rates are constant.
+
+\subsection{Possible kernels and their interpretation}
+
+\subsection{Branching kernel}
+
+We can think of several simplifying cases for which the model can be analysed, including:
+
+\begin{itemize}
+	\item[k1:] $K$ is constant. (This is not very realistic, as it implies that a vessel can branch into daughters of any radii, including larger radii, which is implausible.
+	\item[k2:] $K(r,R)$ is a delta-function, 
+	$K=\delta(r-\alpha R)$. This means that parents of radius $R$ produce only daughters of radius $r=\alpha R$, where it makes sense that $0\le \alpha \le 1$.
+	\item[k3:] $K(r,R)$ is some Gaussian with mean at $r=\alpha R$. This last case can be simulated numerically.
+\end{itemize}
+The cases k1, k2 lead to some possible analysis.
+
+\subsection{Branching rate}
+It is possible that the vessel size affects the branching rate: possibly $\beta(r)$ is an increasing or decreasing function of radius.
+We will consider $\beta \ge 0 =$ constant in some cases that can be analysed.
+
+\subsection{Anastomosis kernel}
+
+It is not clear whether fusion of a tip with a vessel depends on the radii of these two. As in the case of the branching kernel, there are several possible assumptions, including
+
+\begin{itemize}
+	\item[a1:] $A$ is constant. (This is not very realistic, as it implies that a vessel of radius $r$ has an equal probability of fusing with a vessel of any other size. 
+	\item[a2:] $A(r,R)$ is a delta-function, 
+	$K=\delta(r-R)$. This means that an $r$-radius tip can only fuse to an $R$-radius vessel if $R=r$ radius.
+	\item[a3:] $A(r,R)$ is some Gaussian with mean at $r=R$. 
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%
+\subsection{Analysis of special cases}
+
+Below, we consider a number of simple situations where, under suitable assumptions, the model can be analysed either qualitatively or quantitatively.
+Even though many simplifications are somewhat unrealistic, we can use these results to get initial insights into how the branching system behaves.
+
+\subsection{Case 1: No anastomosis}
+Consider the case that (1) there is insignificant tip fusions, so that $A\approx 0$ and (2) Branching of $r$-sized tip produces daughters of any size smaller than $r$, with equal probability, that is $\beta=\beta(R)$ depends only on the parent radius, $R$, with $K(R,r)=H(R-r)$ a step function that ``turns on'' when $R=r$.
+(3) There is also some (possibly size-dependent) rate of tip stalling (or death), $\kappa(r)\ge 0$.
+Then the model has the form
+\begin{equation}
+	\label{eq:Case1:noAnast}
+	\frac{\partial n}{\partial t} = -\beta(r) n + \phi  \int_r^\infty \beta(R) n(R,t) \, dR - \kappa(r) n(r,t)     
+\end{equation}
+The first term is as before, while in the second term, we took the branching kernel to be simpler. The last term is 1st order decay of tips.
+
+This model can be analyzed by a method used in section 4 of \cite{edelstein2001model}.Look for a steady state distribution where $\frac{\partial n}{\partial t}=0$, and define
+\[
+z(r)= \int_r^\infty \beta(R) n(R,t) \, dR
+\]
+Then $\phi z$ is a cumulative rate of branching that results in $r$ sized vessels.
+
+By the Fundamental Theorem of Calculus,
+\[
+\frac{dz}{dr}= -\beta(r) n(r)
+\]
+(where we have used an assumption that $n(r)\to 0$ as $r\to \infty$). We can now rewrite the steady state version of Eqn.~\eqref{eq:Case1:noAnast} in the form
+\[
+0=\frac{dz}{dr} + \phi z +\frac{\kappa(r)}{\beta(r)} \frac{dz}{dr}
+\]
+where we have replaced each term by the equivalent term in $z$. But this is just a simple first order ODE for $z(r)$,
+\[
+\frac{dz}{dr}=- \frac{\phi \beta(r)}{\beta(r)+\kappa(r)} \, z.
+\]
+This can be easily integrated, by the usual technique of separation, i.e.,
+\[
+\frac{dz}{z}=- \frac{\phi \beta(r)}{\beta(r)+\kappa(r)} \, dr.
+\]
+It would be interesting to consider this solution for various specific assumptions about the functions $\beta(r),\kappa(r)$.
+We can then get an analytic formula for $n(r)$, and also for the vessel distribution $\rho(r)$.
+For example, if $\beta,\kappa$ are constants, we immediately see an exponential distribution, as expected.
+
+\subsection{Case 2: Delta function kernels}
+
+One special case that can be analyzed is that of delta function kernels $K(r,R)=\delta(r-\alpha R), A(r, R) = \delta(r-R)$. 
+We also assume that all coefficients are constant parameters.
+In this case, the model equations reduce to
+\begin{align}
+	\frac{\partial n}{\partial t} &= -\beta n + \phi \beta n(r/\alpha,t)  - \psi n(r,t) \rho(r,t)\\
+	\frac{\partial \rho}{\partial t} &= v n(r,t) - \gamma \rho(r,t)  .
+\end{align}
+That is, branching or parents with radius $R$ produces daughters at radius $r=\alpha R$, and fusion is only with vessels of the same radius.
+
+The steady state of this system satisfies
+\begin{align}
+	0 &= -\beta n(r) + \phi \beta n(r/\alpha)  - \psi n(r) \rho(r)\\
+	0 &= nv(r) - \gamma \rho(r)  .
+\end{align}
+We can eliminate $\rho$ using the second equation, obtaining
+\begin{equation}
+	\label{eq:discreteEq1}
+	0= -\beta n(r) + \phi \beta n(r/\alpha)  - \frac{v\psi}{\gamma} n^2(r)
+\end{equation}
+This equation links daughter tips of radius $r$ to parent tips whose (larger) radius is $r/\alpha$.
+Let us define the ith generation as the parents and the i+1 generation as the daughters.
+Assign the discrete variables
+\begin{align}
+	n_{i+1}&= n(r)\\
+	n_i&=n(r/\alpha).
+\end{align}
+Then the above equation can be written as
+\begin{equation}
+	\label{eq:discreteEq2}
+	0= -\beta n_{i+1} + \phi \beta n_i - \frac{v\psi}{\gamma} n^2_{i+1}
+\end{equation}
+This is a nonlinear discrete dynamical ``map'', let us rewrite it as
+\[
+n_i=f(n_{i+1}), \quad \text{where} \quad f(x)=\frac{1}{\phi}\left(x+\frac{\psi}{\beta}x^2\right).
+\]
+The ``generator'' of this iterated map is a quadratic function.
+While we cannot solve it in closed form, we can use the usual ``cobwebbing'' method to solve it graphically, as shown in \cite{edelstein2005mathematical}.
+We can also iterate it on a computer.
+
+We can also determine conditions on the parameters for which there is a progression $n_i \to 0$ as the generation number increases, or where $n_i \to N_0$, where $N_0>0$ is some constant tip density.
+
+Note: this map is related to the famous Robert May's discrete logistic map, but our iteration goes in reverse, from i+1 to i.
+That does not matter in the ability to understand the qualitative behaviour of the system.

### 2024-08-21 Findings (Shared Folder)/Chapters/branchingMorphogenesis.tex
new models added

@@
 \begin{itemize}[nosep]
-	\item Fisher-KPP model for branching morphogenesis in VisualPDE: \href{https://visualpde.com/sim/?options=N4IgxiBcIG4JYCc4BM4GcQBoTIKq6hABEscA1MwgBlOQHU7raNoBGKj0sA6G7Aa0JEABAF5hrAHRVhcAHbCA2q0zsAugG4EAfQBGY4VWntZCxVVWa52meIDM000osBWTToCmBo6xdPzlhoANt7SMvLOmAAsmgC2oRz+FjEapLEAhtSSdgDsaXBM2AAOhADCAK5oAC4A9rGkRQg1uh4AYuVyhOUNTS0AMh5yAOZVABZQdhwWIAg8M3oAVOULABSsALQr5QDUMACUAPQr1lR7e8LrwrFLpAiU0J5LwtvCKzq6Byd7S6s7+xdBBYwW6MXikNAlaB2UjAyCKUDpBBNADupRqQRq5QQUAAbC4XHYXNhESiiIM0HAqgBPKBGIkgEk1ZEDYZjACy6QAHoRWKRGciAMpgdJBDyEdLlWp8pFMgAahD62k5BwATBdhNziTLkQBNBXaKmq9U04mSmoCjxVNEYrEAJXSwzFkAAZiK0B5sMLYh4EOkACqjS2ZSCTT3pb2+gAKowKIemXp96QAWjU6rTJAAOT3ozEIDKQ2CIFDoLg5rG6RFQV1Bd3ZuS1LEsau18A1eu56252l1hsIACiRQpGM6kCMVBUrfbWIAcuV6pB6WBKrVYgKsa6wE7mx6QB5Yroamgm26d3uD0eAIL7uCDKrp+lnw9oIhwZ3OypOow47CPo-R2NSC4ORUFmu77k+AoxnINKQOwP7gUeAqxKmYxyB4R7piq8HnmgApFB4S5BJWo6SFhYE4QGQbpjkGYuFE2DOkEcBFPhyCdli+ZVie2BMWhdAoOy5RBFAE4ZJy7F3CK5ROiARi5PkcgSWQUkyXwIA1DAPpETB27YBpWnpFSEndupmkINpA5Dm26YcPS+nmYZfaco0PKkPZ2l9PIHj8cggnCbBxQYlUfpUvhhBFERaE9M08hDFxNY7o0zRtB0XTRf0gwjOMcYcMUvQeCFYXQPIVQeEMvrCXlyXymCVUtHqtUgGg67pJu7QjrJpBjAgHgeEQAASHhwEMoxVEKIqfpIOTYJpYC1AgrQ3kEyDxS2yKjOkwU1JGQVpdgcjhjJvIAL6YAi2rGZAeIEvS-JknIFLUveWooiyWUctybDSii42iuKZrfXK+rKmqlyagy2oNSAiqGqDwgmgyZoWlaZYIPajqrTuCa+pRVTBqG4CHVGMYTPGRPJqm85GKBYCo5x0DwEgqAYNmNoIBW2Iutxk69seCU9h2qMmbTU79oOcDDjZE4i72s7zouy51GuCAblu3O-nzLYa1eug3vWz3kU+L5vh+6bfobf4k7BkhASB2EQVBMFwRbuHITUqHoSwRhkRreEEUJxHe-bR648GRg0XRDFMSxHhsXT6SQrpIC8d5AmjGyQkidgYlKSpWTydn8i50E0mFKZBk6dz7mGZd0zV1SlkS9ZJG2XpZkWc5nMgLybcV55fFpxn-kThF7uFTJEUOmKdWxZjdUpR13Tz29Yyk7lIBJS04+ECVZUVelHg1Z189Q2pzUq61C9l91vUDUNI1jcKf0kdNsAEfNi0eMtc8gOtm1+ttXaDNSAHW9FAFUx01DYGRCtaAS9f4FngRSWIfpKTPxAItNAgYEDrAANKRkjMINkNRkBf2EM6GoCBhAACFfRyDAFBIYRDKFFFGDUIYgx0IlmOkAA}{Link}.
+	\item Modified Fisher-KPP model for branching morphogenesis in VisualPDE (we have taken into account the segment): \href{https://visualpde.com/sim/?options=N4IgxiBcIG4JYCc4BM4GcQBoTIKq6hABEscA1MwgBlOQHU7raNoBGKj0sA6G7Aa0JEABAF5hrAHRVhcAHbCA2q0zsAugG4EAfQBGY4VWntZCxVVWa52meIDM000osBWTToCmBo6xdPzlhoANt7SMvLOmAAsmgC2oRz+FjEapLEAhtSSdgDsaXBM2AAOhADCAK5oAC4A9rGkRQg1uh4AYuVyhOUNTS0AMh5yAOZVABZQdhwWIAg8M3oAVOULABSsALQr5QDUMACUAPQr1lR7e8LrwrFLpAiU0J5LwtvCKzq6Byd7S6s7+xdBBYwW6MXikNAlaB2UjAyCKUDpBBNADupRqQRq5QQUAAbC4XHYXNhESiiIM0HAqgBPKBGIkgEk1ZEDYZjACy6QAHoRWKRGciAMpgdJBDyEdLlWp8pFMgAahD62k5BwATBdhNziTLkQBNBXaKmq9U04mSmoCjxVNEYrEAJXSwzFkAAZiK0B5sMLYh4EOkACqjS2ZSCTT3pb2+gAKowKIemXp96QAWjU6rTJAAOT3ozEIDKQ2CIFDoLg5rG6RFQV1Bd3ZuS1LEsau18A1eu56252l1hsIACiRQpGM6kCMVBUrfbWIAcuV6pB6WBKrVYgKsa6wE7mx6QB5Yroamgm26d3uD0eAIL7uCDKrp+lnw9oIhwZ3OypOow47CPo-R2NSC4ORUFmu77k+AoxnINKQOwP7gUeAqxKmYxyB4R7piq8HnmgApFB4S5BJWo6SFhYE4QGQbpjkGYuFE2DOkEcBFPhyCdli+ZVie2BMWhdAoOy5RBFAE4ZJy7F3CK5ROiARi5PkcgSWQUkyXwIA1DAPpETB27YBpWnpFSEndupmkINpA5Dm26YcPS+nmYZfaco0PKkPZ2l9PIHj8cggnCbBxQYlUfpUvhhBFERaE9M08hDFxNY7o0zRtB0XTRf0gwjOMcYcMUvQeCFYXQPIVQeEMvrCXlyXymCVUtHqtUgGg67pJu7QjrJpBjAgHgeEQAASHhwEMoxVEKIqfpIOTYJpYC1AgrQ3kEyDxS2yKjOkwU1JGQVpdgcjhjJvIAL6YAi2rGZAeIEvS-JknIFLUveWooiyWUctybDSii42iuKZrfXK+rKmqlyagy2oNSAiqGqDwgmgyZoWlaZYIPajqrTuCa+pRVTBqG4CHVGMYTPGRPJqm85GKBYCo5x0DwEgqAYNmNoIBW2Iutxk69seCU9h2qMmbTU79oOcDDjZE4i72s7zouy51GuCAblu3O-nzLYa1eug3vWz3kU+L5vh+6bfobf4k7BkhASB2EQVBMFwRbuHITUqHoSwRhkRreEEUJxHe-bR648GRg0XRDFMSxHhsXT6SQrpIC8d5AmjGyQkidgYlKSpWTydn8i50E0mFKZBk6dz7mGZd0zV1SlkS9ZJG2XpZkWc5nMgLybcV55fFpxn-kThF7uFTJEUOmKdWxZjdUpR13Tz29Yyk7lIBJS04+ECVZUVelHg1Z189Q2pzUq61C9l91vUDUNI1jcKf0kdNsAEfNi0eMtc8gOtm1+ttXaDNSAHW9FAFUx01DYGRCtaAS9f4FngRSWIfpKTPxAItNAgYEDrAANKRkjMINkNRkBf2EM6GoCBhAACFfRyDAFBIYRDKFFFGDUIYgx0IlmOkAA}{Link}.
 	\item Guo pattern formation model for side branching: \href{https://visualpde.com/sim/?options=N4IgRgTiBcIAwDoCsIA0IDGMQDcCWEeAJngM5ohECqV2AIgPoCCFRAam-QwJKsDqfLgGVWAFxiI4ARnQZasAEpMAcnQBU3BkLVS4FDJ1hSKAay5MABAF4LkgEwW8AOwsBtOKikBdANyNu1rYIdgBsji7unr6MQoGSYc5uHt4+GHEIcHAOiZEpALYArulSCRHJvhAAFgD2zOlwAMzhSVE+TkU2kgAszbkVNTz1mY295akMcPXZZa0A5gCGeXnzU6OtAKYADqR4ADbVLp0IABxIaynrDFLFkzljREM9d60AZoG6574Uy9iIDQDs3zwvwoTmwXQom2wAGECqRRNU8pCINUwOsAGIFMGwFjoTYotEAGXWTlmokqMAaw3QEHkmDUTAAenY1EIAPTcAC0hQZAGoqrUmGoAJoUCCGelMlmxTkWdoaCz8gbcEViwSwDATCyyhZLeas7UWLY7fZOEWssUARWwRAZhsuUjUFmFiosAAphcyAJRst1SXkvEXeiikbEgSwBWKi9CkACOMFEEAK63QOBgrlA8wgKIA7tDqvsClBoCEkEgGkh0Fnc3QSTtRABPCTIKvZ6o54mk8kAWXmAA9sMZW7mhBh5rt1th5gUERRq+2ABrYQkMPtshyygfD9ui2Arhvrw1Nqsz6pCdaifOFiAKeakyfQF7j0gpzCLdYQeYAFUqF-mlI8N88g-eYAAVKmBaAqVkd9PwALWqRFm2OWQC2qItlihWB8EIEhyFQ68wCzGAn12F9UKcBEi3IR9n1fDADioiAr3Q4tAIYyjWIAUW2PYDmbaQKKY5QCiRaBK0wOEETyIQiyfDAH1I8iQHWPIwGqUgaKU19VPUzSmDUvASXEaBEAk3SNNIOg8BeF44QfRAQnQCzNPAyCpGQf44BQlS1MsoQIKcJtoF0Zy-M0oQ8kQ8knHWTTmzsMK9NIIRNnWDACl2YjTOCJLLJ-P9m3+U4unQF5djwTY0qIFiMPmLDtPQCrYr4Yge0ymAZBAZY+1q8Vx2TX4EABIEnD6tgBofeAKGqHAPyy4LGpAWb5vmBs+okdAVogBaeJNficsyCTtt2vt8UHGa5p2tbCWcdZWqIdrdk6vF9lEL8GzS7BNiy2LkVRZxZhIui8QJDEsWwXEQHxVF1k7MkKSg6lobBj6vtgZxRHWWZP2e0HYaXWA9HxtFd2mmM5PmBTMTDYmQHJCB1nWOgAAl1jwWZKlEUdxwchB-lTdKEQgdEjN2IhgbI18c0qeZ3uqUC3sh0F3wugBfLx0BzCWcQoHMsPDEM8DyL88FECdsAAcQKaoLG7aoiHWXYLCEYh1gsAAhT8nAwQKgbVoA}{Link}

### 2024-08-25 Findings (Shared Folder)/Chapters/MolecularBiology.tex
new notes added

@@
 \end{figure}
+
+\section{Important Random Facts}
+\begin{itemize}
+	\item The over expression of ANG1 (angiopoeiting1) induces vascular remodeling that leads to the formation of vessels with a wider diameter (\cite{Augustin2009}).
+	\item  TIE2-mediated EC activation controls the expression of endothelial apelin, which in turn acts in an autocrine manner on EC-expressed G-protein-coupled APJ receptors, the downstream signalling of which contributes to the control of vessel diameter (\cite{Augustin2009,Kidoya2008}).
+	\item  Differences in arterio-venous shear stress also control Ang–Tie signalling (\cite{Augustin2009}).
+	\item The quiescent EC phenotype is maintained by constitutive ANG1–TIE2 signalling.
+ANG1 clusters TIE2 junctionally at inter-endothelial cell junctions in trans to transduce survival signals.
+Differences in arterio-venous shear stress also control Ang–Tie signalling.
+During the transition from the quiescent to the activated phenotype, ECs liberate their endogenously stored pools of ANG2, and this antagonizes ANG1–TIE2 signalling to facilitate EC responsiveness to exogenous cytokines.
+As such, the absence or presence of stored ANG2 contributes to the control of the \textbf{adaptive plasticity of the vascular endothelium} (\cite{Augustin2009}).
+	\item  Oscillatory flow has also been measured in humans and reported in \cite{Rodgers1984} (\cite{Carr2005}).
+	\item Observed oscillation frequencies range up to 240 cycles per minute.
+High frequency oscillations (greater than 50 cycles per minute) have been attributed to either heart pulse or breathing rhythms.
+Lower frequency oscillations are thought to be caused by vasomotion.
+Observed frequencies of vasomotion have been reported to range from 2.7 to 32 cycles per minute.6,23 Recent analysis of RBC velocities and arteriole diameter dynamics by Parthimos et al.
+suggests, however, that low frequency oscillations may not be solely due to vasomotion.
+Their analysis demonstrates that two important measures (correlation dimension
+	and Lyapunov exponent) of the RBC velocity oscillations
+	depends on whether vasomotion is present or not. This indicates that something other than vasomotion is also driving
+	the RBC velocity dynamics \cite{Carr2005}
+	\item The responses outlined above lead to continued vessel shrinking and eventually to pruning of vessels that are nonfunctional with respect to both convective and diffusive transport.
+Typically, such vessels have low shear stress and moderate or high oxygen levels and therefore receive a negative net growth stimulus, causing decrease in diameter \cite{Pries2014}.
+	\item The need for distribution of capillaries throughout the tissue implies the presence of both short and long flow pathways connecting the feeding arteriole to the draining venule \cite{Pries2014}.
+	\item If the shrinking of a given vessel leads neither to local hypoxia nor to increased wall shear stress, then it receives no increasing growth stimulus and continues to shrink, eventually being pruned \cite{Pries2014}.
+	\item 
+	\begin{figure}
+		\centering
+		\includegraphics[width=0.5\linewidth]{images/pillarFormationInEmbryoLigation}
+		\caption{Splitting angiogenesis (intussusception) in ligation embryos.
+(A-H) Scanning electron micrographs of vascular corrosion casts of control (A,B) and ligation (C-H) chicken embryos.
+(A) Low magnification of the normal vasculature.
+The boxed region is shown at higher magnification in B.
+Splitting angiogenesis is not apparent.
+(C) Micrograph of ligation embryo showing extensive pillar formation.
+(D) Magnification of the boxed region in C, showing rows of pillars (red arrowheads) delineating the future arteriolar segment (dashed line).
+(E,F) Overview (E) and detail of the boxed area (F) showing fusion of pillars leading to segregation of the capillaries (asterisks).
+(G,H) Another example showing the advanced splitting by pillars (arrowheads); rows of pillars align (dashed line) and subsequent fusion will lead to separation of the feed vessel from the surrounding capillary plexus.
+(I,J) Fluorescent labeling of endothelial cells in vivo shows pillar formation (arrowheads) in distal arterioles and the connected capillary network.
+art, artery.
+Scale bars: 500 μm in A,C; 200 μm in E; 100 μm in B,D,G; 50 μm in F; 20 μm in H; 30 μm in I,J.
+Figure is from \cite{Buschmann2010}}.
+	\end{figure}
+
+\end{itemize}

### 2024-08-25 Findings (Shared Folder)/Chapters/modelAttempts.tex
new notes added

@@
 
+\section{Continues model to capture the distribution of useful quantities}
+After our literature review we observed that the flow of blood through a network of capillaries strongly depends on the radius of the vessel segments, as well as the length of those segments \cite{Koery2024}.
+So we had an attempt to have a mean field approach to capture the density of vasculature with different radius in the environment.
+Before going into the details of this model, consider the following important box:
+\begin{strangeObs}[The assumptions of the model above are not really realistic]
+	When we were developing this model, we knew that there is a notion of radius distribution in the vascular networks, meaning that in a given vascular network we can find vessel segments of different densities \cite{Jafarnejad2019}.
+At that time we thought that this distribution in length is happening because when a leading tip cell divides, with some mechanism, the radius of the daughter vessels are determined from the radius of the parent vessel.
+But this assumption is not correct and is completely wrong.
+Research studies show that the vascular network forms as a branching-annihilating random walker (with or without global guidance and local self interaction), and then, in a process known as ``vascular renormalization'', ``vascular adaptation'', or ``vascular pruning'', the vascular network changes based on the flow in the network, leading to thickening or thinning of some vessels, some branching joining to each other, etc.
+See \cite{Pries2014,Secomb2021,Bakker2021,Maringanti2021} for more details.
+\textbf{However, for pedagogical reasons and to master the technique in this great idea proposed by Leah we will continue analyzing this model.}
+\end{strangeObs}
+
+This model captures the time evolution of two following variables:
+\begin{itemize}
+	\item $ n(r,t) $: The density (number/Area) of tip cells of vessels with radius $ r $ at time $ t $.
+	\item $ \rho(r,t) $: The density (length/Area) of vessels with a given radius $ r $ at time $ t $.
+\end{itemize}
+We assume that the tip cells move in the environment, and as they move they deposit the vessel segments.
+We assume that each vessel segment as a predetermined radius which is determined by the radius of its parent vessel (see the strange observation box above for a discussion about why this is not a correct assumption).
+We also assume that the tip cells are can divide, which will lead to the branching event.
+Also, for tip cells, there is a chance that they encounter another vessel segment along their way, which will lead to the anastomoses event.
+Also note that we assume there is no side branching (despite the fact that there are some very recent evidences that side branching is very important in biological networks \cite{Ucar2023}).
+Two following PDEs captures all of these assumptions.
+
+\begin{align*}
+	\frac{\partial n}{\partial t} &= -\beta(r)n +  m \int K(r,R)  \beta(R) n(R,t) dR - \alpha n\int A(r,R) \rho(R,t) dR - \kappa n \\
+	\frac{\partial \rho}{\partial t} &= v n - \lambda \rho
+\end{align*}
+The parameters in this model are as the following
+\begin{itemize}
+	\item $ \beta(r) $: the branching rate,
+	\item $ m $: the stoichiometry coefficient in $ T \to m T $, where $ T $ represents a tip cell.
+We assume a symmetrical splitting of the parent vessel into the two daughter vessels of the same radius.
+	\item $ \alpha $: anastomoses rate.
+	\item $ v $: the speed at which the tip cells are moving
+	\item $ \lambda $: the death rate of the tip cells.
+	\item $ \kappa $: the death rate (or stalling rate) of the tip cells.
+\end{itemize}
+Furthermore, in the PDEs above we have two non-local terms that the kernels $ K $ and $ A $ plays a major role.
+The kernel $ K(r,R) $ represents the rate at which tip cells of radius $ R $ give birth to the tip cell of radius $ r $, and $ A(r,R) $ represents the rate at which the tip cells of vessels with radius $ r $ do anastomoses with the vessel segments of radius $ R $.
+Here we assume that the tip cells of a vessel with radius $ R $ can do anastomoses with segments with specific radius and this is captured by $ A(r,R) $.
+Different assumptions and hypothesis will lead to different designs of these kernels.
+
+\subsection{Case I: No Anastomoses} 
+For this case we assume that the kernel $ A(r,R) $ is identically zero.
+This can be a realistic assumption when we are considering the growth of vascular networks in 3D, where for several early generations of the angiogenesis, the anastomoses is almost not present (see \cite{Hannezo2017} for the 3D model for the kidney morphogenesis).
+For this case the model reduces to
+\begin{align*}
+	\frac{\partial n}{\partial t} &= -\beta(r)n +  m \int K(r,R) \beta(R) n(R,t) dR - \kappa n \\
+	\frac{\partial \rho}{\partial t} &= v n - \lambda \rho
+\end{align*}
+
+\subsubsection{Sub-case (i): $ K(r,R) = H(R-r) $}
+For this case we assume that $ K(r,R) $ is a step function given by $ H(R-r) $, meaning that only parent vessels of bigger radius can produce the daughter vessels of smaller radius.
+We can re-write the system of PDEs as
+\begin{align*}
+	\frac{\partial n}{\partial t} &= -\beta(r)n +  m \int_{r}^{\infty} \beta(R) n(R,t) dR - \kappa n \\
+	\frac{\partial \rho}{\partial t} &= v n - \lambda \rho
+\end{align*}
+We seek to find an steady state solution of this differential equation, i.e. when $ t\to \infty $ and we have $ \partial n/\partial t  = 0, \partial \rho/\partial t = 0$. Define
+\[ z(r) = \int_{r}^{\infty} \beta(R)n(R)dR, \]
+and from FTC we have $ dz/dr = -\beta(r)n(r) $. The steady state solution will satisfy
+\begin{align*}
+	\frac{d z}{dr} + mz(r) + \frac{\kappa}{\beta(r)}\frac{dz}{dr} = 0 \quad\Leftrightarrow\quad \frac{dz}{dr} = \frac{- m \beta(r)}{\beta(r)+\kappa} z.
+\end{align*}
+Assuming $ \beta(r)\equiv \beta $, and by $ \gamma = -m\beta/(\beta + \kappa) $ we will get
+\[ z(r) = A e^{-\gamma r}. \]
+From the relation $ dz/dr = - \beta(r) n(r) $, as well as $ vn - \lambda \rho = 0 $ we will get
+\[ \boxed{n(r) = \frac{A\gamma}{\beta} e^{-\gamma r}}, \qquad \boxed{\rho(r) = \frac{vA\gamma}{\beta\lambda}e^{-\gamma r}}.\]
+Thus we conclude that under the assumptions as above, the steady state solution of the PDEs will exponential function.
+
+\subsubsection{Sub-case (ii): $ K(r,R) = \delta(r-\eta R) $}
+See (CrossRef here) for a more generalized version of this analysis where the anastomoses kernel is assumed to be a delta function.
+
+\subsection{Case II: Delta function for anastomoses}
+Let's assume that a vessel segment of radius $ r $ can do the anastomoses only with a vessel of radius $ R $. Thus $ A(r,R) = \delta(r-R) $.
+
+\subsubsection{Sub-case (i): $ K(r,R) = \delta(r-\eta R) $}
+In this sub-case we assume that only parents of radius $ R = r/\alpha $ for $ \alpha \leq  1 $ give birth to the vessels of radius $ r $.
+Also, we assume $ \beta(r) = \beta $ is a constant.
+The system of PDEs will be
+\begin{align*}
+	\frac{\partial n}{\partial t} &= -\beta(r)n +  m\beta n(r/\eta,t) - \alpha n\rho(r,t) - \kappa n \\
+	\frac{\partial \rho}{\partial t} &= v n - \lambda \rho
+\end{align*}
+To find the steady state solution of this system, we will have
+\begin{align*}
+	-\beta n(r) +  m\beta n(r/\eta) - \alpha n(r)\rho(r) - \kappa n(r)  = 0, \qquad v n(r) - \lambda \rho(r) = 0.
+\end{align*}
+By substituting $ \rho $ from the second equation in the first equation we will get
+\[ (\beta + \kappa) n(r) + \frac{\alpha v}{\lambda}n^2(r) = m\beta n(r/\eta). \]
+To find the function $ n(r) $ that satisfies the recursive equation above, we can try different guesses (like $ n(r) = Ar^p $) and find the specify the unknowns of the function.
+However, here we do a different kind of analysis instead of trying to find $ n(r) $.
+Let $ n_i = n(r/\eta) $ (the parent generation) and $ n_{i+1} = n(r) $ (the daughter generation).
+Then we will have
+\[ \underbrace{(\beta + \kappa)}_A n_{i+1} + \underbrace{\frac{\alpha v}{\lambda}}_Bn^2_{i+1} = \underbrace{m\beta}_C n_i. \]
+Solving $ n_{i+1} $ in terms of $ n_i $ we will get
+\[ n_{i+1} = \frac{-B \pm \sqrt{B^2 + 4ACn_i}}{2A}. \]
+
 \newpage
-\section{Idea Bin!}
+\section{Ideas!}
 This section might have very simple, basic and sometimes silly ideas that came into my mind during developing some models and I thought they might worth trying

### 2024-08-28 Findings (Shared Folder)/Chapters/MolecularBiology.tex
new references added

@@
 Typically, such vessels have low shear stress and moderate or high oxygen levels and therefore receive a negative net growth stimulus, causing decrease in diameter \cite{Pries2014}.
-	\item The need for distribution of capillaries throughout the tissue implies the presence of both short and long flow pathways connecting the feeding arteriole to the draining venule \cite{Pries2014}.
+	\item The need for distribution of capillaries throughout the tissue implies the presence of both short and long flow pathways connecting the feeding arteriole to the draining venule.
+As illustrated in FIGURE 4, a short pathway has a very high pressure drop per length and thus very high wall shear stress compared with the feeding arteriole from which it branches.
+On the other hand, the local oxygen partial pressure and metabolic environment of the two segments are similar.
+Responses to local signals alone would therefore favor growth in the short pathway, generating a functional arterio-venous (A-V) shunt.
+To avoid such behavior, an additional mechanism is required that signals differences between arterioles supplying a large number of capillaries and those forming short A-V connections.
+This mechanism must provide transfer of information upstream along arterioles within vascular networks.
+A similar consideration applies to vessels in the venular network, where information transfer in the downstream direction, from distal to proximal vessels, is necessary.
+In this case, convective transport of metabolites may provide the needed signals (55, 63).
+However, upstream information transfer is not so simply explained \cite{Pries2014}.
 	\item If the shrinking of a given vessel leads neither to local hypoxia nor to increased wall shear stress, then it receives no increasing growth stimulus and continues to shrink, eventually being pruned \cite{Pries2014}.
+	\item Conversely, shrinkage of a vessel that is needed for diffusive transport leads to hypoxia and an increased metabolic stimulus for growth, whereas shrinkage of a vessel that is needed for convective transport leads to increased wall shear stress, also a stimulus for vessel growth.
+In each case, the resulting negative feedback loop stabilizes vessel diameter \cite{Pries2014}.
+	\item Third, angiogenesis, remodeling, and pruning occur in parallel and not as separate processes \cite{Pries2014}.
+	\item An interesting prediction of the theory is that this system exhibits hysteresis: the increase in vascular density generated during a period of higher demand is only partially reversed if demand returns to its former level \cite{Pries2014}.
+	\item Under steady-state conditions, endothelial cells are relatively quiescent, exhibiting a turnover time in the range of ∼30–300 days (see reference 18,35 of \cite{Pries2014}).
+	\item In tumors, the microvasculature is typically seen to be more tortuous and disorganized than in normal tissues.
+Tumors often show a relatively high proportion of hypoxic tissue, even if the vascular volume and perfusion are relatively high.
+This hypoxia has important effects on tumor responses to radiation and chemotherapies, generally reducing their effectiveness.
+Analysis of hemodynamics and remodeling in tumor microvessel networks suggests that poor oxygenation may result from functional shunting, i.e., failure to distribute flow appropriately between short and long flow pathways, as a consequence of impaired conducted responses (55, 56).
+Such impairment is plausible given that endogenous VEGF levels are typically elevated in tumor tissues and that VEGF has disruptive effects on vascular wall integrity and gap junction function (74) \cite{Pries2014}.
+	\item The dense but highly disordered and functionally deficient vascular networks often observed in tumors can be interpreted as the result of excessive angiogenesis combined with weak or defective remodeling and pruning.
+Similar patterns with increases of vessel density but not perfusion have been reported to result from VEGF overexpression (78) \cite{Pries2014}.
+	\item Therefore, understanding the formation of
+	vascular networks requires consideration of the integrated processes of angiogenesis, structural adaptation, and pruning \cite{Pries2014}.
+	\item These structural adaptations may involve alterations in the dimensions and wall composition of individual vessel segments (remodelling), the growth of new segments (angiogenesis) and the loss of existing segments (pruning)\cite{Secomb2012}.
+	\item  in the development of hypertension, inward remodelling accompanied by wall thickening is a typical structural feature \cite{Secomb2012}.
+	\item Some experimental references for the vascular remodeling and adaptation: Direct experimental approaches have provided much information about structural remodelling of blood vessels.
+Ex vivo perfusion systems allow manipulation of haemodynamic conditions and observation of resulting structural changes over periods of hours to days [3–5].
+In vivo animal models can be used to observe structural responses to surgical alterations in flow conditions over periods of days to weeks [6–10].
+Such experimental approaches have been used to investigate the responses of individual vessels (generally arteries) to mechanical stimuli, including fluid shear stress acting on the endothelial cell layer and circumferential and axial stresses acting on the wall, and to metabolic and pharmacological stimuli \cite{Secomb2012}.
+	\item Early theories of structural adaptation assumed that the diameter of each segment is controlled so as to achieve a target level of the wall shear stress [18,19] \cite{Secomb2012}.
+	\item However, analyses of microvascular networks in the rat mesentery [22] showed a systematic increase in wall shear stress with intravascular pressure from the venules to the arterioles.
+Such behaviour was incorporated in the model by assuming a pressure-dependent set point for wall shear stress.
+This implies that venous vessels are larger than corresponding arterial vessels carrying the same blood flow.
+Another consequence is that the pressure drop is much larger in the arterioles than in the venules, and capillary pressure is much lower than arterial pressure \cite{Secomb2012}.
+	\item Analysis of the model under the assumption that diameters respond only to the haemodynamic signals of pressure and wall shear stress shows that parallel flow pathways are unstable (fig.
+2A) [19].
+Furthermore, this assumption neglects the obvious need for network structure to respond to metabolic needs.
+Introduction of a signal dependent on local oxygen level (fig.
+1B) provides a metabolic response and can be shown to stabilize parallel pathways (fig.
+2B).
+\cite{Secomb2012}
+	\item Mechanical parameters such as tissue elasticity, viscosity and friction can also specify time and length scales of morphogenetic processes (Fig.
+1a).
+For instance, the length scale of stress propagation, the so-called hydrodynamic length, depends on the relative contribution of viscosity and friction within a cell [30] or a tissue.
+Viscosity can also define rates of deformation upon a given mechanical stress.
+The ratio between the viscous modulus and elastic modulus defines the Maxwell time, that is, the time above which deformations become irreversible, typical of a viscous response[31].
+Mechanics can thus direct morphogenesis in a manner similar to biochemical information.
+For instance, dissipation of a localized stress by friction can generate gradients of stress similar to those better known for biochemical gradients of morphogens [32,33] \cite{Collinet2021}
+	\item We delineated two idealized and distinct modalities of information flow during morphogenesis.
+Programmes specify deterministically and hierarchically all operations required for the development of a shape.
+Programmed morphogenesis results explicitly and predictably from the spatially organized initial conditions, a pre-pattern, and relies upon deterministic rules.
+For instance, a genetically encoded morphogen gradient specifies a battery of downstream decisions that themselves dictate mechanical states in cells such as cell contractility.
+By contrast, self-organization is characterized by the emergence of ordered structures from a purely homogeneous initial state.
+It relies on stochastic rules, local activity associated with molecular motors and dissipation of elastic energy, driving irreversible deformation and amplification of local activity via feedbacks that operate across scales.
+Thereby, the system transits to a steady state that minimizes its free energy \cite{Collinet2021}.
+	\item Programmes are usually hard-wired and exhibit redundancy, such that they are mostly insensitive to genetic perturbations.
+However, once affected, for example by a mutation, they cannot be repaired to generate the expected shape because of the absence of feedbacks and a strict dependency on initial conditions, which may be lost as morphogenesis proceeds.
+By contrast, owing to their internal feedbacks, rapid dynamics and insensitivity to initial conditions, self-organized systems can reform after complete perturbations and constantly adapt to a changing environment.
+Thus, programmes may be most suited to robustly guide a few critical steps of morphogenesis where failing to properly time or position singular shape changes of the tissue may affect the entire subsequent steps of the morphogenesis, such as during embryo gastrulation or the specification of primary sulci in the developing brain cortex \cite{Collinet2021}.
+	\item The energy minimisation formalism used here is consistent with Murray’s law.
+In Murray’s approach, it is assumed that there are two contributions to the energy cost of maintaining the flow: the power required to overcome viscous drag forces and the metabolic cost of maintaining the network structure, which is proportional to its volume \cite{Almeida2022}.
+	\item Animals and plants may benefit from loop structures in many ways.
+For example, loops are important in mitigating damages of networks [3] and optimizing energy consumption with fluctuating flow distributions [3,4]\cite{Hu2013}.
+	\item biological transport networks are thought to have undergone a process of gradual optimization through evolution [1], culminating in organizational principles such as Murray’s law [2–4].
+A particular class of such networks that minimizes flow resistance under biologically relevant constraints has been studied to reveal a wealth of phenomena such as phase transitions [5,6], the interdependence of flow and conduit geometry [7], and predictions about allometric scaling relations in biology [8].
+When the optimization models are generalized to require resilience to damage or to consider fluctuations in the load, optimal networks reproduce the reticulate network patterns observed in biological systems [9,10].
+The optimality principles that often determine these networks also appear in nonbiological context, e.g., river basins [11,12], and are relevant for man-made systems such as gas or sewage pipe networks [13,14] \cite{Ronellenfitsch2016}.
+	\item The topology of such efficient networks is characterized by the tendency to reuse the same edge to supply large parts of the network, as opposed to directly connecting each node to the source.
+This is reflected in the mean number of edges between two bifurcations (the mean branch length).
+Efficient networks tend to exhibit fewer nonbranching nodes [Fig.
+3(c)].
+Temporally fluctuating sources (similar to [27]) during the adaptive process can produce loops [33], reminiscent of real reticulate biological networks.
+In addition, variable branching angles [7], growth anisotropies, and steric effects [33] may also play a role \cite{Ronellenfitsch2016}.
+	\item Growth effectively reduces the dimension of the evolutionary search space to two parameters that can be used to explore the energy landscape \cite{Ronellenfitsch2016}.
 	\item 

### 2024-08-30 Findings (Shared Folder)/Chapters/MolecularBiology.tex
new notes added

@@
 	\item Growth effectively reduces the dimension of the evolutionary search space to two parameters that can be used to explore the energy landscape \cite{Ronellenfitsch2016}.
+	\item Natural and man-made transport webs are frequently dominated by dense sets of nested cycles.
+The architecture of these networks, as defined by the topology and edge weights, determines how efficiently the networks perform their function.
+\cite{Modes2016}
+	\item For many purposes, e.g., when analyzing blood flow in arteries, the dependence of viscosity on shear rate can be neglected and blood can be treated as a Newtonian fluid as defined above \cite{Secomb2021}.
+	 the apparent viscosity in large vessels matches the in vitro result, but the apparent viscosity is substantially higher than the in vitro estimate in smaller vessels.
+For example, the estimated apparent viscosity in a $7\mu m$ capillary at hematocrit $45$ percent is about $8.4\mu_p$, almost seven times higher than would be expected based on data from glass tubes.
 	\item 

### 2024-09-04 Findings (Shared Folder)/Chapters/MolecularBiology.tex
new notes added

@@
 
-\section{Important Random Facts}
+\section{Important Facts}
 \begin{itemize}
@@
 	 the apparent viscosity in large vessels matches the in vitro result, but the apparent viscosity is substantially higher than the in vitro estimate in smaller vessels.
-For example, the estimated apparent viscosity in a $7\mu m$ capillary at hematocrit $45$ percent is about $8.4\mu_p$, almost seven times higher than would be expected based on data from glass tubes.
+For example, the estimated apparent viscosity in a $7\mu m$ capillary at hematocrit $45\%$ is about $8.4\mu_p$, almost seven times higher than would be expected based on data from glass tubes.
+	 \item An important implication of the existence of the endothelial surface layer is that the endothelial cell membrane does not experience a significant level of fluid shear stress due to plasma flow over its surface.
+Instead, shear stress is transmitted to the endothelial cells via the attachment points of the macromolecules that anchor the ESL, including syndecans and glypicans [23, 24].
+Being transmembrane proteins, these molecules may in turn transmit the shear stress to the internal cytoskeleton, consistent with an important role for the cytoskeleton in mechanotransduction of shear stress [25, 26] \cite{Secomb2021}.
+	 \item However, when the flow of blood is considered, additional complexity arises because blood is a concentrated suspension of cells, whose dimensions are comparable to microvessel diameters.
+As mentioned earlier, red blood cells show a tendency to migrate away from microvessel walls.
+The fluid mechanical mechanisms underlying this behavior depend on the deformability of the cells and on the fluid dynamical interactions of cells with the walls and are only partially understood [18, 28,29,30,31,32,33].
+The distribution of red blood cells across the vessel cross-section affects the distribution of hematocrit in each branch when the flow reaches a diverging microvascular bifurcation [34], and the theoretical understanding of this phenomenon is a challenging problem in the low Reynolds number flow [35,36,37] \cite{Secomb2021}.
+	 \item Transport of oxygen is a critical function of the circulation. Being a non-polar molecule, it has relatively low solubility in water and in tissue \cite{Secomb2021}.
+	 \item Therefore, it must be delivered by convective transport within such a distance of all cells that require oxygen.
+This requires a dense network of tiny vessels throughout the tissue.
+On the other hand, viscous resistance to blood flow is very high in vessels with very small diameters.
+For mechanical efficiency, the vascular system must include hierarchical branching structures feeding the microvessels, such that convective transport over larger distances can be achieved by vessels with larger diameters and lower resistance to flow.
+In effect, the vascular system must solve a complex patterning problem, generating a structure in which a dense meshwork of capillaries is combined with a hierarchical structure of arteries, arterioles, venules, and veins of varying lengths and diameters so that all parts of the tissues are adequately supplied with blood flow [40] \cite{Secomb2021}.
+	 \item A quick anatomy: Blood vessels are mainly formed by endothelial cells (ECs) and mural cells (MCs).
+ECs are responsible for the exchange and barrier functions of the vascular network.
+MCs, which are in close contact with the ECs, contribute to vessel stability.
+Vessels with a large caliber (arteries and veins) are sheltered by vascular smooth muscle cells (vSMCs) while capillaries are covered by pericytes (PCs) (Potente and Mäkinen, 2017; van Dijk et al., 2015) \cite{Ouarne2021}.
+	 \item lood vessels develop by distinct mechanisms (Semenza, 2007).
+First, vasculogenesis forms the primary vascular plexus by a de novo production of ECs from the mesoderm (Kolte et al., 2016; Potente and Mäkinen, 2017).
+Then, vascular networks grow through angiogenesis, a process responsible for the formation of new blood vessels from pre-existing ones (Potente and Mäkinen, 2017) \cite{Ouarne2021}.
+	 \item Angiogenesis can occur by two different mechanisms: i) sprouting angiogenesis, characterized by the expansion of pre-existing vascular networks into a web of interconnected capillaries; and ii) intussusceptive angiogenesis, characterized by the formation of an intussusceptive pillar, splitting and remodeling of blood vessels \cite{Ouarne2021}.
+	 \item Vessel pruning plays a crucial role in optimizing blood flow within immature vascular networks, contributing to their hierarchical organization.
+This process involves the selective removal of superfluous or non-perfused vessels.
+Unlike vascular involution, vessel pruning does not rely on endothelial cell (EC) death.
+Instead, it depends on the migration and rearrangement of ECs in response to blood flow (Fonseca et al., 2020; Potente et al., 2011).
+In fact, only 5 to 15\% of vessels undergoing pruning contain apoptotic ECs (Franco et al., 2015; Y.
+Zhang et al., 2018b) \cite{Ouarne2021}.
+	 \item This is a dynamic process of structural adaptation and flow regulation that continually adjusts the vessel caliber to compensate for heterogeneity, leading to a flow redistribution according to metabolic needs to ensure adequate tissue oxygenation (Roy and Secomb, 2020) \cite{Ouarne2021}.
+	 \item Vascular quiescence stabilizes the vascular network by halting its development, allowing endothelial cells (ECs) to specialize and mural cells (MCs) to mature into their specific organ functions.
+During quiescence, ECs stop proliferating and migrating (Ricard et al., 2021), creating a balanced state supported by mature MCs.
+However, this equilibrium is not permanent; quiescence can be disrupted during various physiological processes, such as pregnancy and the menstrual cycle (Carmeliet, 2005; Ricard et al., 2021), as well as in pathological conditions (see Section 6) \cite{Ouarne2021}.
+	 \item where $ \rho $ is the “resistivity”, which is supposed to be the same for all the pipes.
+For $ m = 1 $, the flow in each channel is pluglike (e.g., electric current in metallic wire, liquid flow in porous conduct,…), while for $ m=2 $, the flow is Poiseuille-like (e.g., laminar viscous flow in pipe)\cite{Durand2006}.
+	 \item Therefore, it may be of interest to compare the structure of some natural networks with the results presented in this work.
+Indeed, it has been already shown in various publications [22,23] that Murray’s law is well satisfied in some appropriate portions of human and animal vascular systems.
+In that case, the flow profile is nearly Poiseuille-like ($ m=2 $) and the relevant constraint is a fixed total channel volume ($ n=1 $) [22,23] \cite{Durand2006}.
 	\item 

### 2024-09-04 MeetingSummary/BanffMeeting/document.tex
new notes added

\documentclass[12pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage[left=2cm, right=2cm, top=2cm, bottom=2cm]{geometry}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage{nameref}
\usepackage{enumerate}
\usepackage{placeins}
\usepackage{hyperref}
\title{Formation of Looping Networks - from Nature to Models}
\author{Ali Fele Paranj}

\begin{document}
	\maketitle
	\begin{abstract}
		This is my summary of the international meeting on the ``Formation of Looping Networks - From Nature to Models''.
	\end{abstract}

	\section{Introduction: Annemiek Cornelissen, CNRS \& Université de Paris}
	\begin{itemize}
		\item Loops are not stable by design. There should be lots of background mechanisms in action on order to make them stable. [my note: This can also be seen in the papers by Secomb, where he argues that loop is stable only at special conditions].
		\item Mechanisms for the emergence of the loops: 1) Branching along with anastomoses, 2) Structuring from a ``foam''.
	\end{itemize}

	\section{The scale of river networks: Olivier Devauchelle, Université de Paris}
	\begin{itemize}
		\item River networks are sparse in loops. However, rivers forming in the deltas are usually  very loopy structures. 
		\item Scallop theorem, first discussed by Purcell in the paper ``Life at low Reynolds number''
	\end{itemize}

	\begin{figure}[h!]
		\centering
		\includegraphics[width=0.7\linewidth]{OlivierTalk.pdf}
		\caption{(A) The basin of the Loire river, France. The low number of loops in the network is noticeable. (B) Niger delta, Nigeria. The are plenty of loops in the river structure. (C) Art piece created by Janaina Mello Landini. (D) A slide of the presentation. Note the relation between the daughter and the parent diameter relation and the real data supporting that relation}
	\end{figure}
	\FloatBarrier

	\section{River loops morphodynamics: integrating theory with remote sensing observations :Niccolò Ragno, University of Trento}
	\begin{itemize}
		\item Confluence for the rivers is the same as anastomoses for the vascular networks.
		\item Some main questions. 1) Are the loop in the rivers form by pure chance or it is a result of a selective process? 2) Do loops have a characteristic length scale?
	\end{itemize}

	\begin{figure}[h!]
		\centering
		\includegraphics[width=0.7\linewidth]{NiccoloTalk.png}
		\caption{(A) Some beautiful images of loopy networks of rivers. }
	\end{figure}

	\FloatBarrier

	\section{Tissue geometry and lung branching mode selection: Sharon Lubkin, North Carolina State University}
	\begin{itemize}
		\item A network of networks is a multi-scale network!
	\end{itemize}

	\section{From understanding road networks patterns to modeling their evolution: Claire Lagesse, University of Franche-Comté} 
	\begin{itemize}
		\item Take a look at the ``WayMprph'' software.
	\end{itemize}

	\begin{figure}[h!]
		\centering
		\includegraphics[width=0.7\linewidth]{ClaireTalk.png}
		\caption{(A) Different morphology of cities. (B) Transition of the morphology of the road networks as a function of space. (C) Change of the road network morphology in Calgary. The right part of the graph the road system is more uniform as it is located on a flat ground. However, on the left, the road network is more irregular as it is located in the mountain side of the city.}
	\end{figure}

	\FloatBarrier

	\section{Developmental Dynamics of Complex, Space-Sharing Networks: To Ensnarl or Not To Ensnarl?Carl Modes, Max Planck Institute of Molecular Cell Biology and Genetics (MPI-CBG)}

\end{document}

### 2024-09-07 ReviewPaper_Angiogenesis/main.tex
new notes added

\documentclass[10pt,a4paper,twocolumn]{article}
\usepackage[T1]{fontenc}
\usepackage[left=2cm, right=2cm, top=2cm, bottom=2cm]{geometry}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{thmtools}
\usepackage{xcolor}
\usepackage{enumerate}
\usepackage{nameref}
\usepackage{hyperref}
\title{On the Mathematical Formulation of the Angiogenesis Process in Health and Disease}
\author{Ali Fele-Paranj, Leah Edelstein-Keshet}

\newcommand{\R}{\mathbb{R}}

\begin{document}
	\maketitle
	\begin{abstract}
		We will review the mathematical formulations of the angiogenesis process in health and disease. The implementations of some of the discussed models are shared, as well as sharing some mini-scripts demonstrating use cases of some of the software tools we have described in the paper.
	\end{abstract}
	\section{Introduction}
	Angiogenesis is the biological process in which body generates new vessels out of the pre-existing ones that happens during the development (some citations) as well as the response of body for the change in the demands in the tissue (like in sports, wound healing, or angiogenesis in the case of tumor). Angiogenesis is a complected process because as we will discuss there are many molecular, biochemical, genetic, mechanical, topological (the gap junction signal transduction) as well as statistical mechanisms that are happening in parallel and each of them have a significant effect on the outcome of the network.

	Also we need to distinguish the notion of angiogenesis and angiogenesis. During the development, much of the information about the important and main vessels are pre-determined during the development. However afterwards, since there are $10^9$ vessel segment in the body (citation), it is not possible to code all of that information in a genetic code. So an optimized and working vascular system emerges as a self-organized way.

	Angiogenesis is a simultaneous process of growth a a new network (by sprouting or intussusception) and remodeling at the same time. These two processes happen at the same time, but most of the recent and old papers modeling the angiogenesis are either modeling sprouting, or modeling the vasculature remodeling. Or they assume that there is a pre-existing vasculature mesh and then they apply the remodeling principles on top of that mesh. Neither of these assumptions are realistic and in fact in the paper [Secomb, making vascular network work], the authors show some evidences that the sprouting and remodeling both happen at the same time.

	In this review paper, we will go through the mathematical models capturing the sprouting step in angiogenesis, and then we will discuss the literature modeling the vasculature-remodeling (maturation). We will discuss the classic as well as the recent results and we will conclude with a discussion on the possible improvements on these models.

	\section{Sprouting Angiogenesis}
	\subsection{Biological Mechanisms of Angiogenesis}
	Here we will review some of the important biological aspect of the sprouting in angiogenesis. This will not be a comprehensive review and we refer the readers to other detailed review papers [cited here]. Instead, here we will only cover the details that are highlighted to taken into account in mathematical models presented below.

	\subsection{Anderson-Chaplain Model}
	Is one of the classic models that captures that sprouting step in the angiogenesis and takes into account some important factors like the fibronectin and the tumor angiogenic factor (TAF). This model of angiogenesis has been appeared in many different papers and many different analysis has been done on the topology and statistical features of the formed network.

	\begin{align*}
		&\frac{\partial n}{\partial t} =  D_n\nabla^2 n  - \nabla\cdot(\chi n\nabla c) - \nabla\cdot(\rho n \nabla f), \\
		&\frac{\partial c}{\partial t} = -\lambda n c, \\
		&\frac{\partial f}{\partial t} = \omega n - \mu n f,
	\end{align*}

	\begin{itemize}
		\item $ n = n(X,t): \Omega \times \R \to \R $: the endothelial-cell density (per unit area).
		\item $ c = c(X,t): \Omega \times \R \to \R $: the tumor angiogenic factor (TAF) concentration (nmol per unit area).
		\item $ f = f(X,t): \Omega \times \R \to \R $: the fibronectin concentration (nmol per unit area).
	\end{itemize}

	\subsection{Branching Annihilating Random Walker}
	The Anderson-Chaplain model of angiogenesis is also a form of branching annihilating random walker at its heart. A powerful fact about the BARW formulation of the sprouting mechanism in angiogenesis is that we can have some important analytical predictions about the properties of the network formed by this process.

	The paper A unifying theory of branching annihilating random walker is has demonstrated the capabilities of this model in explaining the observed morphoges in a variety of tissues, as diverse as kidney, mammary glands, and prostate. Also the paper by Ucar 2024 shows that BARW can model the lymphatic network and the authors have found that the global optimization and local adaptation ensues the optimal coverage of vascular tissue.

	\section{Microcirculation, Vascular Remodeling and Adaptation}
	Vascular networks are complex networks that has the ability to adapt to the changing conditions and metabolic demand in tissue. The sprouting step generates a mesh of vasculature in tissue which then adapts according to the signals. In this section, we will review some of the most important biological facts behind this mechanisms and then we will cover the classic and recent mathematical models modeling this phenomena.

	Most of the biology behind the vascular remodeling comes back to the principles of micro circulation in vascular networks. So in this section we will cover the basics of the micro-circulation in vascular networks and we will discuss some of the important non-linearities in the rheology of the blood in microvasculature.

	\subsection{Microcirculation in blood microvasculature}
	The blood is a suspension of red blood cells and proteins in a plasma and this makes vascular networks to have a very interesting non-linear rheology. Three main effects are 
	\begin{enumerate}
		\item Hematocrit splitting effect,
		\item Fahraeus effect,
		\item Fahraeus-Lindqvist effect.
	\end{enumerate}
	Hematocrit splitting effect is basically the fact that at the branching points of the vascular networks,  Faraeus and Fahraeus-Lindqvist effects can be traced down to the plasma skimming effect, where a think layer of void space forms between the red blood cells and the wall of the blood vessels. This causes the ratio between the tube hamtocrite\footnote{Hematocrit is the volumetric ratio of red blood cells in a given sample of blood} and the discharge hematocrite to depend on the tube (which is called the Fahraues effect), as well as the dependence of the relative viscosity of blood (the apparent viscousity of blood\footnote{we can measure the apparent viscousity of blood by considering it as a fluid and running it through a channel (a glass tube) and then usng the relation [blublu] to measure the value for viscosity.} divided by the viscosity of plasma) changes as a function of the tube diameters and retains its minimum value for the tube diameter $ 10\mu m $ that coincides with the typical diameter of the capillary bends.

	 It is very important to mention that all of these experiments where done in a glass tube, and the situation in in-vivo systems are more subtle. There has been discrepency between the tube measurements and the real in-vivo vascular networks. Different mechanisms has been suggested to take into account this (cite to that old paper), but then it turned out the the inner layer of the vessels causes this discrepancy.

	\subsection{Biological Process of the Vascular Network Remodeling}
	The endothelial cells can feel the shear stress and respond to that (through that trans-membrane protein that can act as a mechanical sensor). The higher the shear stress is the larger the vessel gets. However, by an argument first appeared in (some of the Secomb papers), if this was true, then the parallel tracks would be unstable. That is because for the perfectly symmetric parallel track (or loop), disturbing the radius of the any of the parallel tracks can lead to the complete shrinkage of one track and the growth of the other track. To make this not happen, we need to also have metabolic signaling, where tissue under hypoxia initiates a signaling that promotes the growth of the vessel diameter. However, this mechanisms alone can lead to the problem of the functional shunt where the long pathways of the vascular network can get closed and a shunt short pathway can form. So there should be some downstream and upstream signaling happening in the network that is really related to the topology of the network (for instance, length of the downstream daughter branches that a particular branch has) which primarily conducted by the gap-junction receptors between the endothelial cells. This gap junction is one of the most volunreble things in the case of cancer and more of this is covered in the disease section.

	\section{Mathematical Formulation of the Vascular Network Remodeling}
	The problem of the vascular network modeling can be formulated in many different ways and here we will be discussing some of the recent approaches.

	\subsection{Discrete Formulation}
	In this formulation we start with the fact that we already have a mesh of vascular network that can be abstracted as a weighted graph $ G(E,V) $ where $ E $ is the edge set and $ V $ is the vertex set, where the weights on the graph if $ C: E \to \R $ which shows the conductance of each edge. We can also assume that the flow in the network is governed by some potentials defined on the vertices $ P: V \to \R $. The flow at each edge is given by the function $ Q: E \to \R $. We can assume $ Q_{i,j} = F(\Delta P_{i,j},C_{i,j}) $, where different choices for $ F $ will assume different fluids running in the network. For instance, for the case of Poiseuille's flow we have
	\[ Q_{i,j} = -C_{i,j} \frac{\Delta P_{i,j}}{L_{i,j}} \]
	where $ L_{i,j} $ is the length of the edge.

	\subsection{Continuous Formulation}
	In the continuous formulation, instead of the discrete graph as our network, we assume the domain is a continuous domain (for instance $ \Omega \subset \R^3 $). Then we defined the conductance vector field over the domain $ m: \Omega \to \R^3 $ and define the permeability tensor as
	\[ \mathbb{P}[m] = \mathbb{I}r + m\otimes m, \]
	where $ \mathbb{I} $ is identity matrix, $ m\otimes m $ is the tensor product of two vectors\footnote{The tensor product of two vectors $ a=(a_1,a_2) $ and $ b = (b_1,b_2) $ is a $ 2\times 2 $ matrix given as 
	\[ a\otimes b = 
	\begin{pmatrix}
		a_1 b_1 & a_1 b_2 \\
		a_2 b_1 & a_2 b_2
	\end{pmatrix}.
	 \]
	}

	\subsubsection{Well-posedness of the Continuous Formulation}

	\subsection{Time Evolution of the Graph}
	We can come up with all sort of laws and rules that can govern the time evolution of the vascular network. For instance, we can assume that the vascular network evolves as the gradient flow of some energy functional, i.e. there is some functional that the vascular network wants to minimize. Or we can come up with some ad-hoc ODE governing the weights of the network that evolve in a mechanistic way with some biological intuition. Or, we can even try to find this evolution law from data by trying to find a group of diffeomorphisms which are on a geodesics in the space of all diffeomorphisms.

	\subsubsection{Time Evolution as the Gradient Flow of The Energy Functional}
	We can postulate that a normal vascular network remodeling in a normal environment is maximizing some functional $ E $ that depends on the weights of the edges of the graphs. There have been many different suggestions for this functional and one of the most popular ones for the discrete model
	\[ E(C_{i,j}) = \frac12 \sum_{<i,j>}(\frac{Q_{i,j}^2}{C_{i,j}} + \nu C_{i,j}^\gamma)  \]
	with some other variations. In this energy functional the first term takes the pumping energy into account and the second functional takes the metabolic constrain into account. In general we can assume that this function will have three kind of terms: 1) A term about the cost of pumping the fluid through the network, 2) a term quantifying the total volume or the total surface area of the network (as larger networks will need more maintainance cost), and 3) robustness to damage or fluctuating source/load.
	and now we can assume that the weights on the graph changes in a way that this energy minimizes, i.e.
	\[ \frac{d C_{i,j}}{dt} = - \nabla E, \]
	so the energy will be minimized. The steady state solution of this ODE systems is precisely the Euler-Lagrange formulation of the least action principle.

	\subsection{Time Evolution Governed by Mechanistic Ad-Hoc Models}
	We can also describe the time evolution of the conductance in an ad-hoc way. The following ODE has been popular form some specific applications
	\begin{align*}
		&\frac{1}{C_{i,j}}\frac{dC_{i,j}}{dt} = k (\tau_\text{wall} - \tau_0),\\
		&\frac{dC_i,j}{dt} = \alpha |Q_{i,j}|^{2\sigma} - b C_{i,j} + g, \\
		&\frac{d}{dt}\sqrt{C_{i,j}} = F(Q_{i,j}) - c \sqrt{C_{i,j}}.
	\end{align*}

	\section{Data Analysis}
	The vascular networks are very complex networks that their structure carries a lot of important information about their function. There have been many different important measurements/characterization of vascular networks like measuring the distribution of the length of the vascular branches, the distribution of the angles between branches, the distribution of the radius of the vessel segments, and etc. Also there have been some other quantification of the vascular networks like measuring the topological features which we will cover here.

	\subsection{Topological Features of Vascular Networks}
	There has been an active research on the correlating the topological characteristics of vascular network in health and disease and if these measurements can cluster these data points efficiently. Here we will quickly review the important topological features of vascular networks.

	\section{Statistical Features of Vascular Networks}
	Scaling laws, statistical physics, and statistical physics of complex networks.

	\section{Angiogenesis in Health and Disease}
	One important aim of modeling is to understand/predict what happens to a vascular network in disease. Vascular networks get affected in a number of disease, including hypertension, cancer, [any other disease related to eye vascular network]. For instance the notion of vascular remodeling has been a very important topic in cancer and there have been many discussion of enhancing the function of vascular networks before some treatments (like chemotherapy or radiotherapy). There are at least two mechanisms under which the vascular network can affect the treatment of cancer: 1) delivery of the drugs, 2) the response of the tumor the vasculature. In the case of the delivery, an impaired vascular network can lead to impaired delivery of drugs. For instance the chronic hypoxic can affect the reachibility of the tissue of interest by the drug through the vascular network (Or are there any other mechanism leading to this?). For the case of the impaired tumor response to the therapy, the main important thing is that impaired vascular network (for instance the functional shunt) leads to in-efficient oxygenating the tumor microenvironment. It is a well accepted fact that under hypoxia, the therapeutic response becomes worsen about 3 fold for radiotherapy. In the case of the chemo-therapy, there are some studies (which when why) that are showing that the response of tumor to chemotherapy becomes impaired.

	\section{Useful Research Tools}
	In this section we will review some of the important software tools and also open source available datasets that can be used to enhance the research.
	\subsection{Open Microvascular Network Data Sets}
	Often, finding appropriate date to validate a model is a challenging step for the research. Fortunately, due to the growth of the machine learning and deep learning and computer vision community, there are very well organized and large data sets of vascular networks that can be very useful.
	\begin{itemize}
		\item High-Resolution Fundus (HRF) Dataset: Contains retinal fundus images.
		\item ROSE Dataset: Contains retinal OCT-A segmented data.
		\item OCTAGON Dataset: Angiography by OCT-A of foveal avascular zone (FAZ)
		\item VESSEL GRAPH Dataset: Whole brain vessel graphs.
		\item MiniVess Dataset: rodent cerebrovasculature images.
		\item STARE dataset: Stands for STructured Analysis of the Retina
		\item 2-PM Vessel Dataset: Volumetric brain vasculature dataset from two photon microscopy
	\end{itemize}

	\subsection{Open Source Software Solutions}
\end{document}

### 2024-09-10 ReviewPaper_Angiogenesis/main.tex
new notes added

@@
 \usepackage{nameref}
-\usepackage{hyperref}
+\usepackage[hidelinks]{hyperref}
+\hypersetup{
+	colorlinks=true,
+	citecolor=cyan!70!black,
+	linkcolor=cyan!90!black
+}
+\usepackage{natbib} 
+%\setcitestyle{authoryear, open={([},close={)]}}
+%\setcitestyle{numbers,square}
+\bibliographystyle{dinat}
+
 \title{On the Mathematical Formulation of the Angiogenesis Process in Health and Disease}
@@
 The implementations of some of the discussed models are shared, as well as sharing some mini-scripts demonstrating use cases of some of the software tools we have described in the paper.
+In the last 20 years of increased research interest in the optimal transport networks in biological systems, there have been many approaches towards this problem, where some of which share same mathematical structure.
+Here in this paper we will develop the common mathematical formulation of the problem and discuss the results.
+		For the theory of optimal transport networks, since they are abundant in nature and applications, there have been many different modelings for different systems (like the river basins, leaf venation, slime mold, etc).
+One purpose of this review paper is the give a unified picture of the relevance of such attempts in the realm of micro-vascular systems, and cancer.
 	\end{abstract}
@@
 	\subsection{Open Source Software Solutions}
+
+	\section{Construction Zone!}
+	This is a temporally construction zone to keep the information and details of some of the papers and them move them to the appropriate place in the text.
+
+	\subsection{Adaptation and optimization of biological transport networks}
+	\label{Hu_Work}
+	Paper by \cite{Hu2013}.
+The functional relation between $ Q $ (flow) and $ \Delta P $ the pressure difference between the two adjacent nodes given as $ Q = C \Delta P $ is because of the low Reynolds number microvasculature.
+The vessel conductance
+	\[ C = \frac{\pi D^4}{128 \eta L}. \]
+	The model is originally derived for the leaf veins, hence they assume no dependence of the viscosity ($ \eta $) on the diameter of the tube. The energy of the network defined as
+	\[ E = \sum_i\left(\frac{Q_i^2}{\tilde{C}_i} + c_0 \tilde{C}_i^\gamma\right) L_i \]
+	where $ c_0 $ is the metabolic rate, $ \gamma $ an intrinsic factor of the network, $ \tilde{C} = C L $ conductivity (note that $ C $ is the conductance).
+The \textbf{first term} of the energy functional above is the energy required to pump the fluid through the network (this is of the form current times potential difference).
+To understand the rationale behind the \textbf{second term} of the energy functional, we evaluate the case $ \gamma = 1/2 $.
+In the Murray's work \cite{Murray1926}, authors postulates that the vascular networks remodel in a way to minimize their total volume (hence lower total number of the cells building the network, lower the metabolic cost).
+When $ \gamma = 1/2 $, then
+	\[ \tilde{C}_i^{1/2}L_i \propto D^2 L_i, \]
+	i.e.
+when $ \gamma=1/2 $, then the metabolic cost is precisely proportional to the volume of the network.
+Thus $ \gamma $ generalizes the Murray's law for vascular networks.
+Furthermore, when $ \gamma = 1/4 $, then the metabolic rate is proportional to the surface area of the network and minimizing this energy functional is the same as minimizing the total surface area of the network.
+Note that \cite{Durand2006} also has a kind of generalization between surface area and volume.
+
+	The authors mention that the metabolic term is not a constrain in this formulation, and $ c_0 $ is not the Lagrange multiplier, but a measurable quantity.
+I need to figure out the formulation by writing an explicit constraint and using the Lagrange multiplier method to optimize the objective.
+\cite{Durand2006,Bohn2007} are using this formulation (optimization by the use of Lagrange multipliers).
+Furthermore\cite{Almeida2022} uses some form of Lagrange multipliers in a different setting that is worth exploration.
+
+	The authors propose the following ODE that governs the governs the time evolution of the conductance of the links. 
+	\[ \frac{d\tilde{C}_i}{dt} = c\left(\underbrace{\frac{Q_i^2}{\tilde{C}_i^{\gamma+1}}}_{\text{prop to }\tau_w^2} - \tilde{\tau_e}^2\right) \tilde{C}_i, \]
+	where $ \tilde{\tau} $ is the target shear stress that we want to have in the tube.
+Here in this section we will highlight the rationale behind this design.
+First, as derived in \cite{Secomb2012}, the shear stress on the vascular wall in the case of the blood networks is given as
+	\[ \tau_w \propto \frac{Q}{D^3}. \]
+	On the other hand since $ \tilde{C} \propto D^4 $, then for $ \gamma = 1/2 $ the first term in the parenthesis in the RHS of the ODE will become
+	\[ \frac{Q_i^2}{\tilde{C_i}^{3/2}} \propto \frac{Q_i^2}{D^6} \propto (\tau_w)^2.  \]
+	Furthermore, authors claim that for this choice of evolution law for the conductivity, $ dE/dt < 0 $.
+\textbf{I am still trying to figure this out}.
+However, the time derivative of the energy functional is
+	\[ \frac{dE}{dt} = \sum_i\left(-\frac{Q_i^2}{\tilde{C}^2_i} + c_0\gamma \tilde{C}_i^{\gamma-1}\right) L_i \frac{d\tilde{C}_i}{dt}. \]
+
+	\section{Adaptive Hagen-Poiseuille flows on graphs}
+	In this part, I will summarize the methods and the results of the paper \cite{Almeida2022}, and the corresponding theses came out of the same research group that contains more details \cite{Valente2023,Almeida2023}.
+The Hagen-Poiseuille equation is for the steady-state laminar flow of incompressible, Newtonian fluid through a \textbf{long} channel \footnote{Note that the blood is not a Newtonian fluid and experiences shear thickening (or shear thinning) under different shear rates.
+But there are certain limits for the shear rate over which we can assume that the blood is a Newtonian fluid.
+For instance, at a very low shear stress, the viscosity of blood increases a lot at lower tube diameters as the RBC can not squeeze through narrow channels.}.
+The Hagen-Poiseuille flow is given by
+	\[ Q_{ij} = \frac{\pi r_{ij}^4}{8\eta L_{ij}}(p_i - p_j). \]
+	For a derivation of this equation from the first principles see \cite{tecscience2020}.
+The choices for adaptation rule can be due to \cite{Tero2010} where Tero et al propose adaptation rule of the type
+	\[ \frac{d D_{ij}}{dt} = f(|Q_{ij}|) - D_{ij}, \]
+	where the first term is for the increase of conductance of the tube due to flow and the second term is the baseline shrinkage of the vessel segment.
+As suggested by Tero et al, one possible choice for $ f(|Q_{ij}|) $ is
+	\[ f(|Q_{ij}|) = \frac{|Q_{ij}|^\gamma}{1+|Q_{ij}|^\gamma} \]
+	Similar to what we derived at \autoref{Hu_Work}, the total volume of the network is
+	\[ V = \beta \sum_{(i,j)\in E} \sqrt{D_{ij}}L_{ij}. \]
+	Note that the variable $ D_{ij} $ represents the conductivity here whereas in \autoref{Hu_Work} it is used for the diameter. By using the chain rule we can write
+	\[ \frac{d V}{dt} \propto \sum_{(i,j)\in E} \frac{L_{ij}}{2\sqrt{D_{ij}}} \frac{dD_{ij}}{dt} \]
+	where we choose $ d D_{ij}/dt $ according to the ODE governing the time evolution of $ D_{ij} $.
+Note that not every choice of the RHS of this ODE will guarantee that $ dV/dt = 0 $, and thus the volume is conserved.
+\cite{Almeida2023} claims that since the choice of $ f $ by \cite{Tero2010}, does not conserve the volume, thus it violates the fact that the fluid is incompressible.
+I do not think this is true.
+The volume of the network can shrink, and this has nothing to do with the compression of the fluid inside.
+Smaller network carries less fluid.
+
+	\section{Work by Karen Alim}
+
+	\section{Work by Carl D Modes group}
+
+	\section{Work by Katifori group}
+
+	\section{Work by Markovich group}
+
+	\section{Work bCaterina De Bacco group}
+
+	\newpage
+	\bibliography{References/references}
 \end{document}

### 2024-09-10 ReviewPaper_Angiogenesis/main.tex
simulation added

@@
 The Hagen-Poiseuille flow is given by
-	\[ Q_{ij} = \frac{\pi r_{ij}^4}{8\eta L_{ij}}(p_i - p_j). \]
+	\[ Q_{ij} = \frac{\pi r_{ij}^4}{8\eta L_{ij}}(p_i - p_j) = \frac{D_{ij}}{L_{ij}}(p_i - p_j). \]
 	For a derivation of this equation from the first principles see \cite{tecscience2020}.
@@
 
+	The core of the model discussed in \cite{Almeida2023} is an undirected, planar, and connected graph $ G(V,E) $ that is embedded in Euclidean space.
+This models is originally developed for the 2D vein system of leaf.
+But in the case of blood flow, it should be a 3D graph.
+Also, they have assumed that the edges are straight.
+The convention they have used here is that $ Q_{ij}>0 $ means a flow from node $ i $ to $ j $, and vice versa.
+The sources and the sinks are located at fixed nodes.
+In the case of blood networks, the sinks are the cells distributed through the domain.
+We need to figure out some way to model this effectively as sinks on the nodes.
+The conservation law reads
+	\[ \sum_{j\in\mathcal{N}(i)}Q_{ij} = s_i. \]
+	\textbf{Local vs.
+Global Mechanisms:} The authors argue that the decision to conserve the volume through the dynamics is considered a global mechanism.
+That is because if some of the vessels shrink due to the lack of flow, then some other vessel should necessarily expand to keep the total volume conserved.
+They argue that they mechanism proposed by \cite{Tero2010} is a local dynamic.
+
+	Since the adaptation of each vessel segment is due to the change it its cross section (radius) and since $ r_{ij} \propto \sqrt{D_{ij}} $, thus we write the ODE for $ \sqrt{D_{ij}} $.
+One initial suggestion is
+	\[ \frac{d}{dt} \sqrt{D_{ij}} = f(Q) - \mu \sqrt{D_{ij}},\]
+	where $ f $ is an unknown function with $ f(0) = 0 $. Then we will need to make sure that this adaptation rule conserves the total volume of the network.
+
+	To find the pressures on the nodes, given the values of the sources/sinks on the nodes, as well as the conductance of the links, for each node $ i $ we need to solve
+	\[ \sum_{j\in\mathcal{N}(i)}Q_{ij} = \sum_{j\in\mathcal{N}(i)}\frac{C_{ij}}{L_{ij}}(p_i - p_j) = s_i, \]
+	which is a kind of Poisson equation on the graph. This can be formulated as
+	\[ L \mathbf{P} = \mathbf{q}, \]
+	where $ L = D - A $ is the Laplacian matrix of the graph, where $ D $ is the degree matrix and $ A $ is the adjacency matrix.
+
+	\newpage
 	\section{Work by Karen Alim}

### 2024-10-03 ReviewPaper_Angiogenesis/main.tex
new updates added

@@
 	\section{Work by Katifori group}
+	In this research group there are multiple papers that most of them share the basics.
+The note \cite{KatiforiJune2019} covers most of the important basics for flow networks in biology.
+The paper by \cite{Ronellenfitsch2016} authors discuss the role of growth in distribution networks.
+They include the growth effectively by a scaling factor and assume that the distance between two points grow as $ d(t) = \lambda_t d(0) $.
+Furthermore, they assume that all of the dynamical quantities scale like $ \square(t) = \lambda_t^\sigma \square'(t) $, where $ \square $ can be any dynamical quantities in the model like pressure of the nodes $ \mathbf{P} $, and flow on the edges $ \mathbf{F} $.
+Note that the exponent $ \sigma $ depends on the dynamical quantity in $ \square $.
+
+	They define a weighted graph (where weights are the conductance of the edges) and define the pressures on the nodes and the flow on the edges, related to each other by
+	\[ Q_e = C_e (p_i - p_j)/L_e \]
+	where $ Q_e, C_e, $ and $ L_e $ are the flow, conductance, and the length of the edge $ e $ respectively, and $ p_i, p_j $ are the pressure of the adjacent nodes to the edge $ e $.
+This can be written more compactly as
+	\[ \mathbf{Q} = \mathbf{C}\mathbf{L}^{-1}\Delta \mathbf{P} \]
+	where $ \mathbf{C} $, and $ \mathbf{L} $ are the diagonal matrices of the conductance and length of edges, and $ \Delta $ is the transpose of the incidence matrix.
+See \cite{FeleParanj2024} for more details.
+The Kirchhoff equation that they solve is
+	\[ \Delta^T \mathbf{Q} = \mathbf{q}  \]
+	where $ \mathbf{q} $ is the source or sink terms on the nodes. Solving this is equivalent to solve the Poisson-like equation
+	\[ \mathcal{L} \mathbf{P} = \mathbf{q}, \]
+	with the connection that $ \mathcal{L} = \Delta ^T  \mathbf{C}\mathbf{L}^{-1} \Delta  $. The evolution law they use for the conductances is
+	\[ \frac{d C_e}{dt} = a (Q_e/\hat{Q})^{2\gamma} - b C_e + c, \]
+	where $ \hat{Q} $ is a reference flow, and $ a,\gamma, b, c $ are constants.
+By including the growth as a time dependent scaling, they derive the equation for the asymptotic case to study its topology.
+By setting $ C = \lambda_t^{2\gamma\delta} C' $, they write the ODE for asymptotic equations
+	\[ \frac{d C_e'}{dt} = a(Q_e'/\hat{Q})^{2\gamma} - b' C_e' + \lambda_t^{-2\gamma\delta}c. \]
 

### 2024-10-03 Simulations (Shared Folder)/AndersonChaplainCode_JohnNardini/angio_modifiedAli.py

@@ -1,0 +1,355 @@
+import numpy as np
+import cv2
+import os
+import matplotlib.pyplot as plt
+import networkx as nx
+
+# Define the Segment class to represent each vessel segment
+class Segment:
+    def __init__(self, start_point, parent_segment=None):
+        self.start_point = start_point  # Starting position (tuple)
+        self.end_point = start_point    # Updated as the tip cell moves
+        self.parent_segment = parent_segment  # Parent Segment object
+        self.child_segments = []  # List of child Segment objects (branches)
+        self.length = 0.0  # Total length of the segment
+        self.path = [start_point]  # List of positions forming the segment path
+        self.anastomosed = False  # Flag to indicate if the segment has anastomosed
+        self.anastomosed_with = None  # Segment object it anastomosed with
+
+class angio_abm:
+    def __init__(self,
+                 IC='linear',
+                 rho=0.34,
+                 t_final=4.0,
+                 chi=0.38,
+                 chemo_rate='const',
+                 psi=0.5):
+
+        # Parameters
+        self.D = .00035
+        self.alpha = 0.6
+        self.chi = chi
+        self.rho = rho
+        self.beta = 0.05
+        self.gamma = 0.1
+        self.eta = 0.1
+        self.psi = psi
+
+        self.nb_const = 2.5
+        self.chemo_rate = chemo_rate
+
+        # Grids
+        self.eps1 = 0.45
+        self.eps2 = 0.45
+        self.k = 0.75
+        self.nu = (np.sqrt(5) - 0.1)/(np.sqrt(5)-1)
+
+        self.xn = 201
+        self.yn = 201
+        self.x = np.linspace(0, 1, self.xn)
+        self.y = np.linspace(0, 1, self.yn)
+        self.IC = IC
+        self.dx = self.x[1] - self.x[0]
+        self.dy = self.y[1] - self.y[0]
+
+        self.dt = 0.01
+        self.t_final = t_final
+        self.time_grid = np.arange(0, self.t_final, self.dt)
+
+        self.write_folder = 'results/'
+        self.file_name = f'IC_{self.IC}_rho_{round(self.rho, 2)}_chi_{round(self.chi, 2)}'
+
+        if not os.path.isdir(self.write_folder):
+            os.mkdir(self.write_folder)
+
+        # Initialize sprouts
+        self.cell_locs = [[0, .1], [0, .2], [0, .3], [0, .4], [0, .5],
[…]

### 2024-10-03 Simulations (Shared Folder)/NeuronBranching_CanUcarPaper/FullPythonImplementation/main.py

@@ -1,0 +1,464 @@
+import matplotlib.pyplot as plt
+from numpy import linalg as LA
+from pylab import *
+import networkx as nx
+from scipy.spatial import KDTree
+
+
+
+
+class BARW:
+    def __init__(self,tmax):
+        self.G = nx.Graph()
+        self.coordinates = None
+        self.evolve = None
+        self.diff = None
+        self.angle_list = None
+        self.angle = None
+        self.angle_init = None
+        self.node = None
+        self.current_edge = {}
+        self.active_tips = {}  # Track current edges for each active tip
+        self.branch_counter = 0  # To uniquely identify branches
+        self.coordsTemp = None
+
+        self.branchID_TO_StartNode = {} # Give the branchID as the key, it will give you the StartNode
+        self.branchID_TO_endNode = {}
+
+
+        ## Parameters
+        self.Lx = 200
+        self.Lz = 300
+        self.tmax = tmax
+
+        self.lstep = 1                  # Elongation length (step size)
+        self.min_branch = pi / 10       # minimum branching angle
+        self.cutoff = 1.5 * self.lstep  # cutoff length for annihilation via tip-node contact
+        self.radavoid = 10 * self.lstep # Radius and strength of attraction/repulsion (if fav<0, then attraction)
+
+
+
+    def initialize(self):
+        self.node = np.array([[0, self.Lz / 2, 0, 1]])      # [x,y,genID, branchID] for the tip cells only
+        self.angle_init = [0.0]
+        self.angle = np.array(self.angle_init)              # angle for the tip cells only
+        self.angle_list = np.array([[0, 1]])                # [angle, branchID] TODO: the second entry should be 1?
+        self.diff = 0
+        self.evolve = np.array([len(self.node)])
+        self.coordinates = np.array([self.node[0]])
+
+        # initial_point = (self.node[0][0], self.node[0][1])
+        initial_point = tuple(self.node[0][:2])
+        self.G.add_node(initial_point)  # Add the initial point as the first node
+        self.branchID_TO_StartNode[1] = initial_point
+        # self.current_edge[0] = initial_point  # Track the current starting point of the first tip's branch
+
+    def distance(self, vect, tip):
+        diff = np.add(vect[:2], -tip[:2])
+        distance = LA.norm(diff)
+        return distance
+
+
+    def prob_list(self, x, f, pb):
+        # bias for forward steps
+        alpha = np.clip((1 - f * np.sin(x)) / 2, 0, 1)
+        # bias for backward steps
[…]

### 2024-10-03 Simulations (Shared Folder)/NeuronBranching_CanUcarPaper/FullPythonImplementation/run.py

@@ -1,0 +1,37 @@
+from pylab import *
+from tqdm import tqdm
+
+
+from main import BARW
+
+tmax = 200
+prob_branch = 0.03
+fav = -0.1*(0)## -0.1
+fchem = 0.6*0.5
+
+barw = BARW(tmax)
+barw.initialize()
+
+
+for t in tqdm(range(tmax), desc="Simulation Progress"):
+    t+=1
+    if t<5 and len(barw.node)>1:
+        break
+    if len(barw.node)!=0:
+        barw.tissue1(prob_branch,fav,fchem)
+    if len(barw.node)==0:
+        break
+
+barw.remove_arcs()
+barw.plotter_basic()
+barw.plot_graph()
+
+
+
+
+## Useful DEBUG codes
+# plt.scatter(self.coordinates[:,0], self.coordinates[:,1])
+# plt.show()
+
+
+# self.coordinates[:52,3] = np.where(self.coordinates[:52,3]==5, 10, self.coordinates[:52,3])

### 2024-10-04 Simulations (Shared Folder)/NeuronBranching_CanUcarPaper/FullPythonImplementation/agentBased/Duct.py
new code added

@@ -1,0 +1,15 @@
+from Tip import Tip
+import numpy as np
+
+class Duct:
+    def __init__(self,initPos, initVel, network):
+        self.tip = Tip(initPos,initVel, self)
+        self.tailCurve = []
+        self.tailCurve.append(self.tip.position.tolist())
+        self.network = network
+
+    def update(self):
+        if self.tip.isAlive:
+            self.tip.update()
+            self.tailCurve.append(self.tip.position.tolist())
+

### 2024-10-04 Simulations (Shared Folder)/NeuronBranching_CanUcarPaper/FullPythonImplementation/agentBased/Network.py

@@ -1,0 +1,31 @@
+from Duct import Duct
+import numpy as np
+from utils import rotMat
+
+class Network:
+    def __init__(self):
+        self.ductsList = []
+        self.ductsList.append(Duct(np.array([0.0,0.0]),np.array([1.0,0.0]), self))
+
+        self.ductsToBeAdded = []
+
+
+    def update(self):
+        for ducts in self.ductsList:
+            ducts.update()
+
+        for ducts in self.ductsToBeAdded:
+            self.ductsList.append(ducts)
+
+        self.ductsToBeAdded = []
+
+    def handleBranching(self, initPos):
+        v = np.array([1.0,0.0])
+        child1_angle = np.random.uniform(np.pi/4, np.pi/2)
+        child2_angle = np.random.uniform(-np.pi/2, -np.pi/4)
+
+        self.ductsToBeAdded.append(Duct(initPos,np.matmul(rotMat(child1_angle),v),self))
+        self.ductsToBeAdded.append(Duct(initPos,np.matmul(rotMat(child2_angle),v),self))
+        # self.ductsToBeAdded.append(Duct(np.array([0.0,0.0]),np.matmul(rotMat(child1_angle),v),self))
+        # self.ductsToBeAdded.append(Duct(np.array([0.0,0.0]),np.matmul(rotMat(child2_angle),v),self))
+

### 2024-10-04 Simulations (Shared Folder)/NeuronBranching_CanUcarPaper/FullPythonImplementation/agentBased/Tip.py

@@ -1,0 +1,23 @@
+import numpy as np
+from utils import rotMat
+
+
+class Tip:
+    def __init__(self, initPos, initVel, duct):
+        self.position = np.copy(initPos)
+        self.velocity = np.copy(initVel) ## The absolute value of the velocity
+        self.isAlive = True
+        self.duct = duct
+
+        self.dt = 0.01
+
+    def update(self):
+        theta = np.random.uniform(-np.pi / 10, np.pi / 10)
+        self.velocity = np.matmul(rotMat(theta), self.velocity)
+        self.position += self.dt * self.velocity
+
+        if np.random.random() < 0.01:
+            self.duct.network.handleBranching(np.copy(self.position))
+            self.isAlive = False
+
+

### 2024-10-04 Simulations (Shared Folder)/NeuronBranching_CanUcarPaper/FullPythonImplementation/agentBased/main.py

@@ -1,0 +1,31 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from agentBased.Network import Network
+
+
+
+class Main:
+    def __init__(self):
+        self.network = Network()
+
+    def run(self):
+        for i in range(100):
+            self.network.update()
+
+    def plot(self):
+        for duct in self.network.ductsList:
+            coord = np.array(duct.tailCurve)
+            plt.plot(coord[:,0],coord[:,1])
+        plt.axis('equal')
+        plt.show()
+
+
+
+
+if __name__ == '__main__':
+    main = Main()
+    main.run()
+    main.plot()
+
+
+

### 2024-10-04 Simulations (Shared Folder)/NeuronBranching_CanUcarPaper/FullPythonImplementation/main_OLD.py

@@ -319,9 +319,9 @@ class BARW:
         #     self.angle_list = np.append(self.angle_list, np.array(angleEntryList),axis=0)
 
         self.angle_list = np.append(self.angle_list, np.column_stack((np.degrees(self.angle), self.node[:, -1])), axis=0)
-        self.coordsTemp = np.append(self.coordsTemp, np.array(self.node), axis=0)
-        self.coordinates = np.copy(self.coordsTemp)
-        # self.coordinates = np.append(self.coordinates, np.array(self.node), axis=0)
+        # self.coordsTemp = np.append(self.coordsTemp, np.array(self.node), axis=0)
+        # self.coordinates = np.copy(self.coordsTemp)
+        self.coordinates = np.append(self.coordinates, np.array(self.node), axis=0)
 
 
 

### 2024-10-04 Simulations (Shared Folder)/NeuronBranching_CanUcarPaper/FullPythonImplementation/run.py

@@ -2,7 +2,7 @@ from pylab import *
 from tqdm import tqdm
 
 
-from main import BARW
+from main_OLD import BARW
 
 tmax = 200
 prob_branch = 0.03

### 2024-10-04 Simulations (Shared Folder)/NeuronBranching_CanUcarPaper/FullPythonImplementation/agentBased/Network.py
new notes added

@@ -21,8 +21,13 @@ class Network:
 
     def handleBranching(self, initPos):
         v = np.array([1.0,0.0])
-        child1_angle = np.random.uniform(np.pi/4, np.pi/2)
-        child2_angle = np.random.uniform(-np.pi/2, -np.pi/4)
+        # child1_angle = np.random.uniform(np.pi/4, np.pi/2)
+        # child2_angle = np.random.uniform(-np.pi/2, -np.pi/4)
+        #
+        child1_angle = np.random.randn()*np.pi/100.0 + np.pi/8
+        child2_angle = np.random.randn()*np.pi/100.0 - np.pi/8
+
+
 
         self.ductsToBeAdded.append(Duct(initPos,np.matmul(rotMat(child1_angle),v),self))
         self.ductsToBeAdded.append(Duct(initPos,np.matmul(rotMat(child2_angle),v),self))

### 2024-10-04 Simulations (Shared Folder)/NeuronBranching_CanUcarPaper/FullPythonImplementation/agentBased/Tip.py

@@ -12,12 +12,18 @@ class Tip:
         self.dt = 0.01
 
     def update(self):
-        theta = np.random.uniform(-np.pi / 10, np.pi / 10)
+        self.move()
+        self.branch()
+
+
+
+    def move(self):
+        theta = np.random.randn() * np.pi / 50
         self.velocity = np.matmul(rotMat(theta), self.velocity)
         self.position += self.dt * self.velocity
 
+    def branch(self):
         if np.random.random() < 0.01:
             self.duct.network.handleBranching(np.copy(self.position))
             self.isAlive = False
-
 

### 2024-10-05 Simulations (Shared Folder)/NeuronBranching_CanUcarPaper/FullPythonImplementation/agentBased/Network.py
new code

@@ -14,23 +14,19 @@ class Network:
         for ducts in self.ductsList:
             ducts.update()
 
-        for ducts in self.ductsToBeAdded:
+        for ducts in self.ductsToBeAdded: ## The newly created ducts (due to branching) are not added to the
+            # duct list right away as it can lead to buggy behaviour (since the loop is happening on the duct list and
+            # it should not be altered)
             self.ductsList.append(ducts)
 
-        self.ductsToBeAdded = []
+
+        self.ductsToBeAdded = []  ## To make sure that the ducts are added and the list is empty for the next round
 
     def handleBranching(self, initPos):
         v = np.array([1.0,0.0])
-        # child1_angle = np.random.uniform(np.pi/4, np.pi/2)
-        # child2_angle = np.random.uniform(-np.pi/2, -np.pi/4)
-        #
         child1_angle = np.random.randn()*np.pi/100.0 + np.pi/8
         child2_angle = np.random.randn()*np.pi/100.0 - np.pi/8
-
-
-
         self.ductsToBeAdded.append(Duct(initPos,np.matmul(rotMat(child1_angle),v),self))
         self.ductsToBeAdded.append(Duct(initPos,np.matmul(rotMat(child2_angle),v),self))
-        # self.ductsToBeAdded.append(Duct(np.array([0.0,0.0]),np.matmul(rotMat(child1_angle),v),self))
-        # self.ductsToBeAdded.append(Duct(np.array([0.0,0.0]),np.matmul(rotMat(child2_angle),v),self))
+
 

### 2024-10-05 Simulations (Shared Folder)/myAgentBasedImplementation/agentBased/Duct.py

@@ -1,0 +1,17 @@
+from Tip import Tip
+import numpy as np
+
+class Duct:
+    def __init__(self,initPos, initVel, network, parent):
+        self.network = network
+        self.parent = parent
+        self.tailCurve = []
+        self.collisionList = [] # Contains the tips that collided with this duct
+        self.tip = Tip(initPos, initVel, self)
+        self.tailCurve.append(self.tip.position.tolist())
+
+    def update(self):
+        if self.tip.isAlive:
+            self.tip.update()
+            self.tailCurve.append(self.tip.position.tolist())
+

### 2024-10-05 Simulations (Shared Folder)/myAgentBasedImplementation/agentBased/Network.py

@@ -1,0 +1,115 @@
+from numpy import number
+
+from Duct import Duct
+import numpy as np
+from utils import rotMat
+import matplotlib.pyplot as plt
+
+from scipy.spatial import KDTree
+
+class Network:
+    def __init__(self):
+        self.ductsList = []
+        self.ductsToBeAdded = []  ## will add the newly generated ducts (due to branching) after when iteration on the
+        # ductsList to update the ducts is completed.
+        self.dt = 0.01
+        self.vNorm = 1.0 # the absolute value of the velocity. Note that since the velocity will be changing by a rotation
+        # matrix, then its norm will stay the same throughout
+        self.pointToDuct = {}
+        self.r_c = (self.vNorm * self.dt) * 1 # the r collision should be larger than the maximum step size of the walkers
+
+        self.initialize()
+
+    def initialize(self):
+        self.ductsList.append(Duct(np.array([0.0,0.0]),np.array([self.vNorm,0.0]), self, None))
+
+
+    def update(self):
+        self.collisionDetection()
+        for ducts in self.ductsList:
+            ducts.update()
+
+        for ducts in self.ductsToBeAdded: ## The newly created ducts (due to branching) are not added to the
+            # duct list right away as it can lead to buggy behaviour (since the loop is happening on the duct list and
+            # it should not be altered)
+            self.ductsList.append(ducts)
+
+
+
+        self.ductsToBeAdded = []  ## To make sure that the ducts are added and the list is empty for the next round
+
+    def handleBranching(self, initPos, parent):
+        v = parent.tip.velocity
+        # child1_angle = np.random.randn()*np.pi/100.0 + np.pi/8
+        # child2_angle = np.random.randn()*np.pi/100.0 - np.pi/8
+        child1_angle = np.pi/4
+        child2_angle = -np.pi/4
+        self.ductsToBeAdded.append(Duct(initPos,np.matmul(rotMat(child1_angle),v),self, parent))
+        self.ductsToBeAdded.append(Duct(initPos,np.matmul(rotMat(child2_angle),v),self, parent))
+
+
+    def collisionDetection(self):
+        """
+        ** This function will detect the collisions that will happen if we take the next steps for tips.
+        For this we are using the fact that the radius of the collision detection should be larger than the
+        maximum distance any tip can travel at each time step.
+
+        * effect: Will label the tips who will annihilate (collide) with collision flag and also give it the address of the
+        collision (the corresponding duct object that causes the collision as well as the position of the collision).
+
[…]

### 2024-10-05 Simulations (Shared Folder)/myAgentBasedImplementation/agentBased/Tip.py

@@ -1,0 +1,36 @@
+import numpy as np
+from utils import rotMat
+
+
+class Tip:
+    def __init__(self, initPos, initVel, duct):
+        self.position = np.copy(initPos)
+        self.velocity = np.copy(initVel) ## The absolute value of the velocity
+        self.collisionData = None ## this will be replaced with collision information if collides: [Stopperduct, stopperPoint]
+        self.isAlive = True
+        self.duct = duct
+
+        self.dt = self.duct.network.dt
+
+    def update(self):
+        if self.collisionData == None: ## Is not colliding
+            self.move()
+            self.branch()
+        else:
+            stopperDuct = self.collisionData[0]
+            stopperPoint = self.collisionData[1]
+            stopperDuct.collisionList.append(self)
+            self.position = stopperPoint
+            self.isAlive = False
+
+
+    def move(self):
+        theta = np.random.randn() * np.pi / 20
+        self.velocity = np.matmul(rotMat(theta), self.velocity)
+        self.position += self.dt * self.velocity
+
+    def branch(self):
+        if np.random.random() < 0.02:
+            self.duct.network.handleBranching(self.position, self.duct)
+            self.isAlive = False
+

### 2024-10-05 Simulations (Shared Folder)/myAgentBasedImplementation/agentBased/main.py

@@ -1,0 +1,35 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from agentBased.Network import Network
+
+
+
+class Main:
+    def __init__(self):
+        self.network = Network()
+
+    def run(self):
+        for i in range(800):
+            self.network.update()
+
+    def plot(self):
+        for duct in self.network.ductsList:
+            coord = np.array(duct.tailCurve)
+            # plt.scatter(coord[:,0],coord[:,1],s=2,c="k")
+            plt.plot(coord[:,0],coord[:,1],color='steelblue')
+            if duct.tip.isAlive:
+                plt.plot(duct.tip.position[0],duct.tip.position[1],'o',color="orange",ms=3)
+        plt.axis('equal')
+        plt.show()
+
+
+
+
+if __name__ == '__main__':
+    main = Main()
+    main.run()
+    main.plot()
+    # main.network.collisionDetection()
+
+
+

### 2024-10-05 Simulations (Shared Folder)/myAgentBasedImplementation/agentBased/utils.py

@@ -1,0 +1,14 @@
+import numpy as np
+
+def rotMat(theta):
+    """
+    Returns the 2D rotation matrix for a given angle theta (in radians).
+
+    Parameters:
+    theta (float): The rotation angle in radians.
+
+    Returns:
+    numpy.ndarray: A 2x2 rotation matrix.
+    """
+    return np.array([[np.cos(theta), -np.sin(theta)],
+                     [np.sin(theta), np.cos(theta)]])

### 2024-10-05 Simulations (Shared Folder)/myAgentBasedImplementation/initAttempt/main_OLD.py

@@ -1,0 +1,464 @@
+import matplotlib.pyplot as plt
+from numpy import linalg as LA
+from pylab import *
+import networkx as nx
+from scipy.spatial import KDTree
+
+
+
+
+class BARW:
+    def __init__(self,tmax):
+        self.G = nx.Graph()
+        self.coordinates = None
+        self.evolve = None
+        self.diff = None
+        self.angle_list = None
+        self.angle = None
+        self.angle_init = None
+        self.node = None
+        self.current_edge = {}
+        self.active_tips = {}  # Track current edges for each active tip
+        self.branch_counter = 0  # To uniquely identify branches
+        self.coordsTemp = None
+
+        self.branchID_TO_StartNode = {} # Give the branchID as the key, it will give you the StartNode
+        self.branchID_TO_endNode = {}
+
+
+        ## Parameters
+        self.Lx = 200
+        self.Lz = 300
+        self.tmax = tmax
+
+        self.lstep = 1                  # Elongation length (step size)
+        self.min_branch = pi / 10       # minimum branching angle
+        self.cutoff = 1.5 * self.lstep  # cutoff length for annihilation via tip-node contact
+        self.radavoid = 10 * self.lstep # Radius and strength of attraction/repulsion (if fav<0, then attraction)
+
+
+
+    def initialize(self):
+        self.node = np.array([[0, self.Lz / 2, 0, 1]])      # [x,y,genID, branchID] for the tip cells only
+        self.angle_init = [0.0]
+        self.angle = np.array(self.angle_init)              # angle for the tip cells only
+        self.angle_list = np.array([[0, 1]])                # [angle, branchID] TODO: the second entry should be 1?
+        self.diff = 0
+        self.evolve = np.array([len(self.node)])
+        self.coordinates = np.array([self.node[0]])
+
+        # initial_point = (self.node[0][0], self.node[0][1])
+        initial_point = tuple(self.node[0][:2])
+        self.G.add_node(initial_point)  # Add the initial point as the first node
+        self.branchID_TO_StartNode[1] = initial_point
+        # self.current_edge[0] = initial_point  # Track the current starting point of the first tip's branch
+
+    def distance(self, vect, tip):
+        diff = np.add(vect[:2], -tip[:2])
+        distance = LA.norm(diff)
+        return distance
+
+
+    def prob_list(self, x, f, pb):
+        # bias for forward steps
+        alpha = np.clip((1 - f * np.sin(x)) / 2, 0, 1)
+        # bias for backward steps
[…]

### 2024-10-05 Simulations (Shared Folder)/myAgentBasedImplementation/initAttempt/run.py

@@ -1,0 +1,37 @@
+from pylab import *
+from tqdm import tqdm
+
+
+from main_OLD import BARW
+
+tmax = 200
+prob_branch = 0.03
+fav = -0.1*(0)## -0.1
+fchem = 0.6*0.5
+
+barw = BARW(tmax)
+barw.initialize()
+
+
+for t in tqdm(range(tmax), desc="Simulation Progress"):
+    t+=1
+    if t<5 and len(barw.node)>1:
+        break
+    if len(barw.node)!=0:
+        barw.tissue1(prob_branch,fav,fchem)
+    if len(barw.node)==0:
+        break
+
+barw.remove_arcs()
+barw.plotter_basic()
+barw.plot_graph()
+
+
+
+
+## Useful DEBUG codes
+# plt.scatter(self.coordinates[:,0], self.coordinates[:,1])
+# plt.show()
+
+
+# self.coordinates[:52,3] = np.where(self.coordinates[:52,3]==5, 10, self.coordinates[:52,3])
